### 2021-01-20 multivariate.Rmd
multivariate analysis rmd for Team X

---
title: "Multivariate Analysis"
author: "Scott Coffin"
date: "1/19/2021"
output: html_document
---
# Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load packages
library(tidyverse) #General everything
library(RColorBrewer)
library(ggplot2) #General plotting
library(ggrepel) #For adding text labels that repel away from data points
library(calecopal) #Color palette
library(shiny) #Runs shiny
library(shinythemes) #Shiny theme for the page
library(shinyWidgets) #Widgets
library(scales) #SSD - Use the percent format
library(reshape2) #Overview tab - melts bars together
library(ssdtools) #SSD package
library(DT) #Build HTML data tables
#library(plotly) #Make plots interactive
library(viridis) #Colors
library(scales) #To use "percent" function
library(shinyjs) #Exploration tab - reset button
library(tigerstats) #turns things into percents
library(ggbeeswarm) #plot all points
library(fitdistrplus) #alt SSD 
library(RColorBrewer) #colors
library(pheatmap) #pretty heat maps
library(rpart)  #for trees
#library(rattle)    # Fancy tree plot This is a difficult library to install (https://gist.github.com/zhiyzuo/a489ffdcc5da87f28f8589a55aa206dd) 
library(rpart.plot)             # Enhanced tree plots
library(RColorBrewer)       # Color selection for fancy tree plot
library(party)                  # Alternative decision tree algorithm
# library(partykit)               # Convert rpart object to BinaryTree
library(pROC)   #for ROC curves

library(ISLR)  #for the Carseat Data
```
## Data import
```{r data import}
# Load finalized dataset.
aoc <- read_csv("AquaticOrganisms_Clean_final.csv", guess_max = 10000)

#### Introduction Setup ####

# All text inputs below.

#### Overview AO Setup ####

#Final_effect_dataset <- read_csv("Final_effect_dataset.csv")%>%
  #mutate(plot_f = case_when(
    #plot_f == "Polymer" ~ "Polymer",
    #plot_f == "Size" ~ "Size",
    #plot_f == "Shape" ~ "Shape",
    #plot_f == "Organism" ~ "Organism",
    #plot_f == "Lvl1" ~ "Endpoint Category",
    #plot_f == "Life.stage" ~ "Life Stage",
    #plot_f == "Invivo.invivo" ~ "In Vivo or In Vitro",
    #plot_f == "Exposure.route" ~ "Exposure Route"))%>%
  #mutate(plot_f = factor(plot_f))%>%
  #mutate(logEndpoints = log(Endpoints))%>%
  #rename(Percent = Freq)

polydf<-rowPerc(xtabs( ~polymer +effect, aoc)) #pulls polymers by effect 
polyf<-as.data.frame(polydf)%>% #Makes data frame 
  filter(effect %in% c("Y","N"))%>% #Sorts into Yes and No
  mutate(polymer = case_when(
    polymer == "BIO" ~ "Biopolymer",
    polymer == "EVA" ~ "Polyethylene Vinyl Acetate",
    polymer == "LTX" ~ "Latex",
    polymer == "PA" ~ "Polyamide",
    polymer == "PE" ~ "Polyethylene",
    polymer == "PC" ~ "Polycarbonate",
    polymer == "PET" ~ "Polyethylene Terephthalate",
    polymer == "PI" ~ "Polyisoprene",
    polymer == "PMMA" ~ "Polymethylmethacrylate",
    polymer == "PP" ~ "Polypropylene",
    polymer == "PS" ~ "Polystyrene",
    polymer == "PUR" ~ "Polyurathane",
    polymer == "PVC" ~ "Polyvinylchloride",
    polymer == "PLA" ~ "Polylactic Acid"))%>%
  mutate_if(is.numeric, round,0) #rounds percents 
Endpoints<-xtabs(~polymer +effect ,aoc) #Pulls all study obs. for polymer from dataset
polyfinal<- data.frame(cbind(polyf, Endpoints))%>% #adds it as a column
  rename(Endpoints='Freq.1')%>% #renames column
  mutate(logEndpoints = log(Endpoints))%>%
  rename(Percent = Freq)#renames column

sizedf<-rowPerc(xtabs(~size.category +effect, aoc))
sizef<-as.data.frame(sizedf)%>%
  filter(effect %in% c("Y","N"))%>%
  mutate(size.category = case_when(
    size.category == 1 ~ "<1µm",
    size.category == 2 ~ "1µm < 10µm",
    size.category == 3 ~ "10µm < 100µm",
    size.category == 4 ~ "100µm < 1mm",
    size.category == 5 ~ "1mm < 5mm",
    size.category == 0 ~ "unavailable"))%>%
  rename(Type = "size.category")%>%
  mutate_if(is.numeric, round,0)%>%
  mutate(plot="Size")
study_s<-xtabs(~size.category +effect ,aoc)
sizefinal<- data.frame(cbind(sizef, study_s))%>% 
  rename(Endpoints='Freq.1')%>%
  rename(category='size.category')%>%
  mutate(logEndpoints = log(Endpoints))%>%
  rename(Percent = Freq)#renames column

shapedf<-rowPerc(xtabs(~shape + effect, aoc))
shapef<-as.data.frame(shapedf)%>%
  filter(effect %in% c("Y","N"))%>%
  rename(Type="shape")%>%
  mutate_if(is.numeric, round,0)%>%
  mutate(plot="Shape")%>%
  mutate(Type = case_when(
    Type == "cube" ~ "Cube",
    Type == "sphere" ~ "Sphere",
    Type == "fragment" ~ "Fragment",
    Type == "fiber" ~ "Fiber"))
study_sh<-xtabs(~shape + effect,aoc)
shapefinal<- data.frame(cbind(shapef, study_sh))%>% 
  rename(Endpoints='Freq.1')%>%
  rename(category='shape')%>%
  mutate(logEndpoints = log(Endpoints))%>%
  rename(Percent = Freq)#renames column

taxdf<-rowPerc(xtabs(~organism.group +effect, aoc))
taxf<-as.data.frame(taxdf)%>%
  filter(effect %in% c("Y","N"))%>%
  rename(Type= "organism.group")%>%
  mutate_if(is.numeric, round,0)%>%
  mutate(plot="Organism")
study_t<-xtabs(~organism.group +effect,aoc)
taxfinal<- data.frame(cbind(taxf, study_t))%>% 
  rename(Endpoints='Freq.1')%>%
  rename(category='organism.group')%>%
  mutate(logEndpoints = log(Endpoints))%>%
  rename(Percent = Freq)#renames column

lvl1df<-rowPerc(xtabs(~lvl1 +effect, aoc))
lvl1f<-as.data.frame(lvl1df)%>%
  filter(effect %in% c("Y","N"))%>%
  rename(Type= "lvl1")%>%
  mutate_if(is.numeric, round,0)%>%
  mutate(plot="Lvl1")%>%
  mutate(Type = case_when(
    Type == "alimentary.excretory" ~ "Alimentary, Excretory",
    Type == "behavioral.sense.neuro" ~ "Behavioral, Sensory, Neurological",
    Type == "circulatory.respiratory" ~ "Circulatory, Respiratory",
    Type == "community" ~ "Community",
    Type == "fitness" ~ "Fitness",
    Type == "immune" ~ "Immune",
    Type == "metabolism" ~ "Metabolism",
    Type == "microbiome" ~ "Microbiome",
    Type == "stress" ~ "Stress")) 
study_l<-xtabs(~lvl1 +effect,aoc)
lvl1final<- data.frame(cbind(lvl1f, study_l))%>% 
  rename(Endpoints='Freq.1')%>%
  rename(category='lvl1')%>%
  mutate(logEndpoints = log(Endpoints))%>%
  rename(Percent = Freq)#renames column

lifedf<-rowPerc(xtabs(~life.stage +effect, aoc))
lifef<-as.data.frame(lifedf)%>%
  filter(effect %in% c("Y","N"))%>%
  rename(Type= "life.stage")%>%
  mutate_if(is.numeric, round,0)%>%
  mutate(plot="Life.stage")
studyli<-xtabs(~life.stage +effect ,aoc)
lifefinal<- data.frame(cbind(lifef, studyli))%>% 
  rename(Endpoints='Freq.1')%>%
  rename(category='life.stage')%>%
  mutate(logEndpoints = log(Endpoints))%>%
  rename(Percent = Freq)#renames column

vivodf<-rowPerc(xtabs(~invitro.invivo +effect, aoc))
vivof<-as.data.frame(vivodf)%>%
  filter(effect %in% c("Y","N"))%>%
  rename(Type= "invitro.invivo")%>%
  mutate_if(is.numeric, round,0)%>%
  mutate(plot="Invivo.invivo")%>%
  mutate(Type = case_when(
    Type=="invivo"~"In Vivo",
    Type=="invitro"~"In Vitro"))
study_v<-xtabs(~invitro.invivo +effect,aoc)
vivofinal<- data.frame(cbind(vivof, study_v))%>% 
  rename(Endpoints='Freq.1')%>%
  rename(category='invitro.invivo')%>%
  mutate(logEndpoints = log(Endpoints))%>%
  rename(Percent = Freq)#renames column

routedf<-rowPerc(xtabs(~exposure.route +effect, aoc))
routef<-as.data.frame(routedf)%>%
  filter(effect %in% c("Y","N"))%>%
  rename(Type= "exposure.route")%>%
  mutate_if(is.numeric, round,0)%>%
  mutate(plot="Exposure.route")%>%
  mutate(Type = case_when(
    Type == "coparental.exposure" ~"Co-Parental Exposure",
    Type == "paternal.exposure" ~ "Paternal Exposure",
    Type == "maternal.exposure" ~ "Maternal Exposure",
    Type == "food" ~ "Food",
    Type == "water" ~ "Water",
    Type == "sediment" ~ "Sediment",
    Type == "media" ~ "Media"))
study_r<-xtabs(~exposure.route +effect,aoc)
routefinal<- data.frame(cbind(routef, study_r))%>% 
  rename(Endpoints='Freq.1')%>%
  rename(category='exposure.route')%>%
  mutate(logEndpoints = log(Endpoints))%>%
  rename(Percent = Freq)#renames column

#### Exploration AO Setup ####

# Master dataset for scatterplots - for Heili's tab.
aoc_v1 <- aoc %>% # start with original dataset
   # full dataset filters.
  mutate(effect_f = factor(case_when(effect == "Y" ~ "Yes",
    effect == "N" ~ "No"),
    levels = c("No", "Yes"))) %>%
  # removing NAs to make data set nicer
  replace_na(list(size.category = 0, shape = "Not Reported", polymer = "Not Reported", life.stage = "Not Reported"))

aoc_setup <- aoc_v1 %>% # start with original dataset
  mutate(size_f = factor(case_when(
    size.category == 1 ~ "1nm < 100nm",
    size.category == 2 ~ "100nm < 1µm",
    size.category == 3 ~ "1µm < 100µm",
    size.category == 4 ~ "100µm < 1mm",
    size.category == 5 ~ "1mm < 5mm",
    size.category == 0 ~ "Not Reported"),
    levels = c("1nm < 100nm", "100nm < 1µm", "1µm < 100µm", "100µm < 1mm", "1mm < 5mm", "Not Reported"))) %>% # creates new column with nicer names and order by size levels.
  # shape category data tidying.
  mutate(shape_f = factor(case_when(
    shape == "fiber" ~ "Fiber",
    shape == "fragment" ~ "Fragment",
    shape == "sphere" ~ "Sphere",
    shape == "Not Reported" ~ "Not Reported"),
    levels = c("Fiber", "Fragment", "Sphere", "Not Reported"))) %>% # order our different shapes.
  # polymer category data tidying.
  mutate(poly_f = factor(case_when(
    polymer == "BIO" ~ "Biopolymer",
    polymer == "EVA" ~ "Polyethylene Vinyl Acetate",
    polymer == "LTX" ~ "Latex",
    polymer == "PA" ~ "Polyamide",
    polymer == "PE" ~ "Polyethylene",
    polymer == "PC" ~ "Polycarbonate",
    polymer == "PET" ~ "Polyethylene Terephthalate",
    polymer == "PI" ~ "Polyisoprene",
    polymer == "PMMA" ~ "Polymethylmethacrylate",
    polymer == "PP" ~ "Polypropylene",
    polymer == "PS" ~ "Polystyrene",
    polymer == "PUR" ~ "Polyurathane",
    polymer == "PVC" ~ "Polyvinylchloride",
    polymer == "PLA" ~ "Polylactic Acid",
    polymer == "Not Reported" ~ "Not Reported"))) %>%
  # taxonomic category data tidying.
  mutate(org_f = factor(organism.group, levels = c("Algae", "Annelida", "Bacterium", "Cnidaria", "Crustacea",
                                                   "Echinoderm", "Fish", "Insect", "Mollusca", "Nematoda", "Plant", "Rotifera", "Mixed"))) %>% # order our different organisms.
  mutate(lvl1_f = factor(case_when(lvl1 == "alimentary.excretory" ~ "Alimentary, Excretory",
    lvl1 == "behavioral.sense.neuro" ~ "Behavioral, Sensory, Neurological",
    lvl1 == "circulatory.respiratory" ~ "Circulatory, Respiratory",
    lvl1 == "community" ~ "Community",
    lvl1 == "fitness" ~ "Fitness",
    lvl1 == "immune" ~ "Immune",
    lvl1 == "metabolism" ~ "Metabolism",
    lvl1 == "microbiome" ~ "Microbiome",
    lvl1 == "stress" ~ "Stress"))) %>% # creates new column with nicer names.
  # Level 2 Data tidying
  mutate(lvl2_f = factor(case_when(lvl2 == "abundance"~"Abundance",
    lvl2 == "actinobacteria" ~ "Actinobacteria",
    lvl2 == "aggressivity"~"Agressivity",
    lvl2 == "ammonia.excretion" ~ "Ammonia Excretion",
    lvl2 == "bacteroidetes"~ "Bacteriodetes",
    lvl2 == "blood"~"Blood",
    lvl2 == "body.condition"~"Body Condition",
    lvl2 == "boldness"~"Boldness",
    lvl2 == "brain.histo"~"Brain Histological Abnormalities",
    lvl2 == "burrowing"~"Burrowing",
    lvl2 == "carb.metabolism"~"Carb Metabolism",
    lvl2 == "chemokines.cytokines"~"Chemokines",
    lvl2 == "circulatory"~"Circulatory",
    lvl2 == "detoxification"~"Detoxification",
    lvl2 == "development"~"Development",
    lvl2 == "digestion"~"Digestion",
    lvl2 == "digestive.enzymes"~"Digestive Enzymes",
    lvl2 == "digestive.tract.histo"~"Digestive Tract Histological Abnormalities",
    lvl2 == "diversity"~ "Diversity",
    lvl2 == "feeding"~ "Feeding",
    lvl2 == "firmicutes"~ "Firmicutes",
    lvl2 == "gall.bladder.histo" ~ "Gall Bladder Histological Abnormalities",
    lvl2 == "gen.metabolism"~ "General Metabolism",
    lvl2 == "gill.histo"~ "Gill Histological Abnormalities",
    lvl2 == "gonad.histo"~"Gonad Histological Abnormalities",
    lvl2 == "growth"~ "Growth",
    lvl2 == "immune.cells"~"Immune Cells",
    lvl2 == "immune.other"~"Immune Other ",
    lvl2 == "intestinal.permeability"~"Intestinal Permeability",
    lvl2 == "kidney.histo"~"Kidney Histological abnormalities",
    lvl2 == "lipid.metabolism"~"Lipid Metabolism",
    lvl2 == "liver.histo"~"Liver Histological Abnormalities",
    lvl2 == "liver.kidney.products" ~ "Liver and Kidney Products",
    lvl2 == "locomotion"~"Locomotion",
    lvl2 == "mortality"~"Mortality",
    lvl2 == "nervous.system"~"Nervous System",
    lvl2 == "oxidative.stress"~"Oxidative Stress",
    lvl2 == "photosynthesis"~ "Photosynthesis",
    lvl2 == "proteobacteria"~"Protebacteria",
    lvl2 == "reproduction"~"Reproduction",
    lvl2 == "respiration"~"Respiration",
    lvl2 == "sexhormones"~"Sex Hormones",
    lvl2 == "shoaling"~"Shoaling",
    lvl2 == "stress"~"Stress",
    lvl2 == "vision.system"~"Vision System"))) %>% #Renames for widget
  mutate(bio_f = factor(case_when(bio.org == "cell"~"Cell", #Bio Org Data Tidying
    bio.org == "organism"~"Organism",
    bio.org == "population"~ "Population",
    bio.org == "subcell"~"Subcell",
    bio.org == "tissue" ~ "Tissue")))%>%
  mutate(vivo_f = factor(case_when(invitro.invivo == "invivo"~"In Vivo",
    invitro.invivo == "invitro"~"In Vitro")))%>% ##Renames for widget (Not using a widget right now, but saving for human health database)
  mutate(life_f = factor(case_when(life.stage == "Early"~"Early",
    life.stage == "Juvenile"~"Juvenile",
    life.stage == "Adult"~"Adult",
    life.stage == "Not Reported"~"Not Reported")))%>% #Renames for widget
  mutate(env_f = factor(case_when(environment == "Freshwater"~"Freshwater",
    environment == "Marine" ~ "Marine",
    environment == "Terrestrial" ~ "Terrestrial"))) %>%
  mutate(dose.mg.L.master.converted.reported = factor(dose.mg.L.master.converted.reported)) %>%
  mutate(dose.particles.mL.master.converted.reported = factor(dose.particles.mL.master.converted.reported))

#### SSD AO Setup ####

# Master dataset for SSDs
aoc_z <- aoc_setup %>% # start with Heili's altered dataset (no filtration for terrestrial data)
  # environment category data tidying.
  mutate(environment.noNA = replace_na(environment, "Not Reported")) %>% # replaces NA to better relabel.
  mutate(env_f = factor(environment.noNA, levels = c("Marine", "Freshwater", "Terrestrial", "Not Reported")))

# final cleanup and factoring  
aoc_z$Species <- as.factor(paste(aoc_z$genus,aoc_z$species)) #must make value 'Species" (uppercase)
aoc_z$Group <- as.factor(aoc_z$organism.group) #must make value "Group"
aoc_z$Group <- fct_explicit_na(aoc_z$Group) #makes sure that species get counted even if they're missing a group
```
```{r}
# subset data to selected variables

multiVar <- aoc_z %>% dplyr::select(#doi, size.category, 
                                    size_f,
                                    size.length.um.used.for.conversions, 
                                    shape, 
                                    polymer, 
                                    particle.volume.um3, 
                                    density.mg.um.3, 
                                    organism.group,
                                    environment, 
                                    bio.org, #biological level of organization
                                    #af.time, #assessment factor based on exposure time
                                    treatments, #number of doses (no including control)
                                    effect, #yes no
                                    exposure.duration.d, 
                                    exposure.route, #Factor
                                    organism.group, #factor
                                    media.temp, #numeric
                                    lvl1_f, #endpoints
                                    lvl2_f, #endpoints
                                   # lvl3, 
                                     dose.mg.L.master, 
                                    sex, #factor
                                    media.ph, #numeric
                                    media.sal.ppt, #numeric
                                    dose.particles.mL.master,
                                   effect.metric, #NOEC LOEC
                                    functional.group, #factor
                                    charge, #positive or negatibe
                                    zetapotential.mV) %>%  # numeric   
  filter(!size_f == "Not Reported")  #take out not reported

#recode variables
multiVar <- multiVar %>% mutate(effect_10 = case_when(
    effect == "Y" ~ 1,
    effect == "N" ~ 0
  ))# %>% 
  #mutate_all(is.character, ~as.factor())

```

# Analysis
## Data Exploration
### Completeness vby Size
```{r completeness heatmap}
CompletenessSummary_size <- multiVar %>%
  group_by(size_f) %>% 
     summarise_all((name = ~sum(is.na(.))/length(.))) %>% 
  mutate(across(is.numeric,~round(., 2))) %>% 
  mutate(across(is.numeric, ~100 *(1 -.)))
CompletenessSummary_size
```

```{r}
require()
#convert to matrix and transpose
transposed_size <- as.data.frame(t(as.matrix(CompletenessSummary_size[2:25]))) %>% #1 is category, 2-6 are variables
  arrange('1nm < 100nm')
#reassign column names
colnames(transposed_size) <- c("1nm < 100nm", "100nm < 1µm", "1µm < 100µm", "100µm < 1mm", "1mm < 5mm")
#transposedTag
#format as matrix
MissingMatrix_size <- data.matrix(transposed_size)
#build heatmap
pheatmap(MissingMatrix_size,
                           main = "Data Completeness by Size Bin", #title
                           fontsize = 12,
                           cluster_rows = FALSE, cluster_cols = FALSE,#disable dendrograms
                           display_numbers = TRUE,
                           treeheight_row = 0, treeheight_col = 0, #keeps clustering after dropping dendrograms
                           col = rev(brewer.pal(n = 9, name = "PuBu"))) #blue color scheme with 9 colors)

```
### Completeness by Environment
```{r completeness heatmap}
CompletenessSummary_environment <- multiVar %>%
  filter(!environment == "NA") %>% 
  group_by(environment) %>% 
     summarise_all((name = ~sum(is.na(.))/length(.))) %>% 
  mutate(across(is.numeric,~round(., 2))) %>% 
  mutate(across(is.numeric, ~100 *(1 -.)))
CompletenessSummary_environment
```

```{r}
require()
#convert to matrix and transpose
transposed_environment <- as.data.frame(t(as.matrix(CompletenessSummary_environment[2:25])))
#reassign column names
colnames(transposed_environment) <- c("Freshwater", "Marine", "Terrestrial")
#transposedTag
#format as matrix
MissingMatrix_environment <- data.matrix(transposed_environment)
#build heatmap
pheatmap(MissingMatrix_environment,
                           main = "Data Completeness by environment", #title
                           fontsize = 12,
                           cluster_rows = FALSE, cluster_cols = FALSE,#disable dendrograms
                           display_numbers = TRUE,
                           treeheight_row = 0, treeheight_col = 0, #keeps clustering after dropping dendrograms
                           col = rev(brewer.pal(n = 9, name = "PuBu"))) #blue color scheme with 9 colors)

```

## Stepwise Regression

```{r}
## Estimate an OLS Regression
fitols <- lm(effect_10 ~ size.length.um.used.for.conversions + particle.volume.um3 + exposure.duration.d + media.temp + dose.particles.mL.master, 
             na.action = na.omit, 
             data = multiVar)
summary(fitols)
```

```{r}
require(reshape2)
multiVar %>% 
  dplyr::select(size.length.um.used.for.conversions, particle.volume.um3, exposure.duration.d ,media.temp, dose.particles.mL.master) %>% 
  melt() %>%  #convert wide to long
   mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
  ggplot(aes(x = value)) + 
  stat_density() + 
  facet_wrap(~variable, scales = "free")
```

<!-- # ```{r} -->
<!-- # # cdplot(as.factor(effect) ~ particle.volume.um3, data = multiVar, -->
<!-- #        main = "Conditional Density Plot of ", xlab = "Dose", ylab = "Effect") -->
<!-- # ``` -->

## PCA

## Random Forest
### Recursive Partitioning And Regression Trees

The rpart algorithm works by splitting the dataset recursively, which means that the subsets that arise from a split are further split until a predetermined termination criterion is reached.  At each step, the split is made based on the independent variable that results in the largest possible reduction in heterogeneity of the dependent (predicted) variable.

It is important to note that the algorithm works by making the best possible choice at each particular stage, without any consideration of whether those choices remain optimal in future stages. That is, the algorithm makes a locally optimal decision at each stage. It is thus quite possible that such a choice at one stage turns out to be sub-optimal in the overall scheme of things.  In other words,  the algorithm does not find a globally optimal tree.
```{r}
#trim data so effect is always known
multiVar_sub <- multiVar %>% 
  drop_na(effect_10)

# Split data into training and test sets
set.seed(42)
multiVar_sub[,"train"] <- ifelse(runif(nrow(multiVar_sub)) < 0.8, 1, 0)
# Separate trainig and test sets
trainSet <- multiVar_sub[multiVar_sub$train==1,]
testSet <- multiVar_sub[multiVar_sub$train==0,]
#get column index of train flag
trainColNum <- grep("train", names(trainSet))
# Remove train flag column from train and test sets
trainSet <- trainSet[, -trainColNum]
testSet <- testSet[, -trainColNum]
```
Make a classification tree.

```{r, echo = FALSE}
#get column index of predicted variable in dataset
typeColNum <- grep("effect",names(multiVar_sub))

#build tree
require(rpart)
set.seed(15097)
t1 <- rpart(effect ~ size.length.um.used.for.conversions + particle.volume.um3 + exposure.duration.d + media.temp + dose.particles.mL.master,
            method = "class", #classification because response is discrete
            control = rpart.control(minbucket = 20, cp=0.008), #requires that there be at least 19 cases (responded + nonrespondents) in the final grouping of variable values of the terminal node of the tre..
            data = trainSet)
print(t1, digits=4)
```

Plot an interpretable tree.
```{r, echo = FALSE}
require(rpart.plot)
cols <- ifelse(t1$frame$yval == 1, "gray50", "black")
prp(t1, main="Tree for Effect",
    extra=106, # display prob of survival and percent of obs
    nn=TRUE, # display node numbers
    fallen.leaves=TRUE, # put leaves on the bottom of page
    branch=.5, # change angle of branch lines
    faclen=0, # do not abbreviate factor levels
    trace=1, # print automatically calculated cex
    shadow.col="gray", # shadows under the leaves
    branch.lty=1, # draw branches using solid lines
    branch.type=5, # branch lines width = weight(frame$wt), no. of cases here
    split.cex=1.2, # make split text larger than node text
    split.prefix="is ", # put "is " before split text
    split.suffix="?", # put "?" after split text
    col=cols, border.col=cols, # cols[2] if survived
    split.box.col="lightgray", # lightgray split boxes (default is white)
    split.border.col="darkgray", # darkgray border on split boxes
    split.round=0.5) # round the split box corners a tad
```
Next we see how good the model is by seeing how it fares against the test data.
```{r}
t1_predict <- predict(t1, newdata = testSet[,-typeColNum],
                      type="class")
mean(t1_predict==testSet$effect)
# [1] 0.7094088
#confusion matrix
table(pred=t1_predict,true=testSet$effect)
```
```{r}
par(mfrow=c(1,2)) # two plots on one page
#plot approximate R-squared and relative error for different splits (2 plots). labels are only appropriate for the "anova" method.
rsq.rpart(t1)
```

Next, we prune the tree using the cost complexity criterion. Basically, the intent is to see if a shallower subtree can give us comparable results. If so, we’d be better of choosing the shallower tree because it reduces the likelihood of overfitting.

As described earlier, we choose the appropriate pruning parameter (aka cost-complexity parameter) \alpha by picking the value that results in the lowest prediction error. Note that all relevant computations have already been carried out by R when we built the original tree (the call to rpart in the code above). All that remains now is to pick the value of \alpha:

#### Pruning

```{r}
#cost-complexity pruning
printcp(t1)
```
It is clear from the above, that the lowest cross-validation error (xerror in the table) occurs for \alpha =0.008 (this is CP in the table above).   One can find CP programatically like so:
```{r}
# get index of CP with lowest xerror
opt <- which.min(t1$cptable[,"xerror"])
#get its value
cp <- t1$cptable[opt, "CP"]
```
Next, we prune the tree based on this value of CP:

```{r}
# prune the tree
pt1 <- prune(t1,cp)
#pt1<- prune(t1, cp= t1$cptable[which.min(t1$cptable[,"xerror"]),"CP"])

# plot the pruned tree
plot(pt1, uniform=TRUE,
   main="Pruned Classification Tree");text(pt1, use.n=TRUE, all=TRUE, cex=.8)

#post(pfit, file = "c:/ptree.ps",
 #  title = "Pruned Classification Tree for Kyphosis")
```
```{r}
#find proportion of correct predictions using test set
t1_pruned_predict <- predict(pt1, testSet, type="class")
mean(t1_pruned_predict == testSet$effect)
#
```
This is not an improvement over an unprunend tree.. We need to check that this holds up for different training and test sets. This is easily done by creating multiple random partitions of the dataset and checking the efficacy of pruning for each. To do this efficiently, I’ll create a function that takes the training fraction, number of runs (partitions) and the name of the dataset as inputs and outputs the proportion of correct predictions for each run. It also optionally prunes the tree.

```{r}
#function to do multiple runs
multiple_runs_classification <- function(train_fraction,n,dataset,prune_tree=FALSE){
fraction_correct <- rep(NA,n)
set.seed(42)
for (i in 1:n){
  dataset[,"train"] <- ifelse(runif(nrow(dataset))<0.8,1,0)
  trainColNum <- grep("train",names(dataset))
  typeColNum <- grep("effect",names(dataset))
  trainSet <- dataset[dataset$train==1,-trainColNum]
  testSet <- dataset[dataset$train==0,-trainColNum]
  rpart_model <- rpart(effect~ size.length.um.used.for.conversions + particle.volume.um3 + exposure.duration.d + media.temp + dose.particles.mL.master,data = trainSet, method="class")
if(prune_tree==FALSE) {
  rpart_test_predict <- predict(rpart_model,testSet[,-typeColNum],type="class")
  fraction_correct[i] <- mean(rpart_test_predict==testSet$effect)
  }else{
    opt <- which.min(rpart_model$cptable[,"xerror"])
    cp <- rpart_model$cptable[opt, "CP"]
    pruned_model <- prune(rpart_model,cp)
    rpart_pruned_predict <- predict(pruned_model,testSet[,-typeColNum],type="class")
    fraction_correct[i] <- mean(rpart_pruned_predict == testSet$effect)
  }
  }
return(fraction_correct)
}
```

Note that in the above,  I have set the default value of the prune_tree to FALSE, so the function will execute the first branch of the if statement unless the default is overridden.

OK, so let’s do 50 runs with and without pruning, and check the mean and variance of the results for both sets of runs.
```{r}
#50 runs, no pruning
unpruned_set <- multiple_runs_classification(0.8,50,multiVar_sub)
mean(unpruned_set)
#[1] 0.7261882
sd(unpruned_set)
#[1] 0.01554347
#50 runs, with pruning
pruned_set <- multiple_runs_classification(0.8,50,multiVar_sub,prune_tree=TRUE)
mean(pruned_set)
#[1] 0.7261875
sd(pruned_set)
#[1] 0.01552285
```

### CForest
```{r run model}
require(party)
crf <- cforest(as.factor(effect_10) ~ size.length.um.used.for.conversions + particle.volume.um3 + exposure.duration.d +
                 media.temp + dose.particles.mL.master,
               controls = cforest_control(ntree = 500,
                                          mincriterion = qnorm(0.8), 
                                          trace = TRUE), # adds project bar because it's very slow
               data = multiVar_sub)

crf
```

```{r fit results}
train2 <- multiVar_sub %>% dplyr::select(size.length.um.used.for.conversions,
                                  particle.volume.um3,exposure.duration.d,media.temp,dose.particles.mL.master)
train3 <-multiVar_sub %>% dplyr::select(size.length.um.used.for.conversions,
                                  particle.volume.um3,exposure.duration.d,media.temp,dose.particles.mL.master,
                                  effect_10)

fitted <- predict(crf, train2, OOB = TRUE, type ="response")
#rpart.prob <- predict(t1, newdata=imputedSmalls.requested.voluntary,type="prob")

misClasificError <- mean(fitted != train3$effect_10)
print(paste('Training Accuracy', 1 - misClasificError))
```

```{r}
print(crf)
```

```{r}
#Obtaining predicted probabilites for Test data
tree.probs=predict(crf,
                 newdata = train2,
                 type="prob")

#Calculate ROC curve
rocCurve.tree <- roc(train3$effect_10,tree.probs[,'1'])
#plot the ROC curve
plot(rocCurve.tree,col=c(4))
```

Alternative ROC Curve
```{r}
require(pROC)
predicted <- predict(crf, train2, OOB=TRUE, type= "response")
auc(as.numeric(train3$effect_10), as.numeric(predicted))
#Calculate ROC curve
rocCurve.tree <- roc(train3$effect_10,as.numeric(predicted))
#plot the ROC curve
plot(rocCurve.tree,col=c(4))
```
```{r}
#plot reciever operating curve
perf <- performance(preds,"tpr","fpr")
```

```{r}
# compute in-sample results
caret::confusionMatrix(fitted,as.factor(train3$effect_10))
```

```{r}
#plot feature importance
cforestImpPlot <- function(x) {
  cforest_importance <<- v <- varimp(x)
  dotchart(v[order(v)])
}

importancePlot <- cforestImpPlot(crf)
importancePlot
```

```{r}
crf.prob <- matrix(unlist(crfsrvy.prob), ncol=2, byrow=TRUE)
apply(crf.prob,2,mean)
#[1] 0.2524514 0.7475486
tab <- round(cbind(by(rpart.prob[,2], INDICES=t1$where, mean), by(crf.prob[,2], INDICES=t1$where, mean)), 4)
colnames(tab) <- c("rpart", "cforest")
kable(tab)
```