### 2021-02-04 multivariate.Rmd
more modeling

---
title: "Multivariate Analysis"
author: "Scott Coffin"
date: "1/19/2021"
output: html_document
---
# Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load packages
library(tidyverse) #General everything
library(RColorBrewer)
library(ggplot2) #General plotting
library(ggrepel) #For adding text labels that repel away from data points
library(calecopal) #Color palette
library(shiny) #Runs shiny
library(shinythemes) #Shiny theme for the page
library(shinyWidgets) #Widgets
library(scales) #SSD - Use the percent format
library(reshape2) #Overview tab - melts bars together
library(ssdtools) #SSD package
library(DT) #Build HTML data tables
#library(plotly) #Make plots interactive
library(viridis) #Colors
library(scales) #To use "percent" function
library(shinyjs) #Exploration tab - reset button
library(tigerstats) #turns things into percents
library(ggbeeswarm) #plot all points
library(fitdistrplus) #alt SSD 
library(RColorBrewer) #colors
library(pheatmap) #pretty heat maps
library(rpart)  #for trees
#library(rattle)    # Fancy tree plot This is a difficult library to install (https://gist.github.com/zhiyzuo/a489ffdcc5da87f28f8589a55aa206dd) 
library(rpart.plot)             # Enhanced tree plots
library(RColorBrewer)       # Color selection for fancy tree plot
library(party)                  # Alternative decision tree algorithm
# library(partykit)               # Convert rpart object to BinaryTree
library(pROC)   #for ROC curves
library(uwot) #umap
library(ISLR)  #for the Carseat Data
library(quantregForest)
library(caret)
library(tidyverse)
library(tidymodels)
library(skimr)
library(sf)
library(ggspatial)
library(nhdplusTools)
library(patchwork)
library(Metrics)
library(gt)
```
## Data import
```{r data import}
# Load finalized dataset.
aoc <- read_csv("AquaticOrganisms_Clean_final.csv", guess_max = 10000)

#assign row ID
aoc <- rowid_to_column(aoc, "ID")
aoc <- aoc %>% 
  mutate(ID = paste0(ID, "ID"))

#### Introduction Setup ####

# All text inputs below.

#### Overview AO Setup ####

#Final_effect_dataset <- read_csv("Final_effect_dataset.csv")%>%
  #mutate(plot_f = case_when(
    #plot_f == "Polymer" ~ "Polymer",
    #plot_f == "Size" ~ "Size",
    #plot_f == "Shape" ~ "Shape",
    #plot_f == "Organism" ~ "Organism",
    #plot_f == "Lvl1" ~ "Endpoint Category",
    #plot_f == "Life.stage" ~ "Life Stage",
    #plot_f == "Invivo.invivo" ~ "In Vivo or In Vitro",
    #plot_f == "Exposure.route" ~ "Exposure Route"))%>%
  #mutate(plot_f = factor(plot_f))%>%
  #mutate(logEndpoints = log(Endpoints))%>%
  #rename(Percent = Freq)

polydf<-rowPerc(xtabs( ~polymer +effect, aoc)) #pulls polymers by effect 
polyf<-as.data.frame(polydf)%>% #Makes data frame 
  filter(effect %in% c("Y","N"))%>% #Sorts into Yes and No
  mutate(polymer = case_when(
    polymer == "BIO" ~ "Biopolymer",
    polymer == "EVA" ~ "Polyethylene Vinyl Acetate",
    polymer == "LTX" ~ "Latex",
    polymer == "PA" ~ "Polyamide",
    polymer == "PE" ~ "Polyethylene",
    polymer == "PC" ~ "Polycarbonate",
    polymer == "PET" ~ "Polyethylene Terephthalate",
    polymer == "PI" ~ "Polyisoprene",
    polymer == "PMMA" ~ "Polymethylmethacrylate",
    polymer == "PP" ~ "Polypropylene",
    polymer == "PS" ~ "Polystyrene",
    polymer == "PUR" ~ "Polyurathane",
    polymer == "PVC" ~ "Polyvinylchloride",
    polymer == "PLA" ~ "Polylactic Acid"))%>%
  mutate_if(is.numeric, round,0) #rounds percents 
Endpoints<-xtabs(~polymer +effect ,aoc) #Pulls all study obs. for polymer from dataset
polyfinal<- data.frame(cbind(polyf, Endpoints))%>% #adds it as a column
  rename(Endpoints='Freq.1')%>% #renames column
  mutate(logEndpoints = log(Endpoints))%>%
  rename(Percent = Freq)#renames column

sizedf<-rowPerc(xtabs(~size.category +effect, aoc))
sizef<-as.data.frame(sizedf)%>%
  filter(effect %in% c("Y","N"))%>%
  mutate(size.category = case_when(
    size.category == 1 ~ "<1µm",
    size.category == 2 ~ "1µm < 10µm",
    size.category == 3 ~ "10µm < 100µm",
    size.category == 4 ~ "100µm < 1mm",
    size.category == 5 ~ "1mm < 5mm",
    size.category == 0 ~ "unavailable"))%>%
  rename(Type = "size.category")%>%
  mutate_if(is.numeric, round,0)%>%
  mutate(plot="Size")
study_s<-xtabs(~size.category +effect ,aoc)
sizefinal<- data.frame(cbind(sizef, study_s))%>% 
  rename(Endpoints='Freq.1')%>%
  rename(category='size.category')%>%
  mutate(logEndpoints = log(Endpoints))%>%
  rename(Percent = Freq)#renames column

shapedf<-rowPerc(xtabs(~shape + effect, aoc))
shapef<-as.data.frame(shapedf)%>%
  filter(effect %in% c("Y","N"))%>%
  rename(Type="shape")%>%
  mutate_if(is.numeric, round,0)%>%
  mutate(plot="Shape")%>%
  mutate(Type = case_when(
    Type == "cube" ~ "Cube",
    Type == "sphere" ~ "Sphere",
    Type == "fragment" ~ "Fragment",
    Type == "fiber" ~ "Fiber"))
study_sh<-xtabs(~shape + effect,aoc)
shapefinal<- data.frame(cbind(shapef, study_sh))%>% 
  rename(Endpoints='Freq.1')%>%
  rename(category='shape')%>%
  mutate(logEndpoints = log(Endpoints))%>%
  rename(Percent = Freq)#renames column

taxdf<-rowPerc(xtabs(~organism.group +effect, aoc))
taxf<-as.data.frame(taxdf)%>%
  filter(effect %in% c("Y","N"))%>%
  rename(Type= "organism.group")%>%
  mutate_if(is.numeric, round,0)%>%
  mutate(plot="Organism")
study_t<-xtabs(~organism.group +effect,aoc)
taxfinal<- data.frame(cbind(taxf, study_t))%>% 
  rename(Endpoints='Freq.1')%>%
  rename(category='organism.group')%>%
  mutate(logEndpoints = log(Endpoints))%>%
  rename(Percent = Freq)#renames column

lvl1df<-rowPerc(xtabs(~lvl1 +effect, aoc))
lvl1f<-as.data.frame(lvl1df)%>%
  filter(effect %in% c("Y","N"))%>%
  rename(Type= "lvl1")%>%
  mutate_if(is.numeric, round,0)%>%
  mutate(plot="Lvl1")%>%
  mutate(Type = case_when(
    Type == "alimentary.excretory" ~ "Alimentary, Excretory",
    Type == "behavioral.sense.neuro" ~ "Behavioral, Sensory, Neurological",
    Type == "circulatory.respiratory" ~ "Circulatory, Respiratory",
    Type == "community" ~ "Community",
    Type == "fitness" ~ "Fitness",
    Type == "immune" ~ "Immune",
    Type == "metabolism" ~ "Metabolism",
    Type == "microbiome" ~ "Microbiome",
    Type == "stress" ~ "Stress")) 
study_l<-xtabs(~lvl1 +effect,aoc)
lvl1final<- data.frame(cbind(lvl1f, study_l))%>% 
  rename(Endpoints='Freq.1')%>%
  rename(category='lvl1')%>%
  mutate(logEndpoints = log(Endpoints))%>%
  rename(Percent = Freq)#renames column

lifedf<-rowPerc(xtabs(~life.stage +effect, aoc))
lifef<-as.data.frame(lifedf)%>%
  filter(effect %in% c("Y","N"))%>%
  rename(Type= "life.stage")%>%
  mutate_if(is.numeric, round,0)%>%
  mutate(plot="Life.stage")
studyli<-xtabs(~life.stage +effect ,aoc)
lifefinal<- data.frame(cbind(lifef, studyli))%>% 
  rename(Endpoints='Freq.1')%>%
  rename(category='life.stage')%>%
  mutate(logEndpoints = log(Endpoints))%>%
  rename(Percent = Freq)#renames column

vivodf<-rowPerc(xtabs(~invitro.invivo +effect, aoc))
vivof<-as.data.frame(vivodf)%>%
  filter(effect %in% c("Y","N"))%>%
  rename(Type= "invitro.invivo")%>%
  mutate_if(is.numeric, round,0)%>%
  mutate(plot="Invivo.invivo")%>%
  mutate(Type = case_when(
    Type=="invivo"~"In Vivo",
    Type=="invitro"~"In Vitro"))
study_v<-xtabs(~invitro.invivo +effect,aoc)
vivofinal<- data.frame(cbind(vivof, study_v))%>% 
  rename(Endpoints='Freq.1')%>%
  rename(category='invitro.invivo')%>%
  mutate(logEndpoints = log(Endpoints))%>%
  rename(Percent = Freq)#renames column

routedf<-rowPerc(xtabs(~exposure.route +effect, aoc))
routef<-as.data.frame(routedf)%>%
  filter(effect %in% c("Y","N"))%>%
  rename(Type= "exposure.route")%>%
  mutate_if(is.numeric, round,0)%>%
  mutate(plot="Exposure.route")%>%
  mutate(Type = case_when(
    Type == "coparental.exposure" ~"Co-Parental Exposure",
    Type == "paternal.exposure" ~ "Paternal Exposure",
    Type == "maternal.exposure" ~ "Maternal Exposure",
    Type == "food" ~ "Food",
    Type == "water" ~ "Water",
    Type == "sediment" ~ "Sediment",
    Type == "media" ~ "Media"))
study_r<-xtabs(~exposure.route +effect,aoc)
routefinal<- data.frame(cbind(routef, study_r))%>% 
  rename(Endpoints='Freq.1')%>%
  rename(category='exposure.route')%>%
  mutate(logEndpoints = log(Endpoints))%>%
  rename(Percent = Freq)#renames column

#### Exploration AO Setup ####

# Master dataset for scatterplots - for Heili's tab.
aoc_v1 <- aoc %>% # start with original dataset
   # full dataset filters.
  mutate(effect_f = factor(case_when(effect == "Y" ~ "Yes",
    effect == "N" ~ "No"),
    levels = c("No", "Yes"))) %>%
  # removing NAs to make data set nicer
  replace_na(list(size.category = 0, shape = "Not Reported", polymer = "Not Reported", life.stage = "Not Reported"))

aoc_setup <- aoc_v1 %>% # start with original dataset
  mutate(size_f = factor(case_when(
    size.category == 1 ~ "1nm < 100nm",
    size.category == 2 ~ "100nm < 1µm",
    size.category == 3 ~ "1µm < 100µm",
    size.category == 4 ~ "100µm < 1mm",
    size.category == 5 ~ "1mm < 5mm",
    size.category == 0 ~ "Not Reported"),
    levels = c("1nm < 100nm", "100nm < 1µm", "1µm < 100µm", "100µm < 1mm", "1mm < 5mm", "Not Reported"))) %>% # creates new column with nicer names and order by size levels.
  # shape category data tidying.
  mutate(shape_f = factor(case_when(
    shape == "fiber" ~ "Fiber",
    shape == "fragment" ~ "Fragment",
    shape == "sphere" ~ "Sphere",
    shape == "Not Reported" ~ "Not Reported"),
    levels = c("Fiber", "Fragment", "Sphere", "Not Reported"))) %>% # order our different shapes.
  # polymer category data tidying.
  mutate(poly_f = factor(case_when(
    polymer == "BIO" ~ "Biopolymer",
    polymer == "EVA" ~ "Polyethylene Vinyl Acetate",
    polymer == "LTX" ~ "Latex",
    polymer == "PA" ~ "Polyamide",
    polymer == "PE" ~ "Polyethylene",
    polymer == "PC" ~ "Polycarbonate",
    polymer == "PET" ~ "Polyethylene Terephthalate",
    polymer == "PI" ~ "Polyisoprene",
    polymer == "PMMA" ~ "Polymethylmethacrylate",
    polymer == "PP" ~ "Polypropylene",
    polymer == "PS" ~ "Polystyrene",
    polymer == "PUR" ~ "Polyurathane",
    polymer == "PVC" ~ "Polyvinylchloride",
    polymer == "PLA" ~ "Polylactic Acid",
    polymer == "Not Reported" ~ "Not Reported"))) %>%
  # taxonomic category data tidying.
  mutate(org_f = factor(organism.group, levels = c("Algae", "Annelida", "Bacterium", "Cnidaria", "Crustacea",
                                                   "Echinoderm", "Fish", "Insect", "Mollusca", "Nematoda", "Plant", "Rotifera", "Mixed"))) %>% # order our different organisms.
  mutate(lvl1_f = factor(case_when(lvl1 == "alimentary.excretory" ~ "Alimentary, Excretory",
    lvl1 == "behavioral.sense.neuro" ~ "Behavioral, Sensory, Neurological",
    lvl1 == "circulatory.respiratory" ~ "Circulatory, Respiratory",
    lvl1 == "community" ~ "Community",
    lvl1 == "fitness" ~ "Fitness",
    lvl1 == "immune" ~ "Immune",
    lvl1 == "metabolism" ~ "Metabolism",
    lvl1 == "microbiome" ~ "Microbiome",
    lvl1 == "stress" ~ "Stress"))) %>% # creates new column with nicer names.
  # Level 2 Data tidying
  mutate(lvl2_f = factor(case_when(lvl2 == "abundance"~"Abundance",
    lvl2 == "actinobacteria" ~ "Actinobacteria",
    lvl2 == "aggressivity"~"Agressivity",
    lvl2 == "ammonia.excretion" ~ "Ammonia Excretion",
    lvl2 == "bacteroidetes"~ "Bacteriodetes",
    lvl2 == "blood"~"Blood",
    lvl2 == "body.condition"~"Body Condition",
    lvl2 == "boldness"~"Boldness",
    lvl2 == "brain.histo"~"Brain Histological Abnormalities",
    lvl2 == "burrowing"~"Burrowing",
    lvl2 == "carb.metabolism"~"Carb Metabolism",
    lvl2 == "chemokines.cytokines"~"Chemokines",
    lvl2 == "circulatory"~"Circulatory",
    lvl2 == "detoxification"~"Detoxification",
    lvl2 == "development"~"Development",
    lvl2 == "digestion"~"Digestion",
    lvl2 == "digestive.enzymes"~"Digestive Enzymes",
    lvl2 == "digestive.tract.histo"~"Digestive Tract Histological Abnormalities",
    lvl2 == "diversity"~ "Diversity",
    lvl2 == "feeding"~ "Feeding",
    lvl2 == "firmicutes"~ "Firmicutes",
    lvl2 == "gall.bladder.histo" ~ "Gall Bladder Histological Abnormalities",
    lvl2 == "gen.metabolism"~ "General Metabolism",
    lvl2 == "gill.histo"~ "Gill Histological Abnormalities",
    lvl2 == "gonad.histo"~"Gonad Histological Abnormalities",
    lvl2 == "growth"~ "Growth",
    lvl2 == "immune.cells"~"Immune Cells",
    lvl2 == "immune.other"~"Immune Other ",
    lvl2 == "intestinal.permeability"~"Intestinal Permeability",
    lvl2 == "kidney.histo"~"Kidney Histological abnormalities",
    lvl2 == "lipid.metabolism"~"Lipid Metabolism",
    lvl2 == "liver.histo"~"Liver Histological Abnormalities",
    lvl2 == "liver.kidney.products" ~ "Liver and Kidney Products",
    lvl2 == "locomotion"~"Locomotion",
    lvl2 == "mortality"~"Mortality",
    lvl2 == "nervous.system"~"Nervous System",
    lvl2 == "oxidative.stress"~"Oxidative Stress",
    lvl2 == "photosynthesis"~ "Photosynthesis",
    lvl2 == "proteobacteria"~"Protebacteria",
    lvl2 == "reproduction"~"Reproduction",
    lvl2 == "respiration"~"Respiration",
    lvl2 == "sexhormones"~"Sex Hormones",
    lvl2 == "shoaling"~"Shoaling",
    lvl2 == "stress"~"Stress",
    lvl2 == "vision.system"~"Vision System"))) %>% #Renames for widget
  mutate(bio_f = factor(case_when(bio.org == "cell"~"Cell", #Bio Org Data Tidying
    bio.org == "organism"~"Organism",
    bio.org == "population"~ "Population",
    bio.org == "subcell"~"Subcell",
    bio.org == "tissue" ~ "Tissue")))%>%
  mutate(vivo_f = factor(case_when(invitro.invivo == "invivo"~"In Vivo",
    invitro.invivo == "invitro"~"In Vitro")))%>% ##Renames for widget (Not using a widget right now, but saving for human health database)
  mutate(life_f = factor(case_when(life.stage == "Early"~"Early",
    life.stage == "Juvenile"~"Juvenile",
    life.stage == "Adult"~"Adult",
    life.stage == "Not Reported"~"Not Reported")))%>% #Renames for widget
  mutate(env_f = factor(case_when(environment == "Freshwater"~"Freshwater",
    environment == "Marine" ~ "Marine",
    environment == "Terrestrial" ~ "Terrestrial"))) %>%
  mutate(dose.mg.L.master.converted.reported = factor(dose.mg.L.master.converted.reported)) %>%
  mutate(dose.particles.mL.master.converted.reported = factor(dose.particles.mL.master.converted.reported)) %>% 
  mutate(af.time_noNA = replace_na(af.time, "Unavailable")) %>% 
  mutate(acute.chronic_f = factor(case_when(af.time_noNA == 10 ~ "Acute",
                                            af.time_noNA == 1 ~ "Chronic",
                                            af.time_noNA == "Unavailable" ~ "Unavailable"))) %>%    #factorize assesment factor time into chronic/acute
  mutate(dose.mg.L.master.AF.noec = dose.mg.L.master * af.noec) %>% #converts ECxx, LC50, HONEC, LOEC, NOEC all to NOEC
  mutate(dose.particles.mL.master.AF.noec = dose.particles.mL.master * af.noec)  #converts ECxx, LC50, HONEC, LOEC, NOEC all to NOEC

#### SSD AO Setup ####

# Master dataset for SSDs
aoc_z <- aoc_setup %>% # start with Heili's altered dataset (no filtration for terrestrial data)
  # environment category data tidying.
  mutate(environment.noNA = replace_na(environment, "Not Reported")) %>% # replaces NA to better relabel.
  mutate(env_f = factor(environment.noNA, levels = c("Marine", "Freshwater", "Terrestrial", "Not Reported")))

# final cleanup and factoring  
aoc_z$Species <- as.factor(paste(aoc_z$genus,aoc_z$species)) #must make value 'Species" (uppercase)
aoc_z$Group <- as.factor(aoc_z$organism.group) #must make value "Group"
aoc_z$Group <- fct_explicit_na(aoc_z$Group) #makes sure that species get counted even if they're missing a group
```
```{r}
# subset data to selected variables
multiVar <- aoc_z %>% dplyr::select(#doi, size.category, 
                                    ID,
                                    size_f,
                                    size.length.um.used.for.conversions, 
                                    shape, 
                                    polymer, 
                                    particle.volume.um3, 
                                    density.mg.um.3, 
                                    organism.group,
                                    environment, 
                                    bio.org, #biological level of organization
                                    #af.time, #assessment factor based on exposure time
                                    treatments, #number of doses (no including control)
                                    effect_f, #yes no
                                    exposure.duration.d, 
                                    exposure.route, #Factor
                                    organism.group, #factor
                                    media.temp, #numeric
                                    lvl1_f, #endpoints
                                    lvl2_f, #endpoints
                                   # lvl3, 
                                     dose.mg.L.master, 
                                    sex, #factor
                                    media.ph, #numeric
                                    media.sal.ppt, #numeric
                                    dose.particles.mL.master,
                                   effect.metric, #NOEC LOEC
                                    functional.group, #factor
                                    charge, #positive or negatibe
                                    zetapotential.mV, # numeric   
                                   max.size.ingest.mm,#max ingestible size
                                   acute.chronic_f,
                                   dose.mg.L.master.AF.noec,
                                   dose.particles.mL.master.AF.noec,
                                   effect.score) %>%  #1 = minor, 2 = photosynthesis, feeding, 3 = growth, chlorophyll content, 4 = reproduction, 5  = population growth, 6 = survival
  filter(!size_f == "Not Reported") %>%   #take out not reported 
  mutate_if(is.character, as.factor) %>% mutate(effect_10 = case_when(
     effect_f == "Yes" ~ 1,
     effect_f == "No" ~ 0))
```

# Analysis
## Data Exploration
### Completeness vby Size
```{r completeness heatmap}
CompletenessSummary_size <- multiVar %>%
  group_by(size_f) %>% 
     summarise_all((name = ~sum(is.na(.))/length(.))) %>% 
  mutate(across(is.numeric,~round(., 2))) %>% 
  mutate(across(is.numeric, ~100 *(1 -.)))
CompletenessSummary_size
```

```{r}
require()
#convert to matrix and transpose
transposed_size <- as.data.frame(t(as.matrix(CompletenessSummary_size[2:28]))) %>% #1 is category, 2-6 are variables
  arrange('1nm < 100nm')
#reassign column names
colnames(transposed_size) <- c("1nm < 100nm", "100nm < 1µm", "1µm < 100µm", "100µm < 1mm", "1mm < 5mm")
#transposedTag
#format as matrix
MissingMatrix_size <- data.matrix(transposed_size)
#build heatmap
pheatmap(MissingMatrix_size,
                           main = "Data Completeness by Size Bin", #title
                           fontsize = 12,
                           cluster_rows = FALSE, cluster_cols = FALSE,#disable dendrograms
                           display_numbers = TRUE,
                           treeheight_row = 0, treeheight_col = 0, #keeps clustering after dropping dendrograms
                           col = rev(brewer.pal(n = 9, name = "PuBu"))) #blue color scheme with 9 colors)

```
### Completeness by Environment
```{r completeness heatmap}
CompletenessSummary_environment <- multiVar %>%
  filter(!environment == "NA") %>% 
  group_by(environment) %>% 
     summarise_all((name = ~sum(is.na(.))/length(.))) %>% 
  mutate(across(is.numeric,~round(., 2))) %>% 
  mutate(across(is.numeric, ~100 *(1 -.)))
CompletenessSummary_environment
```

```{r}
require()
#convert to matrix and transpose
transposed_environment <- as.data.frame(t(as.matrix(CompletenessSummary_environment[2:28])))
#reassign column names
colnames(transposed_environment) <- c("Freshwater", "Marine", "Terrestrial")
#transposedTag
#format as matrix
MissingMatrix_environment <- data.matrix(transposed_environment)
#build heatmap
pheatmap(MissingMatrix_environment,
                           main = "Data Completeness by environment", #title
                           fontsize = 12,
                           cluster_rows = FALSE, cluster_cols = FALSE,#disable dendrograms
                           display_numbers = TRUE,
                           treeheight_row = 0, treeheight_col = 0, #keeps clustering after dropping dendrograms
                           col = rev(brewer.pal(n = 9, name = "PuBu"))) #blue color scheme with 9 colors)

```

## Stepwise Regression

```{r}
## Estimate an OLS Regression
fitols <- lm(effect_10 ~ size.length.um.used.for.conversions + particle.volume.um3 + exposure.duration.d + media.temp + dose.particles.mL.master +mouth, 
             na.action = na.omit, 
             data = multiVar)
summary(fitols)
```

```{r}
require(reshape2)
multiVar %>% 
  dplyr::select(size.length.um.used.for.conversions, particle.volume.um3, exposure.duration.d ,media.temp, dose.particles.mL.master) %>% 
  melt() %>%  #convert wide to long
   mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
  ggplot(aes(x = value)) + 
  stat_density() + 
  facet_wrap(~variable, scales = "free")
```

<!-- # ```{r} -->
<!-- # # cdplot(as.factor(effect) ~ particle.volume.um3, data = multiVar, -->
<!-- #        main = "Conditional Density Plot of ", xlab = "Dose", ylab = "Effect") -->
<!-- # ``` -->

## PCA

# UMAP

```{r}
require(uwot)
require(Rtsne)
require(vizier) #devtools::install_github("jlmelville/vizier")

# For some functions we need to strip out non-numeric columns and convert data to matrix
x2m <- function(X) {
  if (!methods::is(X, "matrix")) {
    m <- as.matrix(X[, which(vapply(X, is.numeric, logical(1)))])
  }
  else {m <- X} 
  m}

#choose values with most completeness
multiVar2 <- multiVar %>% 
  filter(!environment == "Terrestrial") %>% 
  dplyr::select(size_f, size.length.um.used.for.conversions, shape, polymer, particle.volume.um3, density.mg.um.3, organism.group, bio.org, treatments, effect, exposure.duration.d, exposure.route, lvl1_f, dose.mg.L.master) %>% 
  mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
  drop_na() #drop missing

#convert discrete variables to numeric
multiVar2[] <- data.matrix(multiVar2)

# build umap for small dataset (<10,000 points)
multiVar_map  <- umap(multiVar2, pca = 10)

# Remove duplicates for t-SNE
#multiVar2_noNa_dup <- multiVar2_noNa[-which(duplicated(x2m(multiVar2_noNa))), ]

#build t-SNE
#multiVar_tsne <- Rtsne::Rtsne(multiVar2_noNa_dup, perplexity = 15, initial_dims = 100, partical_pca = TRUE, exaggeration_factor = 4)

# Non-numeric columns are ignored, so in a lot of cases you can pass a data
# frame directly to umap
#iris_umap <- umap(iris, n_neighbors = 50, learning_rate = 0.5, init = "random")

#visualize umap
embed_img <- function(X, Y, k = 15, ...) {
  args <- list(...)
  args$coords <- Y
  args$x <- X

  do.call(vizier::embed_plot, args)
}

#plot
embed_img(multiVar2, multiVar_map, pc_axes = TRUE, equal_axes = TRUE, alpha_scale = 0.5, title = "Tox UMAP", cex = 1)
```

```{r}
#PCA
pca <- stats::prcomp(multiVar2[,-4], retx = TRUE, rank. = 2)
#build color pallete
my_colors = colorRampPalette(c("red", "yellow", "green"))(nrow(multiVar2))
#plot
embed_plot(pca$x, multiVar2$polymer, color_scheme = palette.colors(palette = "Okabe-Ito"), #turbo, #rainbow, #my_colors, 
           title = "Polymer PCA", alpha_scale = 0.5, equal_axes = TRUE)
```

```{r}
require(plotly)
#PCA for discrete variables
multiVar_discrete <- multiVar %>% 
  select(size.length.um.used.for.conversions, polymer, dose.mg.L.master, exposure.duration.d, effect) %>% 
  drop_na %>% 
  mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
  mutate(effect_10 = case_when(
    effect == "Y" ~ 1,
    effect == "N" ~ 0
  )) %>% 
  select(-effect)

#buildPCA
pca_discrete <- stats::prcomp(multiVar_discrete[,-2], retx = TRUE, rank. = 5, scale. = TRUE)

embed_plot(pca_discrete$x, multiVar_discrete$polymer)

embed_plotly(pca_discrete$x, multiVar_discrete$polymer, color_scheme = palette.colors(palette = "Okabe-Ito"), 
           title = "Polymer PCA", alpha_scale = 0.5, equal_axes = TRUE,
           tooltip = paste("Polymer:", multiVar_discrete$polymer))

#embed_plotly also option
```

# Random Forest
## Recursive Partitioning And Regression Trees

The rpart algorithm works by splitting the dataset recursively, which means that the subsets that arise from a split are further split until a predetermined termination criterion is reached.  At each step, the split is made based on the independent variable that results in the largest possible reduction in heterogeneity of the dependent (predicted) variable.

It is important to note that the algorithm works by making the best possible choice at each particular stage, without any consideration of whether those choices remain optimal in future stages. That is, the algorithm makes a locally optimal decision at each stage. It is thus quite possible that such a choice at one stage turns out to be sub-optimal in the overall scheme of things.  In other words,  the algorithm does not find a globally optimal tree.

The size of the data set usually imposes a hard limit on how many features may be considered due to Curse of Dimensionality, i.e. your data becomes sparser and sparser as you increase the number of features considered, which usually leads to overfitting. While there is no rule of thumb relating to how many features vs.  the number of observations you should use, I try to keep $e^Nf < No$ (Nf = number of features, No = number of observations) to minimise overfitting [this is not always possible and it does not ensure that we won’t overfit]. In this case, our training set has 3200, which suggests that using 8 features (e^8 ~ 2980) is not entirely absurd. Obviously, this depends on your data, so we will cover some further overfitting checks later on.
```{r}
#trim data so effect is always known
multiVar_sub <- multiVar %>% 
  drop_na(effect_10)

# Split data into training and test sets
set.seed(42)
multiVar_sub[,"train"] <- ifelse(runif(nrow(multiVar_sub)) < 0.8, 1, 0)
# Separate trainig and test sets
trainSet <- multiVar_sub[multiVar_sub$train==1,]
testSet <- multiVar_sub[multiVar_sub$train==0,]
#get column index of train flag
trainColNum <- grep("train", names(trainSet))
# Remove train flag column from train and test sets
trainSet <- trainSet[, -trainColNum]
testSet <- testSet[, -trainColNum]
```
Make a classification tree.

```{r, echo = FALSE}
#get column index of predicted variable in dataset
typeColNum <- grep("effect",names(multiVar_sub))

#build tree
require(rpart)
set.seed(15097)
t1 <- rpart(effect ~ size.length.um.used.for.conversions + particle.volume.um3 + exposure.duration.d + media.temp + dose.particles.mL.master,
            method = "class", #classification because response is discrete
            control = rpart.control(minbucket = 20, cp=0.008), #requires that there be at least 19 cases (responded + nonrespondents) in the final grouping of variable values of the terminal node of the tre..
            data = trainSet)
print(t1, digits=4)
```

Plot an interpretable tree.
```{r, echo = FALSE}
require(rpart.plot)
cols <- ifelse(t1$frame$yval == 1, "gray50", "black")
prp(t1, main="Tree for Effect",
    extra=106, # display prob of survival and percent of obs
    nn=TRUE, # display node numbers
    fallen.leaves=TRUE, # put leaves on the bottom of page
    branch=.5, # change angle of branch lines
    faclen=0, # do not abbreviate factor levels
    trace=1, # print automatically calculated cex
    shadow.col="gray", # shadows under the leaves
    branch.lty=1, # draw branches using solid lines
    branch.type=5, # branch lines width = weight(frame$wt), no. of cases here
    split.cex=1.2, # make split text larger than node text
    split.prefix="is ", # put "is " before split text
    split.suffix="?", # put "?" after split text
    col=cols, border.col=cols, # cols[2] if survived
    split.box.col="lightgray", # lightgray split boxes (default is white)
    split.border.col="darkgray", # darkgray border on split boxes
    split.round=0.5) # round the split box corners a tad
```
Next we see how good the model is by seeing how it fares against the test data.
```{r}
t1_predict <- predict(t1, newdata = testSet[,-typeColNum],
                      type="class")
mean(t1_predict==testSet$effect)
# [1] 0.7094088
#confusion matrix
table(pred=t1_predict,true=testSet$effect)
```
```{r}
par(mfrow=c(1,2)) # two plots on one page
#plot approximate R-squared and relative error for different splits (2 plots). labels are only appropriate for the "anova" method.
rsq.rpart(t1)
```

Next, we prune the tree using the cost complexity criterion. Basically, the intent is to see if a shallower subtree can give us comparable results. If so, we’d be better of choosing the shallower tree because it reduces the likelihood of overfitting.

As described earlier, we choose the appropriate pruning parameter (aka cost-complexity parameter) \alpha by picking the value that results in the lowest prediction error. Note that all relevant computations have already been carried out by R when we built the original tree (the call to rpart in the code above). All that remains now is to pick the value of \alpha:

### Pruning

```{r}
#cost-complexity pruning
printcp(t1)
```
It is clear from the above, that the lowest cross-validation error (xerror in the table) occurs for \alpha =0.008 (this is CP in the table above).   One can find CP programatically like so:
```{r}
# get index of CP with lowest xerror
opt <- which.min(t1$cptable[,"xerror"])
#get its value
cp <- t1$cptable[opt, "CP"]
```
Next, we prune the tree based on this value of CP:

```{r}
# prune the tree
pt1 <- prune(t1,cp)
#pt1<- prune(t1, cp= t1$cptable[which.min(t1$cptable[,"xerror"]),"CP"])

# plot the pruned tree
plot(pt1, uniform=TRUE,
   main="Pruned Classification Tree");text(pt1, use.n=TRUE, all=TRUE, cex=.8)

#post(pfit, file = "c:/ptree.ps",
 #  title = "Pruned Classification Tree for Kyphosis")
```
```{r}
#find proportion of correct predictions using test set
t1_pruned_predict <- predict(pt1, testSet, type="class")
mean(t1_pruned_predict == testSet$effect)
#
```
This is not an improvement over an unprunend tree.. We need to check that this holds up for different training and test sets. This is easily done by creating multiple random partitions of the dataset and checking the efficacy of pruning for each. To do this efficiently, I’ll create a function that takes the training fraction, number of runs (partitions) and the name of the dataset as inputs and outputs the proportion of correct predictions for each run. It also optionally prunes the tree.

```{r}
#function to do multiple runs
multiple_runs_classification <- function(train_fraction,n,dataset,prune_tree=FALSE){
fraction_correct <- rep(NA,n)
set.seed(42)
for (i in 1:n){
  dataset[,"train"] <- ifelse(runif(nrow(dataset))<0.8,1,0)
  trainColNum <- grep("train",names(dataset))
  typeColNum <- grep("effect",names(dataset))
  trainSet <- dataset[dataset$train==1,-trainColNum]
  testSet <- dataset[dataset$train==0,-trainColNum]
  rpart_model <- rpart(effect~ size.length.um.used.for.conversions + particle.volume.um3 + exposure.duration.d + media.temp + dose.particles.mL.master,data = trainSet, method="class")
if(prune_tree==FALSE) {
  rpart_test_predict <- predict(rpart_model,testSet[,-typeColNum],type="class")
  fraction_correct[i] <- mean(rpart_test_predict==testSet$effect)
  }else{
    opt <- which.min(rpart_model$cptable[,"xerror"])
    cp <- rpart_model$cptable[opt, "CP"]
    pruned_model <- prune(rpart_model,cp)
    rpart_pruned_predict <- predict(pruned_model,testSet[,-typeColNum],type="class")
    fraction_correct[i] <- mean(rpart_pruned_predict == testSet$effect)
  }
  }
return(fraction_correct)
}
```

Note that in the above,  I have set the default value of the prune_tree to FALSE, so the function will execute the first branch of the if statement unless the default is overridden.

OK, so let’s do 50 runs with and without pruning, and check the mean and variance of the results for both sets of runs.
```{r}
#50 runs, no pruning
unpruned_set <- multiple_runs_classification(0.8,50,multiVar_sub)
mean(unpruned_set)
#[1] 0.7261882
sd(unpruned_set)
#[1] 0.01554347
#50 runs, with pruning
pruned_set <- multiple_runs_classification(0.8,50,multiVar_sub,prune_tree=TRUE)
mean(pruned_set)
#[1] 0.7261875
sd(pruned_set)
#[1] 0.01552285
```

## CForest
```{r run model}
require(party)
crf <- cforest(as.factor(effect_10) ~ size.length.um.used.for.conversions + particle.volume.um3 + exposure.duration.d +
                 media.temp + dose.particles.mL.master,
               controls = cforest_control(ntree = 500,
                                          mincriterion = qnorm(0.8), 
                                          trace = TRUE), # adds project bar because it's very slow
               data = multiVar_sub)

crf
```

```{r fit results}
train2 <- multiVar_sub %>% dplyr::select(size.length.um.used.for.conversions,
                                  particle.volume.um3,exposure.duration.d,media.temp,dose.particles.mL.master)
train3 <-multiVar_sub %>% dplyr::select(size.length.um.used.for.conversions,
                                  particle.volume.um3,exposure.duration.d,media.temp,dose.particles.mL.master,
                                  effect_10)

fitted <- predict(crf, train2, OOB = TRUE, type ="response")
#rpart.prob <- predict(t1, newdata=imputedSmalls.requested.voluntary,type="prob")

misClasificError <- mean(fitted != train3$effect_10)
print(paste('Training Accuracy', 1 - misClasificError))
```

```{r}
print(crf)
```

```{r}
#Obtaining predicted probabilites for Test data
tree.probs=predict(crf,
                 newdata = train2,
                 type="prob")

#Calculate ROC curve
rocCurve.tree <- roc(train3$effect_10,tree.probs[,'1'])
#plot the ROC curve
plot(rocCurve.tree,col=c(4))
```

Alternative ROC Curve
```{r}
require(pROC)
predicted <- predict(crf, train2, OOB=TRUE, type= "response")
auc(as.numeric(train3$effect_10), as.numeric(predicted))
#Calculate ROC curve
rocCurve.tree <- roc(train3$effect_10,as.numeric(predicted))
#plot the ROC curve
plot(rocCurve.tree,col=c(4))
```
```{r}
#plot reciever operating curve
perf <- performance(preds,"tpr","fpr")
```

```{r}
# compute in-sample results
caret::confusionMatrix(fitted,as.factor(train3$effect_10))
```

```{r}
#plot feature importance
cforestImpPlot <- function(x) {
  cforest_importance <<- v <- varimp(x)
  dotchart(v[order(v)])
}

importancePlot <- cforestImpPlot(crf)
importancePlot
```

```{r}
crf.prob <- matrix(unlist(crfsrvy.prob), ncol=2, byrow=TRUE)
apply(crf.prob,2,mean)
#[1] 0.2524514 0.7475486
tab <- round(cbind(by(rpart.prob[,2], INDICES=t1$where, mean), by(crf.prob[,2], INDICES=t1$where, mean)), 4)
colnames(tab) <- c("rpart", "cforest")
kable(tab)
```
# Random Forest Package

## Preparation
```{r}
library(quantregForest)
library(caret)
library(tidyverse)
library(tidymodels)
library(skimr)
library(sf)
library(ggspatial)
library(nhdplusTools)
library(patchwork)
library(Metrics)
library(gt)
```

```{r}
multiVar <- multiVar %>% 
  mutate_if(is.character, as.factor) %>% 
  mutate(effect_10 = case_when( #convert ordinal to numeric
     effect_f == "Yes" ~ 1,
     effect_f == "No" ~ 0))

#choose relevant predictors and log-transform
multiVar_small_log <- multiVar %>% 
  dplyr::select(ID, size_f, size.length.um.used.for.conversions, shape, polymer, particle.volume.um3, density.mg.um.3, organism.group, bio.org, treatments, effect_f, exposure.duration.d, exposure.route, lvl1_f, dose.mg.L.master) %>% 
  mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
  drop_na() %>%  #drop missing
 mutate(effect_10 = case_when( #convert ordinal to numeric
     effect_f == "Yes" ~ 1,
     effect_f == "No" ~ 0
   ))# %>% 
 #  select(-effect_f)

#no log transform for comparison
multiVar_small <- multiVar %>% 
  dplyr::select(ID, size_f, size.length.um.used.for.conversions, shape, polymer, particle.volume.um3, density.mg.um.3, organism.group, bio.org, treatments, effect_f, #effect_10 , 
                exposure.duration.d, exposure.route, lvl1_f, dose.mg.L.master) %>% 
  drop_na() #%>%  #drop missing
# mutate(effect_10 = case_when( #convert ordinal to numeric
#     effect == "Y" ~ 1,
#     effect == "N" ~ 0
#   )) %>% 
  # select(-effect)

#ensure completeness
skim(multiVar_small)
```
## Training Data

```{r}
# Create calibration and validation splits with tidymodels initial_split() function.
set.seed(4)
multiVar_small_split <- multiVar_small %>%
  initial_split(prop = 0.75, strata = polymer) # splits data into training and testing set.
# default is 3/4ths split (but 75% training, 25% testing).
# Stratification (strata) = grouping training/testing sets by region, state, etc.
# Using the "strata" call ensures the number of data points in the training data is equivalent to the proportions in the original data set. (Strata below 10% of the total are pooled together.)

# Create a training data set with the training() function
# Pulls from training and testing sets created by initial_split()
multiVar_small_train <- training(multiVar_small_split)
multiVar_small_test <- testing(multiVar_small_split)
# Examine the environment to be sure # of observations looks like the 75/25 split. 3199:1066.

# Create a separate dataset of available IDs that were not used in the training dataset.
notTrain <- aoc %>% # all COMIDS from StreamCat data, sampled or not
  filter(!ID %in% aoc$ID) # Removing sites used to train the model. n = 140,097
```
## Kitchen Sink Model

```{r}
# Step Three - Kitchen Sink model -----------------------------------------

# Create finalized training dataset and include all possible variables. 
rf_dat <- multiVar_small_train %>%
  select(-ID)

# Random forest -- 
# a decision tree model, using predictors to answer dichotomous questions to create nested splits.
# no pruning happens - rather, multiple trees are built (the forest) and then you are looking for consensus across trees
# training data goes down the tree and ends up in a terminal node.
# if testing data goes down the same route, then this upholds our conclusions. Or, if it goes awry, this allows us to look for patterns in how it goes awry.

set.seed(2) # assures the data pulled is random, but sets it for the run below (makes outcome stable)
myrf <- randomForest(y = rf_dat$effect_f, # dependent variable
  x = rf_dat %>%
    select(-effect_f), # selecting all predictor variables
  importance = T, # how useful is a predictor in predicting values (nothing causal)
  proximity = T, 
  ntrees = 100) # 500 trees default.

myrf # examine the results.
```
~20% error rate. Let's compare this with the same model with log-transformed values.
```{r}
# Create calibration and validation splits with tidymodels initial_split() function.
set.seed(4)
multiVar_small_log_split <- multiVar_small_log %>%
  initial_split(prop = 0.75, strata = polymer) # splits data into training and testing set.

# Create a training data set with the training() function
# Pulls from training and testing sets created by initial_split()
multiVar_small_log_train <- training(multiVar_small_log_split)
multiVar_small_log_test <- testing(multiVar_small_log_split)

# Create finalized training dataset and include all possible variables. 
rf_dat_log <- multiVar_small_log_train %>%
  select(-ID, -effect_10)

set.seed(2) # assures the data pulled is random, but sets it for the run below (makes outcome stable)
myrf_log <- randomForest(y = rf_dat_log$effect_f, # dependent variable
  x = rf_dat_log %>%
    select(-effect_f), # selecting all predictor variables
  importance = T, # how useful is a predictor in predicting values (nothing causal)
  proximity = T, 
  ntrees = 100) # 500 trees default.

myrf_log # examine the results.
```
No performance enhancement with log-transformed values.

## Continuous Predictors
Repeat with continous variables whenever possible, and max of 8 predictors.
```{r}
skim(multiVar)
```
```{r}
acute <- multiVar %>% 
  filter(acute.chronic_f == "Acute") %>% 
  filter(!environment == "Terrestrial") %>% 
  filter(bio.org == "organism") %>% 
  filter(lvl1_f == "Fitness") %>% 
  filter(!exposure.route == "food") %>% 
  dplyr::select(-c(sex, functional.group, charge, zetapotential.mV, max.size.ingest.mm, media.sal.ppt, media.ph, media.temp, treatments, acute.chronic_f, max.size.ingest.mm, effect.metric, lvl2_f, exposure.route, exposure.duration.d, bio.org, environment, lvl1_f, dose.particles.mL.master.AF.noec, dose.mg.L.master.AF.noec, size_f, density.mg.um.3, dose.particles.mL.master, effect_10, ID)) %>%
  mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
  drop_na()
skim(acute)
```
With a complete dataset for just acute studies in aquatic organisms with aqueous route of exposure, we are left with 453 complete cases with 6 predictor variables, 1 response variable (effect y/n), and an ID. Let's determine if we have enough data for ML. 
```{r}
exp(1)^6
```
$e^6$ is less than n (453), so we may proceed.

```{r}
# Create calibration and validation splits with tidymodels initial_split() function.
set.seed(4)
acute_split <- acute %>%
  initial_split(prop = 0.75, strata = polymer) # splits data into training and testing set.
# default is 3/4ths split (but 75% training, 25% testing).
# Stratification (strata) = grouping training/testing sets by region, state, etc.
# Using the "strata" call ensures the number of data points in the training data is equivalent to the proportions in the original data set. (Strata below 10% of the total are pooled together.)

# Create a training data set with the training() function
# Pulls from training and testing sets created by initial_split()
acute_train <- training(acute_split)
acute_test <- testing(acute_split)
# Examine the environment to be sure # of observations looks like the 75/25 split. 3199:1066.

# Random forest -- 
set.seed(2) # assures the data pulled is random, but sets it for the run below (makes outcome stable)
acuterf <- randomForest(y = acute_train$effect_f, # dependent variable
  x = acute_train %>%
    dplyr::select(-effect_f), # selecting all predictor variables
  importance = T, # how useful is a predictor in predicting values (nothing causal)
  proximity = T, 
  ntrees = 100) # 500 trees default.

acuterf # examine the results.
```

```{r}
plot(acuterf)
# model performance appears to improve most at ~75 trees
```

```{r}
varImpPlot(acuterf)
# displays which variables are most important
# helps to winnow down list of predictors
# recommended to weigh left pane more
# right pane also shows how evenly things split based on the list of predictors
# values close to 0 can be dropped, but don't have to be
```
Dose, organism group, exposure duration are important.

```{r}
importance <- acuterf$importance
View(importance)
# displays the data plotted in the plot above
```

###Validation

```{r fit results}
fitted <- predict(acuterf, acute_test[,-6], OOB = TRUE, type ="response")
misClasificError <- mean(fitted != acute_test$effect_f)
print(paste('Training Accuracy', 1 - misClasificError))
```

Alternative ROC Curve
```{r}
require(pROC)
predicted <- predict(acuterf, acute_test %>%  dplyr::select(-effect_f),
                       OOB=TRUE, type= "response")
#Calculate ROC curve
rocCurve.tree <- roc(as.numeric(acute_test$effect_f),as.numeric(predicted))

##gplot
# rocks <- roc()

#plot the ROC curve
plot(rocCurve.tree,col=c(4))
```
##### Chronic
```{r}
chronic <- multiVar %>% 
  filter(acute.chronic_f == "Chronic") %>% 
  filter(!environment == "Terrestrial") %>% 
  filter(bio.org == "organism") %>% 
  filter(lvl1_f == "Fitness") %>% 
  filter(!exposure.route == "food") %>% 
  dplyr::select(-c(sex, functional.group, charge, zetapotential.mV, max.size.ingest.mm, media.sal.ppt, media.ph, media.temp, treatments, acute.chronic_f, max.size.ingest.mm, effect.metric, lvl2_f, exposure.route, exposure.duration.d, bio.org, environment, lvl1_f, dose.particles.mL.master.AF.noec, dose.mg.L.master.AF.noec, size_f, density.mg.um.3, dose.particles.mL.master, effect_10, ID)) %>%
  mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
  drop_na()
skim(chronic)
```
With a complete dataset for just acute studies in aquatic organisms with aqueous route of exposure, we are left with 453 complete cases with 6 predictor variables, 1 response variable (effect y/n), and an ID. Let's determine if we have enough data for ML. 
```{r}
exp(1)^6
```
$e^6$ is less than n (453), so we may proceed.

```{r}
# Create calibration and validation splits with tidymodels initial_split() function.
set.seed(4)
chronic_split <- chronic %>%
  initial_split(prop = 0.75, strata = polymer) # splits data into training and testing set.
# default is 3/4ths split (but 75% training, 25% testing).
# Stratification (strata) = grouping training/testing sets by region, state, etc.
# Using the "strata" call ensures the number of data points in the training data is equivalent to the proportions in the original data set. (Strata below 10% of the total are pooled together.)

# Create a training data set with the training() function
# Pulls from training and testing sets created by initial_split()
chronic_train <- training(chronic_split)
chronic_test <- testing(chronic_split)
# Examine the environment to be sure # of observations looks like the 75/25 split. 3199:1066.

# Random forest -- 
set.seed(2) # assures the data pulled is random, but sets it for the run below (makes outcome stable)
chronicrf <- randomForest(y = chronic_train$effect_f, # dependent variable
  x = chronic_train %>%
    dplyr::select(-effect_f), # selecting all predictor variables
  importance = T, # how useful is a predictor in predicting values (nothing causal)
  proximity = T, 
  ntrees = 100) # 500 trees default.

chronicrf # examine the results.
```

```{r}
plot(chronicrf)
# model performance appears to improve most at ~75 trees
```

```{r}
varImpPlot(chronicrf)
# displays which variables are most important
# helps to winnow down list of predictors
# recommended to weigh left pane more
# right pane also shows how evenly things split based on the list of predictors
# values close to 0 can be dropped, but don't have to be
```
Dose, organism group, exposure duration are important.

```{r}
importance <- chronicrf$importance
View(importance)
# displays the data plotted in the plot above
```

```{r fit results}
fitted <- predict(chronicrf, chronic_test[,-6], OOB = TRUE, type ="response")
misClasificError <- mean(fitted != chronic_test$effect_f)
print(paste('Training Accuracy', 1 - misClasificError))
```

#### All Data
```{r}
all <- multiVar %>% 
  #filter(acute.chronic_f == "Chronic") %>% 
  filter(!environment == "Terrestrial") %>% 
  filter(bio.org == "organism") %>% 
  filter(lvl1_f == "Fitness") %>% 
  filter(!exposure.route == "food") %>% 
  dplyr::select(-c(sex, functional.group, charge, zetapotential.mV, max.size.ingest.mm, media.sal.ppt, media.ph, media.temp, treatments, acute.chronic_f, max.size.ingest.mm, effect.metric, lvl2_f, exposure.route, exposure.duration.d, bio.org, environment, lvl1_f, dose.particles.mL.master.AF.noec, dose.mg.L.master.AF.noec, size_f, density.mg.um.3, dose.particles.mL.master, effect_10, ID)) %>%
  mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
  drop_na()
skim(all)
```
With a complete dataset for just acute studies in aquatic organisms with aqueous route of exposure, we are left with 453 complete cases with 6 predictor variables, 1 response variable (effect y/n), and an ID. Let's determine if we have enough data for ML. 
```{r}
exp(1)^6
```
$e^6$ is less than n (453), so we may proceed.

```{r}
# Create calibration and validation splits with tidymodels initial_split() function.
set.seed(4)
all_split <- all %>%
  initial_split(prop = 0.75, strata = polymer) # splits data into training and testing set.
# default is 3/4ths split (but 75% training, 25% testing).
# Stratification (strata) = grouping training/testing sets by region, state, etc.
# Using the "strata" call ensures the number of data points in the training data is equivalent to the proportions in the original data set. (Strata below 10% of the total are pooled together.)

# Create a training data set with the training() function
# Pulls from training and testing sets created by initial_split()
all_train <- training(all_split)
all_test <- testing(all_split)
# Examine the environment to be sure # of observations looks like the 75/25 split. 3199:1066.

# Random forest -- 
set.seed(2) # assures the data pulled is random, but sets it for the run below (makes outcome stable)
allrf <- randomForest(y = all_train$effect_f, # dependent variable
  x = all_train %>%
    dplyr::select(-effect_f), # selecting all predictor variables
  importance = T, # how useful is a predictor in predicting values (nothing causal)
  proximity = T, 
  ntrees = 100) # 500 trees default.

allrf # examine the results.
```

```{r}
plot(allrf)
# model performance appears to improve most at ~75 trees
```

```{r}
varImpPlot(allrf)
# displays which variables are most important
# helps to winnow down list of predictors
# recommended to weigh left pane more
# right pane also shows how evenly things split based on the list of predictors
# values close to 0 can be dropped, but don't have to be
```
Dose, organism group, exposure duration are important.

```{r}
importance <- allrf$importance
View(importance)
# displays the data plotted in the plot above
```

```{r fit results}
fitted <- predict(allrf, all_test[,-6], OOB = TRUE, type ="response")
misClasificError <- mean(fitted != all_test$effect_f)
print(paste('Training Accuracy', 1 - misClasificError))
```

#### NO FILTERS
```{r}
nofilter <- multiVar %>% 
  #filter(acute.chronic_f == "Chronic") %>% 
  ##filter(!environment == "Terrestrial") %>% 
  #filter(bio.org == "organism") %>% 
  #filter(lvl1_f == "Fitness") %>% 
  #filter(!exposure.route == "food") %>% 
  dplyr::select(-c(sex, functional.group, charge, zetapotential.mV, max.size.ingest.mm, media.sal.ppt, media.ph, media.temp, treatments, acute.chronic_f, max.size.ingest.mm, effect.metric, lvl2_f, exposure.route, exposure.duration.d, bio.org, environment, lvl1_f, dose.particles.mL.master.AF.noec, dose.mg.L.master.AF.noec, size_f, density.mg.um.3, dose.particles.mL.master, effect_10, ID)) %>%
  mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
  drop_na()
skim(nofilter)
```
With a complete dataset for just acute studies in aquatic organisms with aqueous route of exposure, we are left with 453 complete cases with 6 predictor variables, 1 response variable (effect y/n), and an ID. Let's determine if we have enough data for ML. 
```{r}
exp(1)^6
```
$e^6$ is less than n (453), so we may proceed.

```{r}
# Create calibration and validation splits with tidymodels initial_split() function.
set.seed(4)
nofilter_split <- nofilter %>%
  initial_split(prop = 0.75, strata = polymer) # splits data into training and testing set.
# default is 3/4ths split (but 75% training, 25% testing).
# Stratification (strata) = grouping training/testing sets by region, state, etc.
# Using the "strata" cnofilter ensures the number of data points in the training data is equivalent to the proportions in the original data set. (Strata below 10% of the total are pooled together.)

# Create a training data set with the training() function
# Pulls from training and testing sets created by initial_split()
nofilter_train <- training(nofilter_split)
nofilter_test <- testing(nofilter_split)
# Examine the environment to be sure # of observations looks like the 75/25 split. 3199:1066.

# Random forest -- 
set.seed(2) # assures the data pulled is random, but sets it for the run below (makes outcome stable)
nofilterrf <- randomForest(y = nofilter_train$effect_f, # dependent variable
  x = nofilter_train %>%
    dplyr::select(-effect_f), # selecting nofilter predictor variables
  importance = T, # how useful is a predictor in predicting values (nothing causal)
  proximity = T, 
  ntrees = 100) # 500 trees default.

nofilterrf # examine the results.
```

```{r}
plot(nofilterrf)
# model performance appears to improve most at ~75 trees
```

```{r}
varImpPlot(nofilterrf)
# displays which variables are most important
# helps to winnow down list of predictors
# recommended to weigh left pane more
# right pane also shows how evenly things split based on the list of predictors
# values close to 0 can be dropped, but don't have to be
```
Dose, organism group, exposure duration are important.

```{r}
importance <- nofilterrf$importance
View(importance)
# displays the data plotted in the plot above
```

```{r fit results}
fitted <- predict(nofilterrf, nofilter_test[,-6], OOB = TRUE, type ="response")
misClasificError <- mean(fitted != nofilter_test$effect_f)
print(paste('Training Accuracy', 1 - misClasificError))
```

#### ROC Curve
yet anothyer try
```{r}
###CHRONIC
chronicpredictions <- as.data.frame(predict(chronicrf, chronic_test[,-6], type = "prob"))
# predict class and then attach test class
chronicpredictions$predict <- names(chronicpredictions)[1:2][apply(chronicpredictions[,1:2], 1, which.max)]
chronicpredictions$observed <- chronic_test$effect_f

####ACUTE
predictions <- as.data.frame(predict(acuterf, acute_test[,-6], type = "prob"))
# predict class and then attach test class
predictions$predict <- names(predictions)[1:2][apply(predictions[,1:2], 1, which.max)]
predictions$observed <- acute_test$effect_f
head(predictions)

#### all
allpredictions <- as.data.frame(predict(allrf, all_test[,-6], type = "prob"))
# predict class and then attach test class
allpredictions$predict <- names(allpredictions)[1:2][apply(allpredictions[,1:2], 1, which.max)]
allpredictions$observed <- all_test$effect_f
head(allpredictions)

###nofilter
nofilterpredictions <- as.data.frame(predict(nofilterrf, nofilter_test[,-6], type = "prob"))
# predict class and then attach test class
nofilterpredictions$predict <- names(nofilterpredictions)[1:2][apply(nofilterpredictions[,1:2], 1, which.max)]
nofilterpredictions$observed <- nofilter_test$effect_f
```
```{r}
# 1 ROC curve, yes vs no for acute
roc.acute <- roc(ifelse(predictions$observed=="Yes", "Yes", "No"), as.numeric(predictions$Yes))

#chronic
roc.chronic <- roc(ifelse(chronicpredictions$observed=="Yes", "Yes", "No"), as.numeric(chronicpredictions$Yes))
#all
roc.all <- roc(ifelse(allpredictions$observed=="Yes", "Yes", "No"), as.numeric(allpredictions$Yes))

#nofilter
roc.nofilter <- roc(ifelse(nofilterpredictions$observed=="Yes", "Yes", "No"), as.numeric(nofilterpredictions$Yes))

all <- ggroc(roc.all) + labs(title = "Chronic and Acute",
                             subtitle = "Accuracy = 81.7%, n = 1510") +
  theme_minimal()
acute <- ggroc(roc.acute, col = "red") + labs(title = "Acute",
                                              subtitle = "Accuracy = 84.8%, n = 883") + theme_minimal()
chronic <- ggroc(roc.chronic, col = "blue") + labs(title = "Chronic",
                                                   subtitle = "Accuracy = 67.3%, n = 452") + theme_minimal()

nofilter<- ggroc(roc.nofilter, col = "green") + labs(title = "Entire Dataset",
                                                   subtitle = "Accuracy = 79.9%, n = 4615") + theme_minimal()
require(gridExtra)
grid.arrange(all, acute, chronic, nofilter,
             ncol = 4)
```

```{r}
#plot reciever operating curve
perf <- performance(preds,"tpr","fpr")
```

```{r}
# compute in-sample results
caret::confusionMatrix(fitted,as.factor(acute_test$effect_f))
```

```{r}
# Validation set assessment #1: looking at confusion matrix
prediction_for_table <- predict(acuterf, acute_test[,-6])
table(observed = acute_test[,6], predicted = prediction_for_table)
```

```{r}
# Validation set assessment #2: ROC curves and AUC
# Needs to import ROCR package for ROC curve plotting:
library(ROCR)
# Calculate the probability of new observations belonging to each class
# prediction_for_roc_curve will be a matrix with dimensions data_set_size x number_of_classes
prediction_for_roc_curve <- predict(acuterf ,acute_test[,-6],type="response")
# Use pretty colours:
pretty_colours <- c("#F8766D","#00BA38","#619CFF")
# Specify the different classes 
classes <- levels(acute_test$organism.group)
# For each class
for (i in 1:13)
{
 # Define which observations belong to class[i]
 true_values <- ifelse(acute_test[,6]==classes[i],"Yes","No")
 # Assess the performance of classifier for class[i]
 pred <- prediction(prediction_for_roc_curve[,i],true_values)
 perf <- performance(pred, "tpr", "fpr")
 if (i==1)
 {
     plot(perf,main="ROC Curve",col=pretty_colours[i]) 
 }
 else
 {
     plot(perf,main="ROC Curve",col=pretty_colours[i],add=TRUE) 
 }
 # Calculate the AUC and print it to screen
 auc.perf <- performance(pred, measure = "auc")
 print(auc.perf@y.values)
}
```

```{r}
# predict()
# returns out of bag predictions for training data
# in the bag: every time a tree is built, it uses ~80% of the original 75% we set aside from the original dataset used to create a tree to assure random data selection
# out of bag: looking at the remaining 20% of the training data to predict, when you want to know what your model does at the training location sites

# # Predict CRAM scores state-wide for all COMIDs.
# notTrain_prediction <- notTrain %>% # taking all available IDs, that haven't been used in training
#   na.omit() %>% # remove NAs
#   mutate(effect_predicted = predict(myrf, newdata = notTrain %>% na.omit())) # using developed model (myrf), inputting predictor variables (notTrain - which contains IDs and associated  data) to predict output/dependent variable (effect_predicted a.k.a. effect).

# rePredict CRAM scores for training data.
multiVar_small_train$effect_predicted <- predict(myrf) # Add column of predicted CRAM values to training dataset.

# Creates new dataset of bound rows for both ...
# effect_predictions <- bind_rows(nottrain_prediction %>%
#                             mutate(Set = "Non-training"), # statewide COMIDs (not used for training)
#                             multiVar_smal_train %>%
#                             mutate(Set = "Training")) # COMIDS from training dataset
# This creates the dataset that will be plotted to create a state-wide plot of predicted CRAM scores.

# Plot the data.
multiVar_small_train %>% 
  mutate(effect_10 = case_when( #convert ordinal to numeric
     effect_predicted == "Yes" ~ 1,
     effect_predicted == "No" ~ 0
   )) %>% 
  ggplot(aes(x = log10(dose.mg.L.master), y = effect_10)) +
  geom_point(alpha = 0.7) +#color = "black") +
  labs(x = "Density",
    y = "Predicted CRAM Score") +
  theme_dark() #+
  #facet_wrap(.~polymer)
```
### Predictor Selector
```{r}
# Using caret to select the best predictors
# What are the parameters you want to use to run recursive feature elimination (rfe)?
my_ctrl <- rfeControl(functions = rfFuncs,
                      method = "cv",
                      verbose = FALSE,
                      returnResamp = "all")

# rfe = recursive feature elimination
# THIS STEP TAKES FOR-EV-ER!!!
set.seed(22)
my_rfe <- rfe(y = rf_dat$effect_f, # set dependent variable
              x = rf_dat %>% select(-effect_f), # set predictor variables
              size = c(1:2, 4, 6, 8, 10, 13), # sets how many variables are in the overall model
              # I have 13 total possible variables, so I've chosen increments of 3 to look at.
              rfeControl = my_ctrl) # pull in control from above

# can you make your model even simpler?
# the following will pick a model with the smallest number of predictor variables based on the tolerance ("tol") that you specify (how much less than the best are you willing to tolerate?)
my_size <- pickSizeTolerance(my_rfe$results, metric = "Accuracy", tol = 1, maximize = F)
# higher tol (~10) gives you less variables
# lower tol (~1) gives you more variables - "I'd like the simplest model within 1% of the best model."
pickVars(my_rfe$variables, size = my_size)
```

```{r}
# Predict scores using the above 2 variables:

# Create re-finalized training dataset and include all possible variables. 
rf_dat2 <- multiVar_small_train %>%  select(dose.mg.L.master, organism.group, effect_f)

set.seed(4) # assures the data pulled is random, but sets it for the run below (makes outcome stable)
myrf2 <- randomForest(y = rf_dat2$effect_f, # dependent variable
  x = rf_dat2 %>%
    select(-effect_f),
  importance = T, 
  proximity = T, 
  ntrees = 500)

myrf2 # examine the results. 
```

```{r}
importance <- as.data.frame(as.table(myrf$importance))

# Nicer ggplot variable importance plot.
vip_plot_a <- importance %>%
 filter(Var2 == "MeanDecreaseAccuracy") %>%
  mutate(Var1 = factor(Var1)) %>%
  mutate(Var1_f = fct_reorder(Var1, Freq)) %>%
  ggplot(aes(x = Freq, y = Var1_f)) +
  geom_point(size = 3, alpha = 0.75, color = "black") +
  labs(x = "Importance",
    y = "Variables") +
  theme_bw()

# vip_plot_b <- importance %>%
#   filter(Var2 == "MeanDecreaseGini") %>%
#   mutate(Var1 = factor(Var1)) %>%
#   mutate(Var1_f = fct_reorder(Var1, Freq)) %>%
#   ggplot(aes(x = Freq, y = Var1_f)) +
#   geom_point(size = 3, alpha = 0.75, color = "black") +
#   labs(x = "Node Purity",
#     y = "Variables") +
#   theme_bw()

vip_plot <- vip_plot_a #+ vip_plot_b

vip_plot
# ggsave("cram_vip_plot.png",
#      path = "/Users/heilil/Desktop/R_figures",
#      width = 25,
#      height = 10,
#      units = "cm"
#    )
```
GINI importance measures the average gain of purity by splits of a given variable. If the variable is useful, it tends to split mixed labeled nodes into pure single class nodes. Splitting by a permuted variables tend neither to increase nor decrease node purities. Permuting a useful variable, tend to give relatively large decrease in mean gini-gain. GINI importance is closely related to the local decision function, that random forest uses to select the best available split. Therefore, it does not take much extra time to compute. On the other hand, mean gini-gain in local splits, is not necessarily what is most useful to measure, in contrary to change of overall model performance. Gini importance is overall inferior to (permutation based) variable importance as it is relatively more biased, more unstable and tend to answer a more indirect question.

### Quantile Regression model
```{r}
# Quantile random forest regression mode, instead of looking at the mode of trees, can compare to 10th, 50th, 90th percentiles etc.

# Need to make a new dataset taking the above results of pickVars into account.
# Create finalized training dataset and include all possible variables. 
qrf_dat <- multiVar_small_train  %>%
  select(-effect_predicted, -ID)
#   select(asci, RdCrsWs, PctAgWs, PctUrbWsRp100, PctOpWsRp100, PctOpWs, DamDensWs, RdDensWs, NABD_DensWs, PctUrbWs, PctUrbCatRp100, RdDensWsRp100, PctOpCat, PctUrbCat, RdDensCat, CBNFWs, PctOpCatRp100, PctAgWsRp100, TRIDensWs, AgKffactWs, FertWs)

 set.seed(20)
 myqrf <- quantregForest(y = qrf_dat$effect_f, # dependent variable
              x = qrf_dat %>%
                  select(-effect_f),
              importance = T,
              proximity = T,
              keep.inbag=T,
              ntrees = 500)

#predict(myqrf) # automatically presents 10th %tile, median, and 90th %tile
```

```{r}
#predict(myqrf, what=c(0.2, 0.3, 0.999)) # to print specific quantiles
```

```{r}
plot(myqrf) # plots the results.
```
# Again appears to improve after ~100 trees.

### Model Validation
#### Categorical Response

#### Continuous Response
```{r}
# Compare predicted vs. actual results, including by size.
# Adding lines of slope=1 and linear models to each plot.
val1 <- multiVar_small_train %>% 
  ggplot(aes(x = effect_predicted, y = effect_f)) +
  geom_point(color = "#2A3927", alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE, color = "#2A3927") +
  labs(x = "Predicted Effect",
    y = "Effect measured",
    title = "Training Data\nn=613") +
  geom_abline(intercept = 0, slope = 1) +
  theme_bw()

val1

lm1 <- lm(cram~cram_predicted, data = mydf2_train2)
summary(lm1)

val2 <- ggplot(mydf2_test2, aes(x = cram_predicted, y = cram)) +
  geom_point(color = "#3793EC", alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE, color = "#3793EC") +
  scale_x_continuous(breaks = c(0.5, 0.7, 0.9)) +
  labs(x = "CRAM predicted",
    y = "CRAM measured",
    title = "Testing Data\nn=202") +
  geom_abline(intercept = 0, slope = 1) +
  theme_bw()

val2

lm2 <- lm(cram~cram_predicted, data = mydf2_test2)
summary(lm2)

# Create the full testing + training dataset to plot together.
mydf2_test2$set <- "Testing"
mydf2_train2$set <- "Training"
full_train_test <- bind_rows(mydf2_test2, mydf2_train2) %>%
  mutate(set_f = factor(set, levels = c("Training", "Testing")))

val3 <- ggplot(full_train_test, aes(x = cram_predicted, y = cram, color = set_f)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE) +
  scale_color_manual(name = "Set", values = c("#2A3927", "#3793EC"), drop = FALSE) +
  labs(x = "CRAM predicted",
    y = "CRAM measured",
    title = "All Data\nn=815") +
  geom_abline(intercept = 0, slope = 1, color = "black") +
  facet_wrap(~PSA6) +
  theme_bw()

val3

val_fig <- (val1 + val2) /
  (val3)

val_fig + plot_annotation(
  title = 'CRAM Random Forest Results',
  subtitle = 'All modeling performed using StreamCAT datasets.',
  caption = 'Linear models are colored according to dataset. Lines of slope = 1 are denoted in black.'
)

# Save figure.
# ggsave("cram_rfmodel_validation.png",
#      path = "/Users/heilil/Desktop/R_figures",
#      width = 35,
#      height = 25,
#      units = "cm"
#    )

lm3 <- lm(cram~cram_predicted, 
  data = full_train_test %>%
    filter(PSA6 == "Central_Valley") %>%
    filter(set_f == "Training"))

lm4 <- lm(cram~cram_predicted, 
  data = full_train_test %>%
    filter(PSA6 == "Central_Valley") %>%
    filter(set_f == "Testing"))

lm5 <- lm(cram~cram_predicted, 
  data = full_train_test %>%
    filter(PSA6 == "Chaparral") %>%
    filter(set_f == "Training"))

lm6 <- lm(cram~cram_predicted, 
  data = full_train_test %>%
    filter(PSA6 == "Chaparral") %>%
    filter(set_f == "Testing"))

lm7 <- lm(cram~cram_predicted, 
  data = full_train_test %>%
    filter(PSA6 == "Deserts_Modoc") %>%
    filter(set_f == "Training"))

lm8 <- lm(cram~cram_predicted, 
  data = full_train_test %>%
    filter(PSA6 == "Deserts_Modoc") %>%
    filter(set_f == "Testing"))

lm9 <- lm(cram~cram_predicted, 
  data = full_train_test %>%
    filter(PSA6 == "North_Coast") %>%
    filter(set_f == "Training"))

lm10 <- lm(cram~cram_predicted, 
  data = full_train_test %>%
    filter(PSA6 == "North_Coast") %>%
    filter(set_f == "Testing"))

lm11 <- lm(cram~cram_predicted, 
  data = full_train_test %>%
    filter(PSA6 == "Sierra") %>%
    filter(set_f == "Training"))

lm12 <- lm(cram~cram_predicted, 
  data = full_train_test %>%
    filter(PSA6 == "Sierra") %>%
    filter(set_f == "Testing"))

lm13 <- lm(cram~cram_predicted, 
  data = full_train_test %>%
    filter(PSA6 == "South_Coast") %>%
    filter(set_f == "Training"))

lm14 <- lm(cram~cram_predicted, 
  data = full_train_test %>%
    filter(PSA6 == "South_Coast") %>%
    filter(set_f == "Testing"))

# Chose not to compute confusion matrix / accuracy score since this is more applicable to categorical ouputs from random forest models -
# Instead, calculated Root Mean Squared Error (RMSE) of both training and test datasets.
# If test RMSE values are much greater than training, then possible the model has been over fit.

predtest <- predict(myrf2, mydf2_test2)
rmse(mydf2_test2$cram,predtest)
# 9.03

predtrain <- predict(myrf2, mydf2_train2)
rmse(mydf2_train2$cram,predtrain)
# 4.20

# Double checking using the original random forest dataset (rf_dat) with all 35 possible variables included to see where the error in number of predictors starts to increase dramatically (to help double check our decision to include 25 parameters).
dc <- rfcv(rf_dat %>%
    select(-cram), 
  rf_dat$cram,
  step = 0.7, # default is 0.5
  scale="log")

dc$error.cv
#34        24        17        12         8         6         4         3         2 
# 96.28062  97.29366  99.08743 102.93752 115.52580 120.63049 122.09563 123.74425 139.30181 
# 1 
# 206.58921

# Appears between 34 and 24 variables, there is an insignificant increase in error.
# However, this model is much larger than the CSCI (20) and ASCI (10) models, so we may decide to trim this down in the future.

```

# Works Cited

# Non-linear Dose Response
If we hold all paramaters constant except for dose, we should observe a non-linear dose-response behavior that can be best predicted by a 3- or 4-parameter Hill Equation. The most available information is likely for polystyrene spheres in crustaceans.There are 21 studies available.