### 2021-02-06 Stepwise Multiple Regressions.Rmd
Modeling with acute/chronic and life stage added

---
title: "Stepwise Multiple Regressions"
author: "Leah Thornton Hampton"
date: "2/1/2021"
output: html_document
---
# Load packages and data import

```{r, echo = FALSE, message = FALSE}
#Packages
library(tidyverse)
library(MASS) #Stepwise regression package

#Data Import
aoc <- read_csv("AquaticOrganisms_Clean_final.csv", guess_max = 10000)

```

# Data tidying

```{r, echo = false, message = FALSE}

#This is the same tidying and setup as found in app.R as of 2/1/20
aoc_setup <- aoc %>% 
  mutate(effect_f = factor(case_when(effect == "Y" ~ "Yes",
    effect == "N" ~ "No"),
    levels = c("No", "Yes"))) %>%
  replace_na(list(size.category = 0, shape = "Not Reported", polymer = "Not Reported", life.stage = "Not Reported")) %>% 
  mutate(size_f = factor(case_when(
    size.category == 1 ~ "1nm < 100nm",
    size.category == 2 ~ "100nm < 1µm",
    size.category == 3 ~ "1µm < 100µm",
    size.category == 4 ~ "100µm < 1mm",
    size.category == 5 ~ "1mm < 5mm",
    size.category == 0 ~ "Not Reported"),
    levels = c("1nm < 100nm", "100nm < 1µm", "1µm < 100µm", "100µm < 1mm", "1mm < 5mm", "Not Reported"))) %>% 
  mutate(shape_f = factor(case_when(
    shape == "fiber" ~ "Fiber",
    shape == "fragment" ~ "Fragment",
    shape == "sphere" ~ "Sphere",
    shape == "Not Reported" ~ "Not Reported"),
    levels = c("Fiber", "Fragment", "Sphere", "Not Reported"))) %>% 
  mutate(poly_f = factor(case_when(
    polymer == "BIO" ~ "Biopolymer",
    polymer == "EVA" ~ "Polyethylene Vinyl Acetate",
    polymer == "LTX" ~ "Latex",
    polymer == "PA" ~ "Polyamide",
    polymer == "PE" ~ "Polyethylene",
    polymer == "PC" ~ "Polycarbonate",
    polymer == "PET" ~ "Polyethylene Terephthalate",
    polymer == "PI" ~ "Polyisoprene",
    polymer == "PMMA" ~ "Polymethylmethacrylate",
    polymer == "PP" ~ "Polypropylene",
    polymer == "PS" ~ "Polystyrene",
    polymer == "PUR" ~ "Polyurethane",
    polymer == "PVC" ~ "Polyvinylchloride",
    polymer == "PLA" ~ "Polylactic Acid",
    polymer == "Not Reported" ~ "Not Reported"))) %>%
  mutate(org_f = factor(organism.group, levels = c("Algae", "Annelida", "Bacterium", "Cnidaria", "Crustacea",
                                                   "Echinoderm", "Fish", "Insect", "Mollusca", "Nematoda", "Plant", "Rotifera", "Mixed"))) %>% 
  mutate(lvl1_f = factor(case_when(lvl1 == "alimentary.excretory" ~ "Alimentary, Excretory",
    lvl1 == "behavioral.sense.neuro" ~ "Behavioral, Sensory, Neurological",
    lvl1 == "circulatory.respiratory" ~ "Circulatory, Respiratory",
    lvl1 == "community" ~ "Community",
    lvl1 == "fitness" ~ "Fitness",
    lvl1 == "immune" ~ "Immune",
    lvl1 == "metabolism" ~ "Metabolism",
    lvl1 == "microbiome" ~ "Microbiome",
    lvl1 == "stress" ~ "Stress"))) %>% 
  mutate(lvl2_f = factor(case_when(lvl2 == "abundance"~"Abundance",
    lvl2 == "actinobacteria" ~ "Actinobacteria",
    lvl2 == "aggressivity"~"Agressivity",
    lvl2 == "ammonia.excretion" ~ "Ammonia Excretion",
    lvl2 == "bacteroidetes"~ "Bacteriodetes",
    lvl2 == "blood"~"Blood",
    lvl2 == "body.condition"~"Body Condition",
    lvl2 == "boldness"~"Boldness",
    lvl2 == "brain.histo"~"Brain Histological Abnormalities",
    lvl2 == "burrowing"~"Burrowing",
    lvl2 == "carb.metabolism"~"Carb Metabolism",
    lvl2 == "chemokines.cytokines"~"Chemokines",
    lvl2 == "circulatory"~"Circulatory",
    lvl2 == "detoxification"~"Detoxification",
    lvl2 == "development"~"Development",
    lvl2 == "digestion"~"Digestion",
    lvl2 == "digestive.enzymes"~"Digestive Enzymes",
    lvl2 == "digestive.tract.histo"~"Digestive Tract Histological Abnormalities",
    lvl2 == "diversity"~ "Diversity",
    lvl2 == "feeding"~ "Feeding",
    lvl2 == "firmicutes"~ "Firmicutes",
    lvl2 == "gall.bladder.histo" ~ "Gall Bladder Histological Abnormalities",
    lvl2 == "gen.metabolism"~ "General Metabolism",
    lvl2 == "gill.histo"~ "Gill Histological Abnormalities",
    lvl2 == "gonad.histo"~"Gonad Histological Abnormalities",
    lvl2 == "growth"~ "Growth",
    lvl2 == "immune.cells"~"Immune Cells",
    lvl2 == "immune.other"~"Immune Other ",
    lvl2 == "intestinal.permeability"~"Intestinal Permeability",
    lvl2 == "kidney.histo"~"Kidney Histological abnormalities",
    lvl2 == "lipid.metabolism"~"Lipid Metabolism",
    lvl2 == "liver.histo"~"Liver Histological Abnormalities",
    lvl2 == "liver.kidney.products" ~ "Liver and Kidney Products",
    lvl2 == "locomotion"~"Locomotion",
    lvl2 == "mortality"~"Mortality",
    lvl2 == "nervous.system"~"Nervous System",
    lvl2 == "oxidative.stress"~"Oxidative Stress",
    lvl2 == "photosynthesis"~ "Photosynthesis",
    lvl2 == "proteobacteria"~"Protebacteria",
    lvl2 == "reproduction"~"Reproduction",
    lvl2 == "respiration"~"Respiration",
    lvl2 == "sexhormones"~"Sex Hormones",
    lvl2 == "shoaling"~"Shoaling",
    lvl2 == "stress"~"Stress",
    lvl2 == "vision.system"~"Vision System"))) %>% 
  mutate(bio_f = factor(case_when(bio.org == "cell"~"Cell", 
    bio.org == "organism"~"Organism",
    bio.org == "population"~ "Population",
    bio.org == "subcell"~"Subcell",
    bio.org == "tissue" ~ "Tissue")))%>%
  mutate(vivo_f = factor(case_when(invitro.invivo == "invivo"~"In Vivo",
    invitro.invivo == "invitro"~"In Vitro")))%>% 
  mutate(life_f = factor(case_when(life.stage == "Early"~"Early",
    life.stage == "Juvenile"~"Juvenile",
    life.stage == "Adult"~"Adult",
    life.stage == "Not Reported"~"Not Reported")))%>% 
  mutate(env_f = factor(case_when(environment == "Freshwater"~"Freshwater",
    environment == "Marine" ~ "Marine",
    environment == "Terrestrial" ~ "Terrestrial"))) %>%
  mutate(species_f = as.factor(paste(genus,species))) %>% 
  mutate(dose.mg.L.master.converted.reported = factor(dose.mg.L.master.converted.reported)) %>%
  mutate(dose.particles.mL.master.converted.reported = factor(dose.particles.mL.master.converted.reported)) %>% 
  mutate(effect.metric = factor(effect.metric)) %>%
  mutate(dose.um3.mL.master = particle.volume.um3 * dose.particles.mL.master) %>%  
  mutate(af.time_noNA = replace_na(af.time, "Unavailable")) %>% 
  mutate(acute.chronic_f = factor(case_when(af.time_noNA == 10 ~ "Acute",
                                            af.time_noNA == 1 ~ "Chronic",
                                            af.time_noNA == "Unavailable" ~ "Unavailable"))) %>% 
  mutate(charge_f = factor(case_when(charge == "negative" ~ "Negative",
                                     charge == "positive" ~ "Positive",
                                     charge == "Not Reported" ~ "Not Reported"))) %>%
  mutate(particle_source_f = factor(case_when(particle.source == "commercial" ~ "Commercial",
                                              particle.source == "lab" ~ "Lab",
                                              particle.source == "modified" ~ "Modified",
                                              particle.source == "N" ~ "Not Reported")))

```

#Variables to put into model

* Dose (counts)
* Dose (mass)
* Dose (volume)
* Size (continuous)
* Size (binned)
* Polymer type
* Shape
* Organism Group (this is being used as a temporary substitute for body size)
* Life Stage
* Level of Biological Organization
* Exposure Duration
* Acute/Chronic (this values are only complete for fish, molluscs, crustaceans and algae)
* Charge (negative/positive - categorical)
* Zeta potential 
* Particle Source (commercial, generated in-house, or mmodified commercial particles e.g., milling)

Response Variables
* Effect (Y/N) - Binary Data
* Effect (Y/N) - Binary Data x Effect Score (binned organism level effects)

#Continous Variable Distributions

There are several continuous variables that we want to feed into our model. Before doing so, we're going to check the distribution to see if any of them are skewed and need to be transformed.

```{r, echo = FALSE}

#Dose - Count

Dose_Particles_mL <- aoc_setup$dose.particles.mL.master

hist(Dose_Particles_mL)

hist(log10(Dose_Particles_mL))

#Dose - Mass

Dose_Mass_L <- aoc_setup$dose.mg.L.master

hist(Dose_Mass_L)

hist(log10(Dose_Mass_L))

#Dose - Volume

`Dose_um^3_mL` <- aoc_setup$dose.um3.mL.master

hist(`Dose_um^3_mL`)

hist(log10(`Dose_um^3_mL`))

#Size - Continuous

Size <- aoc_setup$size.length.um.used.for.conversions

hist(Size)

hist(log10(Size))

#Exposure Duration

Exposure_Duration <- aoc_setup$exposure.duration.d

hist(Exposure_Duration)

hist(log10(Exposure_Duration))

#Zeta Potential

Zeta_Potential_mV <- aoc_setup$zetapotential.mV

hist(Zeta_Potential_mV)

#No need to log transform

aoc_setup_select <- aoc_setup %>% 
  mutate(log10(dose.particles.mL.master)) %>% 
  mutate(log10(dose.mg.L.master)) %>%
  mutate(log10(dose.um3.mL.master)) %>%
  mutate(log10(size.length.um.used.for.conversions)) %>%
  mutate(log10(exposure.duration.d))

```

Due to skewed data, the following categories are log10 transformed before modeling:

* Dose (counts)
* Dose (mass)
* Dose (volume)
* Size (continuous)
* Exposure Duration

# Model: All Independent Variables, Response Variable: Effect (Y/N)

Full Model

```{r, echo = FALSE}

#Select the columns that we want to feed into the model for simplicity
aoc_setup_select_1 <- aoc_setup_select %>% 
  dplyr::select(
    `log10(dose.particles.mL.master)`, 
    `log10(dose.mg.L.master)`,
    `log10(dose.um3.mL.master)`,
    `log10(size.length.um.used.for.conversions)`, 
    size_f,     
    poly_f, 
    shape_f,
    org_f,
    life_f,
    bio_f,
    `log10(exposure.duration.d)`,
    acute.chronic_f,
    charge_f,
    zetapotential.mV,
    particle_source_f,
    effect_f)

#Convert all data to numerical data
aoc_setup_select_num <- data.matrix(aoc_setup_select_1)

#Convert data matrix into a data frame
x <- as.data.frame(aoc_setup_select_num)

#Drop NAs to get model to be happy - this might be sketchy!?
y<- na.omit(x)

#Fit the full model
Full_Model <- lm(effect_f ~., data = y,
                 na.action = na.omit)
summary(Full_Model)
```

Stepwise Model - Both Directions

```{r, echo = FALSE}
#Stepwise model
step.model <- stepAIC(Full_Model, direction = "both", 
                      trace = FALSE)

summary(step.model)

```

## Remove Charge and Zeta Potential

Missing categories (NA) means that some data are incomplete. This is likely charge and zeta potential which are only in the database for PS and PC spheres. Removed from model to retain more data.

Full Model

```{r, echo = FALSE}

#Select the columns that we want to feed into the model for simplicity
aoc_setup_select2 <- aoc_setup_select %>% 
  dplyr::select(
    `log10(dose.particles.mL.master)`, 
    `log10(dose.mg.L.master)`,
    `log10(dose.um3.mL.master)`,
    `log10(size.length.um.used.for.conversions)`, 
    size_f,     
    poly_f, 
    shape_f,
    org_f,
    life_f,
    bio_f,
    `log10(exposure.duration.d)`,
    acute.chronic_f,
    # charge_f,
    # zetapotential.mV,
    particle_source_f,
    effect_f)

#Convert all data to numerical data
aoc_setup_select_num <- data.matrix(aoc_setup_select2)

#Convert data matrix into a data frame
x <- as.data.frame(aoc_setup_select_num)

#Drop NAs to get model to be happy - this might be sketchy!?
y<- na.omit(x)

#Fit the full model
Full_Model <- lm(effect_f ~., data = y,
                 na.action = na.omit)
summary(Full_Model)
```

Stepwise Model - Both Directions

```{r, echo = FALSE}
#Stepwise model
step.model <- stepAIC(Full_Model, direction = "both", 
                      trace = FALSE)

summary(step.model)

```

## Remove Exposure Duration

Exposure duration not significant - removed from model.

Full Model

```{r, echo = FALSE}

#Select the columns that we want to feed into the model for simplicity
aoc_setup_select3 <- aoc_setup_select %>% 
  dplyr::select(
    `log10(dose.particles.mL.master)`, 
    `log10(dose.mg.L.master)`,
    `log10(dose.um3.mL.master)`,
    `log10(size.length.um.used.for.conversions)`, 
    size_f,     
    poly_f, 
    shape_f,
    org_f,
    life_f,
    bio_f,
    # `log10(exposure.duration.d)`,
    acute.chronic_f,
    # charge_f,
    # zetapotential.mV,
    particle_source_f,
    effect_f)

#Convert all data to numerical data
aoc_setup_select_num <- data.matrix(aoc_setup_select3)

#Convert data matrix into a data frame
x <- as.data.frame(aoc_setup_select_num)

#Drop NAs to get model to be happy - this might be sketchy!?
y<- na.omit(x)

#Fit the full model
Full_Model <- lm(effect_f ~., data = y,
                 na.action = na.omit)
summary(Full_Model)
```

Stepwise Model - Both Directions

```{r, echo = FALSE}
#Stepwise model
step.model <- stepAIC(Full_Model, direction = "both", 
                      trace = FALSE)

summary(step.model)

```

## Remove Size (binned)

Size (binned) not significant - removed from model.

Full Model

```{r, echo = FALSE}

#Select the columns that we want to feed into the model for simplicity
aoc_setup_select4 <- aoc_setup_select %>% 
  dplyr::select(
    `log10(dose.particles.mL.master)`, 
    `log10(dose.mg.L.master)`,
    `log10(dose.um3.mL.master)`,
    `log10(size.length.um.used.for.conversions)`, 
    # size_f,     
    poly_f, 
    shape_f,
    org_f,
    life_f,
    bio_f,
    # `log10(exposure.duration.d)`,
    acute.chronic_f,
    # charge_f,
    # zetapotential.mV,
    particle_source_f,
    effect_f)

#Convert all data to numerical data
aoc_setup_select_num <- data.matrix(aoc_setup_select4)

#Convert data matrix into a data frame
x <- as.data.frame(aoc_setup_select_num)

#Drop NAs to get model to be happy - this might be sketchy!?
y<- na.omit(x)

#Fit the full model
Full_Model <- lm(effect_f ~., data = y,
                 na.action = na.omit)
summary(Full_Model)
```

Stepwise Model - Both Directions

```{r, echo = FALSE}
#Stepwise model
step.model <- stepAIC(Full_Model, direction = "both", 
                      trace = FALSE)

summary(step.model)

```

## Remove Particle Source and Life Stage

Particle Source and Life Stage - removed from model.

Full Model

```{r, echo = FALSE}

#Select the columns that we want to feed into the model for simplicity
aoc_setup_select5 <- aoc_setup_select %>% 
  dplyr::select(
    `log10(dose.particles.mL.master)`, 
    `log10(dose.mg.L.master)`,
    `log10(dose.um3.mL.master)`,
    `log10(size.length.um.used.for.conversions)`, 
    # size_f,     
    poly_f, 
    shape_f,
    org_f,
    # life_f,
    bio_f,
    # `log10(exposure.duration.d)`,
    acute.chronic_f,
    # charge_f,
    # zetapotential.mV,
    # particle_source_f,
    effect_f)

#Convert all data to numerical data
aoc_setup_select_num <- data.matrix(aoc_setup_select5)

#Convert data matrix into a data frame
x <- as.data.frame(aoc_setup_select_num)

#Drop NAs to get model to be happy - this might be sketchy!?
y<- na.omit(x)

#Fit the full model
Full_Model <- lm(effect_f ~., data = y,
                 na.action = na.omit)
summary(Full_Model)
```

Stepwise Model - Both Directions

```{r, echo = FALSE}
#Stepwise model
step.model <- stepAIC(Full_Model, direction = "both", 
                      trace = FALSE)

summary(step.model)

```

## All Variables: Acute

Add all variables back in but only use acute data (fish, crustaceans, mollucs, and algae)

Full Model

```{r, echo = FALSE}

#Select the columns that we want to feed into the model for simplicity
aoc_setup_select5 <- aoc_setup_select %>% 
  filter(acute.chronic_f == "Acute") %>% 
  dplyr::select(
    `log10(dose.particles.mL.master)`, 
    `log10(dose.mg.L.master)`,
    `log10(dose.um3.mL.master)`,
    `log10(size.length.um.used.for.conversions)`, 
    size_f,
    poly_f, 
    shape_f,
    org_f,
    life_f,
    bio_f,
    `log10(exposure.duration.d)`,
    charge_f,
    zetapotential.mV,
    particle_source_f,
    effect_f)

#Convert all data to numerical data
aoc_setup_select_num <- data.matrix(aoc_setup_select5)

#Convert data matrix into a data frame
x <- as.data.frame(aoc_setup_select_num)

#Drop NAs to get model to be happy - this might be sketchy!?
y<- na.omit(x)

#Fit the full model
Full_Model <- lm(effect_f ~., data = y,
                 na.action = na.omit)
summary(Full_Model)
```

Stepwise Model - Both Directions

```{r, echo = FALSE}
#Stepwise model
step.model <- stepAIC(Full_Model, direction = "both", 
                      trace = FALSE)

summary(step.model)

```

## All Variables: Chronic

Add all variables back in but only use chronic data (fish, crustaceans, mollucs, and algae)

Full Model

```{r, echo = FALSE}

#Select the columns that we want to feed into the model for simplicity
aoc_setup_select6 <- aoc_setup_select %>% 
  filter(acute.chronic_f == "Chronic") %>% 
  dplyr::select(
    `log10(dose.particles.mL.master)`, 
    `log10(dose.mg.L.master)`,
    `log10(dose.um3.mL.master)`,
    `log10(size.length.um.used.for.conversions)`, 
    size_f,
    poly_f, 
    shape_f,
    org_f,
    life_f,
    bio_f,
    `log10(exposure.duration.d)`,
    charge_f,
    zetapotential.mV,
    particle_source_f,
    effect_f)

#Convert all data to numerical data
aoc_setup_select_num <- data.matrix(aoc_setup_select6)

#Convert data matrix into a data frame
x <- as.data.frame(aoc_setup_select_num)

#Drop NAs to get model to be happy - this might be sketchy!?
y<- na.omit(x)

#Fit the full model
Full_Model <- lm(effect_f ~., data = y,
                 na.action = na.omit)
summary(Full_Model)
```

Stepwise Model - Both Directions

```{r, echo = FALSE}
#Stepwise model
step.model <- stepAIC(Full_Model, direction = "both", 
                      trace = FALSE)

summary(step.model)

```

# Model: All Independent Variables, Response Variable: Effect (Y/N) x Effect Score

Full Model

```{r, echo = FALSE}

#Select the columns that we want to feed into the model for simplicity
aoc_setup_select_1 <- aoc_setup_select %>% 
  dplyr::select(
    `log10(dose.particles.mL.master)`, 
    `log10(dose.mg.L.master)`,
    `log10(dose.um3.mL.master)`,
    `log10(size.length.um.used.for.conversions)`, 
    size_f,     
    poly_f, 
    shape_f,
    org_f,
    life_f,
    bio_f,
    `log10(exposure.duration.d)`,
    acute.chronic_f,
    charge_f,
    zetapotential.mV,
    particle_source_f,
    effect_f,
    effect.score)

#Convert all data to numerical data
aoc_setup_select_num <- data.matrix(aoc_setup_select_1)

#Convert data matrix into a data frame
x <- as.data.frame(aoc_setup_select_num)

#Drop NAs to get model to be happy - this might be sketchy!?
y<- na.omit(x)

#Fit the full model
Full_Model <- lm(effect_f*effect.score ~., data = y,
                 na.action = na.omit)
summary(Full_Model)
```

Stepwise Model - Both Directions

```{r, echo = FALSE}
#Stepwise model
step.model <- stepAIC(Full_Model, direction = "both", 
                      trace = FALSE)

summary(step.model)

```

## Remove Charge and Zeta Potential

Missing categories (NA) means that some data are incomplete. This is likely charge and zeta potential which are only in the database for PS and PC spheres. Removed from model to retain more data.

Full Model

```{r, echo = FALSE}

#Select the columns that we want to feed into the model for simplicity
aoc_setup_select_2 <- aoc_setup_select %>% 
  dplyr::select(
    `log10(dose.particles.mL.master)`, 
    `log10(dose.mg.L.master)`,
    `log10(dose.um3.mL.master)`,
    `log10(size.length.um.used.for.conversions)`, 
    size_f,     
    poly_f, 
    shape_f,
    org_f,
    life_f,
    bio_f,
    `log10(exposure.duration.d)`,
    acute.chronic_f,
    # charge_f,
    # zetapotential.mV,
    particle_source_f,
    effect_f,
    effect.score)

#Convert all data to numerical data
aoc_setup_select_num <- data.matrix(aoc_setup_select_2)

#Convert data matrix into a data frame
x <- as.data.frame(aoc_setup_select_num)

#Drop NAs to get model to be happy - this might be sketchy!?
y<- na.omit(x)

#Fit the full model
Full_Model <- lm(effect_f*effect.score ~., data = y,
                 na.action = na.omit)
summary(Full_Model)
```

Stepwise Model - Both Directions

```{r, echo = FALSE}
#Stepwise model
step.model <- stepAIC(Full_Model, direction = "both", 
                      trace = FALSE)

summary(step.model)

```

## Remove Size (binned)

Size (binned) not significant - removed from model.

Full Model

```{r, echo = FALSE}

#Select the columns that we want to feed into the model for simplicity
aoc_setup_select_3 <- aoc_setup_select %>% 
  dplyr::select(
    `log10(dose.particles.mL.master)`, 
    `log10(dose.mg.L.master)`,
    `log10(dose.um3.mL.master)`,
    `log10(size.length.um.used.for.conversions)`, 
    # size_f,     
    poly_f, 
    shape_f,
    org_f,
    life_f,
    bio_f,
    `log10(exposure.duration.d)`,
    acute.chronic_f,
    # charge_f,
    # zetapotential.mV,
    particle_source_f,
    effect_f,
    effect.score)

#Convert all data to numerical data
aoc_setup_select_num <- data.matrix(aoc_setup_select_3)

#Convert data matrix into a data frame
x <- as.data.frame(aoc_setup_select_num)

#Drop NAs to get model to be happy - this might be sketchy!?
y<- na.omit(x)

#Fit the full model
Full_Model <- lm(effect_f*effect.score ~., data = y,
                 na.action = na.omit)
summary(Full_Model)
```

Stepwise Model - Both Directions

```{r, echo = FALSE}
#Stepwise model
step.model <- stepAIC(Full_Model, direction = "both", 
                      trace = FALSE)

summary(step.model)

```

## Remove Particle Source

Particle source not significant - removed from model.

Full Model

```{r, echo = FALSE}

#Select the columns that we want to feed into the model for simplicity
aoc_setup_select_4 <- aoc_setup_select %>% 
  dplyr::select(
    `log10(dose.particles.mL.master)`, 
    `log10(dose.mg.L.master)`,
    `log10(dose.um3.mL.master)`,
    `log10(size.length.um.used.for.conversions)`, 
    # size_f,     
    poly_f, 
    shape_f,
    org_f,
    life_f,
    bio_f,
    `log10(exposure.duration.d)`,
    acute.chronic_f,
    # charge_f,
    # zetapotential.mV,
    # particle_source_f,
    effect_f,
    effect.score)

#Convert all data to numerical data
aoc_setup_select_num <- data.matrix(aoc_setup_select_4)

#Convert data matrix into a data frame
x <- as.data.frame(aoc_setup_select_num)

#Drop NAs to get model to be happy - this might be sketchy!?
y<- na.omit(x)

#Fit the full model
Full_Model <- lm(effect_f*effect.score ~., data = y,
                 na.action = na.omit)
summary(Full_Model)
```

Stepwise Model - Both Directions

```{r, echo = FALSE}
#Stepwise model
step.model <- stepAIC(Full_Model, direction = "both", 
                      trace = FALSE)

summary(step.model)

```