### 2021-02-10 Stepwise Multiple Regressions.Rmd

---
title: "Stepwise Multiple Regressions"
author: "Leah Thornton Hampton"
date: "2/1/2021"
output:
  html_document: 
    toc: true
    toc_depth: 4
    number_sections: true
    toc_float: true
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
#Packages
library(tidyverse)
library(MASS, exclude = 'select') #Stepwise regression package 
library(ggiraphExtra) #For plotting

#Data Import
aoc <- read_csv("AquaticOrganisms_Clean_final.csv", guess_max = 10000)

```

# Set Up

```{r, echo = FALSE, message = FALSE}

#This is the same tidying and setup as found in app.R as of 2/1/20
aoc_setup <- aoc %>% 
  mutate(effect_f = factor(case_when(effect == "Y" ~ "Yes",
    effect == "N" ~ "No"),
    levels = c("No", "Yes"))) %>%
  mutate(effect_10 = case_when(
     effect_f == "Yes" ~ 1,
     effect_f == "No" ~ 0)) %>% 
  replace_na(list(size.category = 0, shape = "Not Reported", polymer = "Not Reported", life.stage = "Not Reported")) %>% 
  mutate(size_f = factor(case_when(
    size.category == 1 ~ "1nm < 100nm",
    size.category == 2 ~ "100nm < 1µm",
    size.category == 3 ~ "1µm < 100µm",
    size.category == 4 ~ "100µm < 1mm",
    size.category == 5 ~ "1mm < 5mm",
    size.category == 0 ~ "Not Reported"),
    levels = c("1nm < 100nm", "100nm < 1µm", "1µm < 100µm", "100µm < 1mm", "1mm < 5mm", "Not Reported"))) %>% 
  mutate(shape_f = factor(case_when(
    shape == "fiber" ~ "Fiber",
    shape == "fragment" ~ "Fragment",
    shape == "sphere" ~ "Sphere",
    shape == "Not Reported" ~ "Not Reported"),
    levels = c("Fiber", "Fragment", "Sphere", "Not Reported"))) %>% 
  mutate(poly_f = factor(case_when(
    polymer == "BIO" ~ "Biopolymer",
    polymer == "EVA" ~ "Polyethylene Vinyl Acetate",
    polymer == "LTX" ~ "Latex",
    polymer == "PA" ~ "Polyamide",
    polymer == "PE" ~ "Polyethylene",
    polymer == "PC" ~ "Polycarbonate",
    polymer == "PET" ~ "Polyethylene Terephthalate",
    polymer == "PI" ~ "Polyisoprene",
    polymer == "PMMA" ~ "Polymethylmethacrylate",
    polymer == "PP" ~ "Polypropylene",
    polymer == "PS" ~ "Polystyrene",
    polymer == "PUR" ~ "Polyurethane",
    polymer == "PVC" ~ "Polyvinylchloride",
    polymer == "PLA" ~ "Polylactic Acid",
    polymer == "Not Reported" ~ "Not Reported"))) %>%
  mutate(org_f = factor(organism.group, levels = c("Algae", "Annelida", "Bacterium", "Cnidaria", "Crustacea",
                                                   "Echinoderm", "Fish", "Insect", "Mollusca", "Nematoda", "Plant", "Rotifera", "Mixed"))) %>% 
  mutate(lvl1_f = factor(case_when(lvl1 == "alimentary.excretory" ~ "Alimentary, Excretory",
    lvl1 == "behavioral.sense.neuro" ~ "Behavioral, Sensory, Neurological",
    lvl1 == "circulatory.respiratory" ~ "Circulatory, Respiratory",
    lvl1 == "community" ~ "Community",
    lvl1 == "fitness" ~ "Fitness",
    lvl1 == "immune" ~ "Immune",
    lvl1 == "metabolism" ~ "Metabolism",
    lvl1 == "microbiome" ~ "Microbiome",
    lvl1 == "stress" ~ "Stress"))) %>% 
  mutate(lvl2_f = factor(case_when(lvl2 == "abundance"~"Abundance",
    lvl2 == "actinobacteria" ~ "Actinobacteria",
    lvl2 == "aggressivity"~"Agressivity",
    lvl2 == "ammonia.excretion" ~ "Ammonia Excretion",
    lvl2 == "bacteroidetes"~ "Bacteriodetes",
    lvl2 == "blood"~"Blood",
    lvl2 == "body.condition"~"Body Condition",
    lvl2 == "boldness"~"Boldness",
    lvl2 == "brain.histo"~"Brain Histological Abnormalities",
    lvl2 == "burrowing"~"Burrowing",
    lvl2 == "carb.metabolism"~"Carb Metabolism",
    lvl2 == "chemokines.cytokines"~"Chemokines",
    lvl2 == "circulatory"~"Circulatory",
    lvl2 == "detoxification"~"Detoxification",
    lvl2 == "development"~"Development",
    lvl2 == "digestion"~"Digestion",
    lvl2 == "digestive.enzymes"~"Digestive Enzymes",
    lvl2 == "digestive.tract.histo"~"Digestive Tract Histological Abnormalities",
    lvl2 == "diversity"~ "Diversity",
    lvl2 == "feeding"~ "Feeding",
    lvl2 == "firmicutes"~ "Firmicutes",
    lvl2 == "gall.bladder.histo" ~ "Gall Bladder Histological Abnormalities",
    lvl2 == "gen.metabolism"~ "General Metabolism",
    lvl2 == "gill.histo"~ "Gill Histological Abnormalities",
    lvl2 == "gonad.histo"~"Gonad Histological Abnormalities",
    lvl2 == "growth"~ "Growth",
    lvl2 == "immune.cells"~"Immune Cells",
    lvl2 == "immune.other"~"Immune Other ",
    lvl2 == "intestinal.permeability"~"Intestinal Permeability",
    lvl2 == "kidney.histo"~"Kidney Histological abnormalities",
    lvl2 == "lipid.metabolism"~"Lipid Metabolism",
    lvl2 == "liver.histo"~"Liver Histological Abnormalities",
    lvl2 == "liver.kidney.products" ~ "Liver and Kidney Products",
    lvl2 == "locomotion"~"Locomotion",
    lvl2 == "mortality"~"Mortality",
    lvl2 == "nervous.system"~"Nervous System",
    lvl2 == "oxidative.stress"~"Oxidative Stress",
    lvl2 == "photosynthesis"~ "Photosynthesis",
    lvl2 == "proteobacteria"~"Protebacteria",
    lvl2 == "reproduction"~"Reproduction",
    lvl2 == "respiration"~"Respiration",
    lvl2 == "sexhormones"~"Sex Hormones",
    lvl2 == "shoaling"~"Shoaling",
    lvl2 == "stress"~"Stress",
    lvl2 == "vision.system"~"Vision System"))) %>% 
  mutate(bio_f = factor(case_when(bio.org == "cell"~"Cell", 
    bio.org == "organism"~"Organism",
    bio.org == "population"~ "Population",
    bio.org == "subcell"~"Subcell",
    bio.org == "tissue" ~ "Tissue")))%>%
  mutate(vivo_f = factor(case_when(invitro.invivo == "invivo"~"In Vivo",
    invitro.invivo == "invitro"~"In Vitro")))%>% 
  mutate(life_f = factor(case_when(life.stage == "Early"~"Early",
    life.stage == "Juvenile"~"Juvenile",
    life.stage == "Adult"~"Adult",
    life.stage == "Not Reported"~"Not Reported")))%>% 
  mutate(env_f = factor(case_when(environment == "Freshwater"~"Freshwater",
    environment == "Marine" ~ "Marine",
    environment == "Terrestrial" ~ "Terrestrial"))) %>%
  mutate(species_f = as.factor(paste(genus,species))) %>% 
  mutate(dose.mg.L.master.converted.reported = factor(dose.mg.L.master.converted.reported)) %>%
  mutate(dose.particles.mL.master.converted.reported = factor(dose.particles.mL.master.converted.reported)) %>% 
  mutate(effect.metric = factor(effect.metric)) %>%
  mutate(dose.um3.mL.master = particle.volume.um3 * dose.particles.mL.master) %>%  
  mutate(af.time_noNA = replace_na(af.time, "Unavailable")) %>% 
  mutate(acute.chronic_f = factor(case_when(af.time_noNA == 10 ~ "Acute",
                                            af.time_noNA == 1 ~ "Chronic",
                                            af.time_noNA == "Unavailable" ~ "Unavailable"))) %>% 
  mutate(charge_f = factor(case_when(charge == "negative" ~ "Negative",
                                     charge == "positive" ~ "Positive",
                                     charge == "Not Reported" ~ "Not Reported"))) %>%
  mutate(particle_source_f = factor(case_when(particle.source == "commercial" ~ "Commercial",
                                              particle.source == "lab" ~ "Lab",
                                              particle.source == "modified" ~ "Modified",
                                              particle.source == "N" ~ "Not Reported")))

```

```{r}

#Custom function that finds factors with only 1 level
#Source: https://stackoverflow.com/questions/44200195/how-to-debug-contrasts-can-be-applied-only-to-factors-with-2-or-more-levels-er

debug_contr_error <- function (dat, subset_vec = NULL) {
  if (!is.null(subset_vec)) {
    ## step 0
    if (mode(subset_vec) == "logical") {
      if (length(subset_vec) != nrow(dat)) {
        stop("'logical' `subset_vec` provided but length does not match `nrow(dat)`")
        }
      subset_log_vec <- subset_vec
      } else if (mode(subset_vec) == "numeric") {
      ## check range
      ran <- range(subset_vec)
      if (ran[1] < 1 || ran[2] > nrow(dat)) {
        stop("'numeric' `subset_vec` provided but values are out of bound")
        } else {
        subset_log_vec <- logical(nrow(dat))
        subset_log_vec[as.integer(subset_vec)] <- TRUE
        } 
      } else {
      stop("`subset_vec` must be either 'logical' or 'numeric'")
      }
    dat <- base::subset(dat, subset = subset_log_vec)
    } else {
    ## step 1
    dat <- stats::na.omit(dat)
    }
  if (nrow(dat) == 0L) warning("no complete cases")
  ## step 2
  var_mode <- sapply(dat, mode)
  if (any(var_mode %in% c("complex", "raw"))) stop("complex or raw not allowed!")
  var_class <- sapply(dat, class)
  if (any(var_mode[var_class == "AsIs"] %in% c("logical", "character"))) {
    stop("matrix variables with 'AsIs' class must be 'numeric'")
    }
  ind1 <- which(var_mode %in% c("logical", "character"))
  dat[ind1] <- lapply(dat[ind1], as.factor)
  ## step 3
  fctr <- which(sapply(dat, is.factor))
  if (length(fctr) == 0L) warning("no factor variables to summary")
  ind2 <- if (length(ind1) > 0L) fctr[-ind1] else fctr
  dat[ind2] <- lapply(dat[ind2], base::droplevels.factor)
  ## step 4
  lev <- lapply(dat[fctr], base::levels.default)
  nl <- lengths(lev)
  ## return
  list(nlevels = nl, levels = lev)
  }

```

#Variables to put into model

* Dose (counts)
* Dose (mass)
* Dose (volume)
* Size (continuous)
* Size (binned)
* Polymer type
* Shape
* Organism Group (this is being used as a temporary substitute for body size)
* Life Stage
* Level of Biological Organization
* Exposure Duration
* Acute/Chronic (this values are only complete for fish, molluscs, crustaceans and algae)
* Charge (negative/positive - categorical)
* Zeta potential 
* Particle Source (commercial, generated in-house, or mmodified commercial particles e.g., milling)

Response Variables
* Effect (Y/N) - Binary Data
* Effect (Y/N) - Binary Data x Effect Score (binned organism level effects)

#Continous Variable Distributions

There are several continuous variables that we want to feed into our model. Before doing so, we're going to check the distribution to see if any of them are skewed and need to be transformed.

```{r, echo = FALSE}

#Dose - Count

Dose_Particles_mL <- aoc_setup$dose.particles.mL.master

hist(Dose_Particles_mL)

hist(log10(Dose_Particles_mL))

#Dose - Mass

Dose_Mass_L <- aoc_setup$dose.mg.L.master

hist(Dose_Mass_L)

hist(log10(Dose_Mass_L))

#Dose - Volume

`Dose_um^3_mL` <- aoc_setup$dose.um3.mL.master

hist(`Dose_um^3_mL`)

hist(log10(`Dose_um^3_mL`))

#Size - Continuous

Size <- aoc_setup$size.length.um.used.for.conversions

hist(Size)

hist(log10(Size))

#Exposure Duration

Exposure_Duration <- aoc_setup$exposure.duration.d

hist(Exposure_Duration)

hist(log10(Exposure_Duration))

#Zeta Potential

Zeta_Potential_mV <- aoc_setup$zetapotential.mV

hist(Zeta_Potential_mV)

#No need to log transform

aoc_setup_select <- aoc_setup %>% 
  mutate(log.dose.particles.mL.master = log10(dose.particles.mL.master)) %>% 
  mutate(log.dose.mg.L.master = log10(dose.mg.L.master)) %>%
  mutate(log.dose.um3.mL.master = log10(dose.um3.mL.master)) %>%
  mutate(log.size.length.um.used.for.conversions = log10(size.length.um.used.for.conversions)) %>%
  mutate(log.exposure.duration.d = log10(exposure.duration.d))

```

Due to skewed data, the following categories are log10 transformed before modeling:

* Dose (counts)
* Dose (mass)
* Dose (volume)
* Size (continuous)
* Exposure Duration

# Model: Kitchen Sink

All Independent Variables, Response Variable: Effect (Y/N)

Full Model

```{r, echo = FALSE}

#Select the columns that we want to feed into the model for simplicity
aoc_setup_select_1 <- aoc_setup_select %>%
  filter(life_f != "Not Reported") %>% 
  filter(acute.chronic_f != "Unavailable") %>% 
  filter(size_f != "Not Reported") %>% 
  filter(poly_f != "Not Reported") %>%
  filter(shape_f != "Not Reported") %>%
  filter(environment != "Terrestrial") %>%
  dplyr::select(
    log.dose.particles.mL.master, 
    log.dose.mg.L.master,
    log.dose.um3.mL.master,
    log.size.length.um.used.for.conversions,
    # size_f,
    # poly_f,
    shape_f,
    org_f,
    life_f,
    bio_f,
    log.exposure.duration.d,
    acute.chronic_f,
    # charge_f,
    # zetapotential.mV,
    # particle_source_f,
    effect_10)

#Check for single level factors
debug_contr_error(aoc_setup_select_1)

#Omit missing data before modeling
aoc_setup_select_1 <- na.omit(aoc_setup_select_1)

#Fit the full model
Full_Model <- lm(effect_10 ~., data = aoc_setup_select_1)

summary(Full_Model)
```

Stepwise Model - Both Directions

```{r, echo = FALSE}

#Stepwise model
step.model <- stepAIC(Full_Model, direction = "both", 
                      trace = FALSE)

summary(step.model)

```

# Model: Crustacean Fitness, Organism Level Endpoints Only
## Dose (mass)

Full Model

```{r, echo = FALSE}

#Select the columns that we want to feed into the model for simplicity
m1_crust <- aoc_setup %>% 
  filter(!effect_f == "NA") %>% 
  #filter(effect.metric == c("HONEC", "LOEC", "NOEC")) %>% 
  #filter(organism.group == "Crustacea") %>% 
  filter(!acute.chronic_f == "Unavailable") %>% 
  filter(bio.org == "organism") %>% 
  filter(!environment == "Terrestrial") %>% 
  filter(lvl1_f == "Fitness") %>% 
  filter(!size_f == "Not Reported") %>% 
  mutate(logdose.mg.L.master = log10(dose.mg.L.master)) %>%
  mutate(logdose.particles.mL.master = log10(dose.particles.mL.master)) %>%
  mutate(logdose.um3.mL.master = log10(dose.um3.mL.master)) %>%
 # filter(acute.chronic_f == "Acute") %>% 
  dplyr::select(c(effect_10, logdose.mg.L.master, logdose.particles.mL.master, logdose.um3.mL.master, size.length.um.used.for.conversions)) %>% 
  drop_na()

#Fit the full model
m1_crust_model_dose <- glm(effect_10 ~ logdose.mg.L.master*size.length.um.used.for.conversions, data = m1_crust, na.action = "na.exclude", family = "binomial")

summary(m1_crust_model_dose)
```

```{r}
#Plots

ggPredict(m1_crust_model_dose,colorn=100,jitter=FALSE)
```

## Dose (volume)

Full Model

```{r, echo = FALSE}

#Select the columns that we want to feed into the model for simplicity
m1_crust <- aoc_setup %>% 
  filter(!effect_f == "NA") %>% 
  #filter(effect.metric == c("HONEC", "LOEC", "NOEC")) %>% 
  #filter(organism.group == "Crustacea") %>% 
  filter(!acute.chronic_f == "Unavailable") %>% 
  filter(bio.org == "organism") %>% 
  filter(!environment == "Terrestrial") %>% 
  filter(lvl1_f == "Fitness") %>% 
  filter(!size_f == "Not Reported") %>% 
  mutate(logdose.mg.L.master = log10(dose.mg.L.master)) %>%
  mutate(logdose.particles.mL.master = log10(dose.particles.mL.master)) %>%
  mutate(logdose.um3.mL.master = log10(dose.um3.mL.master)) %>%
 # filter(acute.chronic_f == "Acute") %>% 
  dplyr::select(c(effect_10, logdose.mg.L.master, logdose.particles.mL.master, logdose.um3.mL.master, size.length.um.used.for.conversions)) %>% 
  drop_na()

#Fit the full model
m1_crust_model_dose <- glm(effect_10 ~ logdose.um3.mL.master*size.length.um.used.for.conversions, data = m1_crust, na.action = "na.exclude", family = "binomial")

summary(m1_crust_model_dose)
```

```{r}
#Plots
ggPredict(m1_crust_model_dose,
          colorn=200,
 #         jitter=TRUE, 
point = FALSE,
          interactive = FALSE, 
show.summary = TRUE,
#digits = 3, 
se = FALSE)
```

```{r}
#particle couunt
#Select the columns that we want to feed into the model for simplicity
m1_crust_part <- aoc_setup %>% 
  filter(!effect_f == "NA") %>% 
  #filter(effect.metric == c("HONEC", "LOEC", "NOEC")) %>% 
  #filter(organism.group == "Crustacea") %>% 
  filter(!acute.chronic_f == "Unavailable") %>% 
  filter(bio.org == "organism") %>% 
  filter(!environment == "Terrestrial") %>% 
  filter(lvl1_f == "Fitness") %>% 
  filter(!size_f == "Not Reported") %>% 
  mutate(logdose.particles.mL.master = log10(dose.particles.mL.master)) %>% 
 # filter(acute.chronic_f == "Acute") %>% 
  dplyr::select(c(effect_10, logdose.particles.mL.master, size.length.um.used.for.conversions)) %>% 
  drop_na()

#Fit the full model
m1_crust_model_particles <- glm(effect_10 ~ logdose.particles.mL.master * size.length.um.used.for.conversions, data = m1_crust_part, na.action = "na.exclude", family = "binomial")

summary(m1_crust_model_particles)
```
plot
```{r}
ggPredict(m1_crust_model_particles, colorn=1000)
```
### volume
```{r}
#particle couunt
#Select the columns that we want to feed into the model for simplicity
m1_crust_volume <- aoc_setup %>% 
  filter(!effect_f == "NA") %>% 
  #filter(effect.metric == c("HONEC", "LOEC", "NOEC")) %>% 
  #filter(organism.group == "Crustacea") %>% 
  filter(!acute.chronic_f == "Unavailable") %>% 
  filter(bio.org == "organism") %>% 
  filter(!environment == "Terrestrial") %>% 
  filter(lvl1_f == "Fitness") %>% 
  filter(!size_f == "Not Reported") %>% 
  mutate(logdose.um3.mL.master = log10(dose.um3.mL.master)) %>% 
 # filter(acute.chronic_f == "Acute") %>% 
  dplyr::select(c(effect_10, logdose.um3.mL.master, size.length.um.used.for.conversions)) %>% 
  drop_na()

#Fit the full model
m1_crust_model_volume <- glm(effect_10 ~ logdose.um3.mL.master * size.length.um.used.for.conversions, data = m1_crust_volume, na.action = "na.exclude", family = "binomial")

summary(m1_crust_model_volume)
```
plot
```{r}
ggPredict(m1_crust_model_volume, colorn=500)
```

```{r}
#alternative method using ggplot

#get probabilities
m1 <- m1_crust %>% 
  mutate(prob = predict(m1_crust_model_dose,
                        type = "response"))

# plot
ggplot(m1, aes(x = logdose.mg.L.master, y = effect_10, color = size.length.um.used.for.conversions))+
  geom_jitter() +
  geom_smooth(method = 'glm',
              aes(x = logdose.mg.L.master, y = prob), #use predicted NOEC probability
              method.args = list(family = binomial(link = 'logit')),
              se = FALSE, color = 'red') #+
  # geom_smooth(method = 'glm',
  #             method.args = list(family = binomial(link = 'probit')),
  #             se = FALSE, color = 'green') +
  # stat_smooth(method = "glm", method.args = list(family = "binomial"), se = FALSE)
```

```{r}
ggPredict(m1_crust_model_dose,colorn=100,jitter=FALSE)
```

## Dose (count)

Full Model

```{r, echo = FALSE}

#Select the columns that we want to feed into the model for simplicity
m1_crust <- aoc_setup %>% 
  filter(!effect_f == "NA") %>% 
  #filter(effect.metric == c("HONEC", "LOEC", "NOEC")) %>% 
  #filter(organism.group == "Crustacea") %>% 
  filter(!acute.chronic_f == "Unavailable") %>% 
  filter(bio.org == "organism") %>% 
  filter(!environment == "Terrestrial") %>% 
  filter(lvl1_f == "Fitness") %>% 
  filter(!size_f == "Not Reported") %>% 
  mutate(logdose.mg.L.master = log10(dose.mg.L.master)) %>%
  mutate(logdose.particles.mL.master = log10(dose.particles.mL.master)) %>%
  mutate(logdose.um3.mL.master = log10(dose.um3.mL.master)) %>%
 # filter(acute.chronic_f == "Acute") %>% 
  dplyr::select(c(effect_10, logdose.mg.L.master, logdose.particles.mL.master, logdose.um3.mL.master, size.length.um.used.for.conversions)) %>% 
  drop_na()

#Fit the full model
m1_crust_model_dose <- glm(effect_10 ~ logdose.particles.mL.master*size.length.um.used.for.conversions, data = m1_crust, na.action = "na.exclude", family = "binomial")

summary(m1_crust_model_dose)

ggPredict(m1_crust_model_dose,colorn=100,jitter=TRUE, interactive = FALSE, digits = 3, se = FALSE)

```

```{r}
#Plots
ggPredict(m1_crust_model_dose,colorn=100,jitter=FALSE)
```

```{r}
#alternative method using ggpredict
require(ggeffects)
require(ggdark)
require(ggsci)
ggpredict(m1_crust_model_dose)
df <- ggpredict(m1_crust_model_dose, se = TRUE)
df <- ggpredict(m1_crust_model_dose,terms=c("logdose.mg.L.master", "size.length.um.used.for.conversions"))

#note that faceting works best with discrete variable as interactive term
ggplot(df, aes(x = x, y = predicted, colour = group)) +
  geom_line() +
  geom_ribbon( aes(ymin = conf.low, ymax = conf.high, fill = group, color = NULL), alpha = .15) +
  facet_wrap(~.) +
  dark_theme_bw() +
  scale_color_tron()
```

```{r}
#alt method
plot(df, facet = TRUE)
```

## Discrete predictor and dose
```{r, echo = FALSE}
#Select the columns that we want to feed into the model for simplicity
m1_crust <- aoc_setup %>% 
  filter(!effect_f == "NA") %>% 
  #filter(effect.metric == c("HONEC", "LOEC", "NOEC")) %>% 
  #filter(organism.group == "Crustacea") %>% 
  filter(!acute.chronic_f == "Unavailable") %>% 
  filter(bio.org == "organism") %>% 
  filter(!environment == "Terrestrial") %>% 
  filter(lvl1_f == "Fitness") %>% 
  filter(!size_f == "Not Reported") %>% 
  mutate(logdose.mg.L.master = log10(dose.mg.L.master)) %>% 
 # filter(acute.chronic_f == "Acute") %>% 
  dplyr::select(c(effect_10, logdose.mg.L.master, size_f)) %>% 
  drop_na()

#Fit the full model
m1_crust_model_discrete <- glm(effect_10 ~ logdose.mg.L.master * size_f, data = m1_crust, na.action = "na.exclude", family = "binomial")

summary(m1_crust_model_discrete)
```
Plot.
```{r}
discrete <- ggpredict(m1_crust_model_discrete,terms = c("logdose.mg.L.master", "size_f"))
plot(discrete, facet = TRUE, add.data = TRUE)
```
### Survival package
```{r}
#include time
m1_crust <- aoc_setup %>% 
  filter(!effect_f == "NA") %>% 
  #filter(effect.metric == c("HONEC", "LOEC", "NOEC")) %>% 
  #filter(organism.group == "Crustacea") %>% 
  filter(!acute.chronic_f == "Unavailable") %>% 
  filter(bio.org == "organism") %>% 
  filter(!environment == "Terrestrial") %>% 
  filter(lvl1_f == "Fitness") %>% 
  filter(!size_f == "Not Reported") %>% 
  mutate(logdose.mg.L.master = log10(dose.mg.L.master)) %>% 
 # filter(acute.chronic_f == "Acute") %>% 
  dplyr::select(c(effect_10, logdose.mg.L.master, size_f, exposure.duration.d)) %>% 
  drop_na()

require(survival)
survival <- survival::coxph(survival::Surv(exposure.duration.d, effect_10) ~ logdose.mg.L.master + size_f, data = m1_crust)
#cumulative hazard
cumHaz <- ggpredict(survival,terms = c("logdose.mg.L.master", "size_f"), type = "cumhaz")
plot(cumHaz, facet = TRUE, add.data = TRUE, colors = "flat")
```
```{r}
#survival probability over time
pr <- ggpredict(survival, c("logdose.mg.L.master", "size_f"), type = "surv")
plot(pr, colors = "social")
```