### 2021-02-19 DALEX.Rmd
many more models!

---
title: "DALEX"
author: "Scott Coffin"
date: "2/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
https://uc-r.github.io/dalex#:~:text=DALEX%20is%20an%20R%20package,performance%20across%20multiple%20models%20convenient.
```{r}
# load required packages
library(rsample)
library(dplyr)
library(DALEX)
library(tidyverse)
library(tigerstats)
library(caret)
```
# Data Import
```{r}
# Load finalized dataset.
aoc <- read_csv("AquaticOrganisms_Clean_final.csv", guess_max = 10000) #%>% 
  #mutate(ID = paste0("ID",row_number()))

#### Introduction Setup ####

# All text inputs below.

#### Overview AO Setup ####

#Final_effect_dataset <- read_csv("Final_effect_dataset.csv")%>%
  #mutate(plot_f = case_when(
    #plot_f == "Polymer" ~ "Polymer",
    #plot_f == "Size" ~ "Size",
    #plot_f == "Shape" ~ "Shape",
    #plot_f == "Organism" ~ "Organism",
    #plot_f == "Lvl1" ~ "Endpoint Category",
    #plot_f == "Life.stage" ~ "Life Stage",
    #plot_f == "Invivo.invivo" ~ "In Vivo or In Vitro",
    #plot_f == "Exposure.route" ~ "Exposure Route"))%>%
  #mutate(plot_f = factor(plot_f))%>%
  #mutate(logEndpoints = log(Endpoints))%>%
  #rename(Percent = Freq)

polydf<-rowPerc(xtabs( ~polymer +effect, aoc)) #pulls polymers by effect 
polyf<-as.data.frame(polydf)%>% #Makes data frame 
  filter(effect %in% c("Y","N"))%>% #Sorts into Yes and No
  mutate(polymer = case_when(
    polymer == "BIO" ~ "Biopolymer",
    polymer == "EVA" ~ "Polyethylene Vinyl Acetate",
    polymer == "LTX" ~ "Latex",
    polymer == "PA" ~ "Polyamide",
    polymer == "PE" ~ "Polyethylene",
    polymer == "PC" ~ "Polycarbonate",
    polymer == "PET" ~ "Polyethylene Terephthalate",
    polymer == "PI" ~ "Polyisoprene",
    polymer == "PMMA" ~ "Polymethylmethacrylate",
    polymer == "PP" ~ "Polypropylene",
    polymer == "PS" ~ "Polystyrene",
    polymer == "PUR" ~ "Polyurathane",
    polymer == "PVC" ~ "Polyvinylchloride",
    polymer == "PLA" ~ "Polylactic Acid"))%>%
  mutate_if(is.numeric, round,0) #rounds percents 
Endpoints<-xtabs(~polymer +effect ,aoc) #Pulls all study obs. for polymer from dataset
polyfinal<- data.frame(cbind(polyf, Endpoints))%>% #adds it as a column
  rename(Endpoints='Freq.1')%>% #renames column
  mutate(logEndpoints = log(Endpoints))%>%
  rename(Percent = Freq)#renames column

sizedf<-rowPerc(xtabs(~size.category +effect, aoc))
sizef<-as.data.frame(sizedf)%>%
  filter(effect %in% c("Y","N"))%>%
  mutate(size.category = case_when(
    size.category == 1 ~ "<1µm",
    size.category == 2 ~ "1µm < 10µm",
    size.category == 3 ~ "10µm < 100µm",
    size.category == 4 ~ "100µm < 1mm",
    size.category == 5 ~ "1mm < 5mm",
    size.category == 0 ~ "unavailable"))%>%
  rename(Type = "size.category")%>%
  mutate_if(is.numeric, round,0)%>%
  mutate(plot="Size")
study_s<-xtabs(~size.category +effect ,aoc)
sizefinal<- data.frame(cbind(sizef, study_s))%>% 
  rename(Endpoints='Freq.1')%>%
  rename(category='size.category')%>%
  mutate(logEndpoints = log(Endpoints))%>%
  rename(Percent = Freq)#renames column

shapedf<-rowPerc(xtabs(~shape + effect, aoc))
shapef<-as.data.frame(shapedf)%>%
  filter(effect %in% c("Y","N"))%>%
  rename(Type="shape")%>%
  mutate_if(is.numeric, round,0)%>%
  mutate(plot="Shape")%>%
  mutate(Type = case_when(
    Type == "cube" ~ "Cube",
    Type == "sphere" ~ "Sphere",
    Type == "fragment" ~ "Fragment",
    Type == "fiber" ~ "Fiber"))
study_sh<-xtabs(~shape + effect,aoc)
shapefinal<- data.frame(cbind(shapef, study_sh))%>% 
  rename(Endpoints='Freq.1')%>%
  rename(category='shape')%>%
  mutate(logEndpoints = log(Endpoints))%>%
  rename(Percent = Freq)#renames column

taxdf<-rowPerc(xtabs(~organism.group +effect, aoc))
taxf<-as.data.frame(taxdf)%>%
  filter(effect %in% c("Y","N"))%>%
  rename(Type= "organism.group")%>%
  mutate_if(is.numeric, round,0)%>%
  mutate(plot="Organism")
study_t<-xtabs(~organism.group +effect,aoc)
taxfinal<- data.frame(cbind(taxf, study_t))%>% 
  rename(Endpoints='Freq.1')%>%
  rename(category='organism.group')%>%
  mutate(logEndpoints = log(Endpoints))%>%
  rename(Percent = Freq)#renames column

lvl1df<-rowPerc(xtabs(~lvl1 +effect, aoc))
lvl1f<-as.data.frame(lvl1df)%>%
  filter(effect %in% c("Y","N"))%>%
  rename(Type= "lvl1")%>%
  mutate_if(is.numeric, round,0)%>%
  mutate(plot="Lvl1")%>%
  mutate(Type = case_when(
    Type == "alimentary.excretory" ~ "Alimentary, Excretory",
    Type == "behavioral.sense.neuro" ~ "Behavioral, Sensory, Neurological",
    Type == "circulatory.respiratory" ~ "Circulatory, Respiratory",
    Type == "community" ~ "Community",
    Type == "fitness" ~ "Fitness",
    Type == "immune" ~ "Immune",
    Type == "metabolism" ~ "Metabolism",
    Type == "microbiome" ~ "Microbiome",
    Type == "stress" ~ "Stress")) 
study_l<-xtabs(~lvl1 +effect,aoc)
lvl1final<- data.frame(cbind(lvl1f, study_l))%>% 
  rename(Endpoints='Freq.1')%>%
  rename(category='lvl1')%>%
  mutate(logEndpoints = log(Endpoints))%>%
  rename(Percent = Freq)#renames column

lifedf<-rowPerc(xtabs(~life.stage +effect, aoc))
lifef<-as.data.frame(lifedf)%>%
  filter(effect %in% c("Y","N"))%>%
  rename(Type= "life.stage")%>%
  mutate_if(is.numeric, round,0)%>%
  mutate(plot="Life.stage")
studyli<-xtabs(~life.stage +effect ,aoc)
lifefinal<- data.frame(cbind(lifef, studyli))%>% 
  rename(Endpoints='Freq.1')%>%
  rename(category='life.stage')%>%
  mutate(logEndpoints = log(Endpoints))%>%
  rename(Percent = Freq)#renames column

vivodf<-rowPerc(xtabs(~invitro.invivo +effect, aoc))
vivof<-as.data.frame(vivodf)%>%
  filter(effect %in% c("Y","N"))%>%
  rename(Type= "invitro.invivo")%>%
  mutate_if(is.numeric, round,0)%>%
  mutate(plot="Invivo.invivo")%>%
  mutate(Type = case_when(
    Type=="invivo"~"In Vivo",
    Type=="invitro"~"In Vitro"))
study_v<-xtabs(~invitro.invivo +effect,aoc)
vivofinal<- data.frame(cbind(vivof, study_v))%>% 
  rename(Endpoints='Freq.1')%>%
  rename(category='invitro.invivo')%>%
  mutate(logEndpoints = log(Endpoints))%>%
  rename(Percent = Freq)#renames column

routedf<-rowPerc(xtabs(~exposure.route +effect, aoc))
routef<-as.data.frame(routedf)%>%
  filter(effect %in% c("Y","N"))%>%
  rename(Type= "exposure.route")%>%
  mutate_if(is.numeric, round,0)%>%
  mutate(plot="Exposure.route")%>%
  mutate(Type = case_when(
    Type == "coparental.exposure" ~"Co-Parental Exposure",
    Type == "paternal.exposure" ~ "Paternal Exposure",
    Type == "maternal.exposure" ~ "Maternal Exposure",
    Type == "food" ~ "Food",
    Type == "water" ~ "Water",
    Type == "sediment" ~ "Sediment",
    Type == "media" ~ "Media"))
study_r<-xtabs(~exposure.route +effect,aoc)
routefinal<- data.frame(cbind(routef, study_r))%>% 
  rename(Endpoints='Freq.1')%>%
  rename(category='exposure.route')%>%
  mutate(logEndpoints = log(Endpoints))%>%
  rename(Percent = Freq)#renames column

#### Exploration AO Setup ####

# Master dataset for scatterplots - for Heili's tab.
aoc_v1 <- aoc %>% # start with original dataset
   # full dataset filters.
  mutate(effect_f = factor(case_when(effect == "Y" ~ "Yes",
    effect == "N" ~ "No"),
    levels = c("No", "Yes"))) %>%
  # removing NAs to make data set nicer
  replace_na(list(size.category = 0, shape = "Not Reported", polymer = "Not Reported", life.stage = "Not Reported"))

aoc_setup <- aoc_v1 %>% # start with original dataset
  mutate(size_f = factor(case_when(
    size.category == 1 ~ "1nm < 100nm",
    size.category == 2 ~ "100nm < 1µm",
    size.category == 3 ~ "1µm < 100µm",
    size.category == 4 ~ "100µm < 1mm",
    size.category == 5 ~ "1mm < 5mm",
    size.category == 0 ~ "Not Reported"),
    levels = c("1nm < 100nm", "100nm < 1µm", "1µm < 100µm", "100µm < 1mm", "1mm < 5mm", "Not Reported"))) %>% # creates new column with nicer names and order by size levels.
  # shape category data tidying.
  mutate(shape_f = factor(case_when(
    shape == "fiber" ~ "Fiber",
    shape == "fragment" ~ "Fragment",
    shape == "sphere" ~ "Sphere",
    shape == "Not Reported" ~ "Not Reported"),
    levels = c("Fiber", "Fragment", "Sphere", "Not Reported"))) %>% # order our different shapes.
  # polymer category data tidying.
  mutate(poly_f = factor(case_when(
    polymer == "BIO" ~ "Biopolymer",
    polymer == "EVA" ~ "Polyethylene Vinyl Acetate",
    polymer == "LTX" ~ "Latex",
    polymer == "PA" ~ "Polyamide",
    polymer == "PE" ~ "Polyethylene",
    polymer == "PC" ~ "Polycarbonate",
    polymer == "PET" ~ "Polyethylene Terephthalate",
    polymer == "PI" ~ "Polyisoprene",
    polymer == "PMMA" ~ "Polymethylmethacrylate",
    polymer == "PP" ~ "Polypropylene",
    polymer == "PS" ~ "Polystyrene",
    polymer == "PUR" ~ "Polyurathane",
    polymer == "PVC" ~ "Polyvinylchloride",
    polymer == "PLA" ~ "Polylactic Acid",
    polymer == "Not Reported" ~ "Not Reported"))) %>%
  # taxonomic category data tidying.
  mutate(org_f = factor(organism.group, levels = c("Algae", "Annelida", "Bacterium", "Cnidaria", "Crustacea",
                                                   "Echinoderm", "Fish", "Insect", "Mollusca", "Nematoda", "Plant", "Rotifera", "Mixed"))) %>% # order our different organisms.
  mutate(lvl1_f = factor(case_when(lvl1 == "alimentary.excretory" ~ "Alimentary, Excretory",
    lvl1 == "behavioral.sense.neuro" ~ "Behavioral, Sensory, Neurological",
    lvl1 == "circulatory.respiratory" ~ "Circulatory, Respiratory",
    lvl1 == "community" ~ "Community",
    lvl1 == "fitness" ~ "Fitness",
    lvl1 == "immune" ~ "Immune",
    lvl1 == "metabolism" ~ "Metabolism",
    lvl1 == "microbiome" ~ "Microbiome",
    lvl1 == "stress" ~ "Stress"))) %>% # creates new column with nicer names.
  # Level 2 Data tidying
  mutate(lvl2_f = factor(case_when(lvl2 == "abundance"~"Abundance",
    lvl2 == "actinobacteria" ~ "Actinobacteria",
    lvl2 == "aggressivity"~"Agressivity",
    lvl2 == "ammonia.excretion" ~ "Ammonia Excretion",
    lvl2 == "bacteroidetes"~ "Bacteriodetes",
    lvl2 == "blood"~"Blood",
    lvl2 == "body.condition"~"Body Condition",
    lvl2 == "boldness"~"Boldness",
    lvl2 == "brain.histo"~"Brain Histological Abnormalities",
    lvl2 == "burrowing"~"Burrowing",
    lvl2 == "carb.metabolism"~"Carb Metabolism",
    lvl2 == "chemokines.cytokines"~"Chemokines",
    lvl2 == "circulatory"~"Circulatory",
    lvl2 == "detoxification"~"Detoxification",
    lvl2 == "development"~"Development",
    lvl2 == "digestion"~"Digestion",
    lvl2 == "digestive.enzymes"~"Digestive Enzymes",
    lvl2 == "digestive.tract.histo"~"Digestive Tract Histological Abnormalities",
    lvl2 == "diversity"~ "Diversity",
    lvl2 == "feeding"~ "Feeding",
    lvl2 == "firmicutes"~ "Firmicutes",
    lvl2 == "gall.bladder.histo" ~ "Gall Bladder Histological Abnormalities",
    lvl2 == "gen.metabolism"~ "General Metabolism",
    lvl2 == "gill.histo"~ "Gill Histological Abnormalities",
    lvl2 == "gonad.histo"~"Gonad Histological Abnormalities",
    lvl2 == "growth"~ "Growth",
    lvl2 == "immune.cells"~"Immune Cells",
    lvl2 == "immune.other"~"Immune Other ",
    lvl2 == "intestinal.permeability"~"Intestinal Permeability",
    lvl2 == "kidney.histo"~"Kidney Histological abnormalities",
    lvl2 == "lipid.metabolism"~"Lipid Metabolism",
    lvl2 == "liver.histo"~"Liver Histological Abnormalities",
    lvl2 == "liver.kidney.products" ~ "Liver and Kidney Products",
    lvl2 == "locomotion"~"Locomotion",
    lvl2 == "mortality"~"Mortality",
    lvl2 == "nervous.system"~"Nervous System",
    lvl2 == "oxidative.stress"~"Oxidative Stress",
    lvl2 == "photosynthesis"~ "Photosynthesis",
    lvl2 == "proteobacteria"~"Protebacteria",
    lvl2 == "reproduction"~"Reproduction",
    lvl2 == "respiration"~"Respiration",
    lvl2 == "sexhormones"~"Sex Hormones",
    lvl2 == "shoaling"~"Shoaling",
    lvl2 == "stress"~"Stress",
    lvl2 == "vision.system"~"Vision System"))) %>% #Renames for widget
  mutate(bio_f = factor(case_when(bio.org == "cell"~"Cell", #Bio Org Data Tidying
    bio.org == "organism"~"Organism",
    bio.org == "population"~ "Population",
    bio.org == "subcell"~"Subcell",
    bio.org == "tissue" ~ "Tissue")))%>%
  mutate(vivo_f = factor(case_when(invitro.invivo == "invivo"~"In Vivo",
    invitro.invivo == "invitro"~"In Vitro")))%>% ##Renames for widget (Not using a widget right now, but saving for human health database)
  mutate(life_f = factor(case_when(life.stage == "Early"~"Early",
    life.stage == "Juvenile"~"Juvenile",
    life.stage == "Adult"~"Adult",
    life.stage == "Not Reported"~"Not Reported")))%>% #Renames for widget
  mutate(env_f = factor(case_when(environment == "Freshwater"~"Freshwater",
    environment == "Marine" ~ "Marine",
    environment == "Terrestrial" ~ "Terrestrial"))) %>%
  mutate(dose.mg.L.master.converted.reported = factor(dose.mg.L.master.converted.reported)) %>%
  mutate(dose.particles.mL.master.converted.reported = factor(dose.particles.mL.master.converted.reported)) %>% 
   mutate(dose.um3.mL.master = particle.volume.um3 * dose.particles.mL.master) %>%   #calculate volume/mL
  mutate(af.time_noNA = replace_na(af.time, "Unavailable")) %>% 
  mutate(acute.chronic_f = factor(case_when(af.time_noNA == 10 ~ "Acute",
                                            af.time_noNA == 1 ~ "Chronic",
                                            af.time_noNA == "Unavailable" ~ "Unavailable"))) %>%    #factorize assesment factor time into chronic/acute
  mutate(dose.mg.L.master.AF.noec = dose.mg.L.master * af.noec) %>% 
  mutate(dose.particles.mL.master.AF.noec = dose.particles.mL.master * af.noec) %>% 
  mutate(effect_f = factor(effect)) %>% 
  mutate_if(is.character, as.factor) %>% 
  mutate(effect_10 = case_when(
     effect_f == "Y" ~ 1,
     effect_f == "N" ~ 0))

#### SSD AO Setup ####

# Master dataset for SSDs
aoc_z <- aoc_setup %>% # start with Heili's altered dataset (no filtration for terrestrial data)
  # environment category data tidying.
  mutate(environment.noNA = replace_na(environment, "Not Reported")) %>% # replaces NA to better relabel.
  mutate(env_f = factor(environment.noNA, levels = c("Marine", "Freshwater", "Terrestrial", "Not Reported")))

# final cleanup and factoring  
aoc_z$Species <- as.factor(paste(aoc_z$genus,aoc_z$species)) #must make value 'Species" (uppercase)
aoc_z$Group <- as.factor(aoc_z$organism.group) #must make value "Group"
aoc_z$Group <- fct_explicit_na(aoc_z$Group) #makes sure that species get counted even if they're missing a group
```
## Data Selection
```{r}
df <- aoc_z %>% 
  dplyr::select(c(organism.group, exposure.duration.d, lvl1_f, dose.um3.mL.master, dose.mg.L.master, dose.particles.mL.master, life.stage, bio.org, polymer, shape, size.length.um.used.for.conversions, treatments, effect)) %>% 
  drop_na() %>%  #drop missing
mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
  mutate(effect_10 = case_when( #convert ordinal to numeric
      effect == "Y" ~ 1,
      effect == "N" ~ 0
    )) %>%
  mutate(effect_10 = factor(effect_10)) %>% 
  dplyr::select(-(effect))

#ensure completeness
skim(df)
```

```{r}
# create train, validation, and test splits
# Create calibration and validation splits with tidymodels initial_split() function.
set.seed(4)
df_split <- df %>%
  initial_split(prop = 0.75)
# default is 3/4ths split (but 75% training, 25% testing).

# Create a training data set with the training() function
# Pulls from training and testing sets created by initial_split()
train <- training(df_split)
test <- testing(df_split)

# variable names for resonse & features
y <- "effect_10"
x <- setdiff(names(df), y) 
```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
# Examine the environment to be sure # of observations looks like the 75/25 split. 3199:1066.
classif_rf <- train(effect_10~., data = train, method = "rf", ntree = 100, tuneLength = 1)

classif_glm <- train(effect_10~., data = train, method = "glm", family = "binomial")

classif_svm <- train(effect_10~., data = train, method = "svmRadial", prob.model = TRUE, tuneLength = 1,
                     preProcess = c("pca","scale","center"))

classif_decTree <- train(effect_10~., data = train, method = "C5.0", preProcess = c("scale","center"))

classif_nnet <- train(effect_10~., data = train, method = "nnet", preProcess=c("scale","center"))

classif_xgbTree <- train(effect_10~., data = train, method = "xgbTree")

#classif_logicBag <- train(effect_10~., data = train, method = "logicBag", ntree = 100)

# explain
yTest <- as.numeric(as.character(test$effect_10))

explainer_classif_rf <- DALEX::explain(classif_rf, label = "rf",
                                       data = test, 
                                       y = yTest)

explainer_classif_glm <- DALEX::explain(classif_glm, label = "glm", 
                                        data = test, y = yTest)

explainer_classif_svm <- DALEX::explain(classif_svm,  label = "svm", 
                                        data = test, y = yTest)

explainer_classif_decTree <- DALEX::explain(classif_rf, label = "Decision Tree",
                                       data = test, 
                                       y = yTest)

explainer_classif_nnet <- DALEX::explain(classif_glm, label = "Neural Net", 
                                        data = test, y = yTest)

explainer_classif_xgbTree <- DALEX::explain(classif_svm,  label = "eXtreme Gradient Boosting Trees", 
                                        data = test, y = yTest)

#explainer_classif_logicBag <- DALEX::explain(classif_svm,  label = "Bagged Logic Regression", 
                                       # data = test, y = yTest)

```
### Model Performance
```{r}
mp_classif_rf <- model_performance(explainer_classif_rf)
mp_classif_glm <- model_performance(explainer_classif_glm)
mp_classif_svm <- model_performance(explainer_classif_svm)
mp_classif_decTree <- model_performance(explainer_classif_decTree)
mp_classif_nnet <- model_performance(explainer_classif_nnet)
mp_classif_xgbTree <- model_performance(explainer_classif_xgbTree)
#mp_classif_logicBag <- model_performance(explainer_classif_logicBag)

# compare residuals plots
resid_dist <- plot(mp_classif_rf, mp_classif_glm, mp_classif_svm, mp_classif_decTree, mp_classif_nnet,
                   mp_classif_xgbTree) + #, mp_classif_logicBag) +
  theme_minimal() +
        theme(legend.position = 'bottom',
              plot.title = element_text(hjust = 0.5)) + 
        labs(y = '')
resid_dist
```
#### Residuals
```{r}
resid_box <- plot(mp_classif_rf, mp_classif_glm, mp_classif_svm, mp_classif_decTree, mp_classif_nnet, mp_classif_xgbTree,
     #mp_classif_logicBag, 
     geom = "boxplot") +
  theme_minimal() +
        theme(legend.position = 'bottom',
              plot.title = element_text(hjust = 0.5)) 
resid_box
```
```{r}
require(gridExtra)
grid.arrange(resid_box,resid_dist, ncol=2)
```
#### Roc Curves
```{r}
plot(mp_classif_rf, 
     #mp_classif_glm, 
     mp_classif_svm, 
     #mp_classif_decTree, 
     mp_classif_nnet,
     #mp_classif_xgbTree,
     geom = "roc") +
  ggtitle("ROC Curves - All Models",  
          paste("AUC_rf = ",round(mp_classif_rf$measures$auc,2), 
                paste("AUC_svm = ",round(mp_classif_svm$measures$auc,2)),
          paste("AUC_nnet = ",round(mp_classif_nnet$measures$auc,2))
          )) +
#,  "AUC_glm = 0.799  AUC_svm = 0.798 AUC_decTree = AUC_nnet = AUC_xgbTree = ") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

### Variable Importance
```{r fig.height=7, fig.width=5}
vi_classif_rf <- model_parts(explainer_classif_rf, loss_function = loss_root_mean_square)
vi_classif_glm <- model_parts(explainer_classif_glm, loss_function = loss_root_mean_square)
vi_classif_svm <- model_parts(explainer_classif_svm, loss_function = loss_root_mean_square)
vi_classif_decTree <- model_parts(explainer_classif_decTree, loss_function = loss_root_mean_square)
vi_classif_nnet <- model_parts(explainer_classif_nnet, loss_function = loss_root_mean_square)
vi_classif_xgbTree <- model_parts(explainer_classif_xgbTree, loss_function = loss_root_mean_square)
#vi_classif_logicBag <- model_parts(explainer_classif_logicBag, loss_function = loss_root_mean_square)

plot(vi_classif_rf, vi_classif_glm, vi_classif_svm, vi_classif_decTree, vi_classif_nnet, vi_classif_xgbTree)#, #vi_classif_logicBag)
```
#### Partial Dependence Plot
```{r}
pdp_classif_rf  <- model_profile(explainer_classif_rf, variable = "dose.um3.mL.master", type = "partial")
pdp_classif_glm  <- model_profile(explainer_classif_glm, variable = "dose.um3.mL.master", type = "partial")
pdp_classif_svm  <- model_profile(explainer_classif_svm, variable = "dose.um3.mL.master", type = "partial")
pdp_classif_decTree  <- model_profile(explainer_classif_decTree, variable = "dose.um3.mL.master", type = "partial")
pdp_classif_nnet  <- model_profile(explainer_classif_nnet, variable = "dose.um3.mL.master", type = "partial")
pdp_classif_xgbTree  <- model_profile(explainer_classif_xgbTree, variable = "dose.um3.mL.master", type = "partial")
#pdp_classif_logicBag  <- model_profile(explainer_classif_logicBag, variable = "dose.um3.mL.master", type = "partial")

plot(pdp_classif_rf, pdp_classif_glm, pdp_classif_svm, pdp_classif_decTree, pdp_classif_nnet, pdp_classif_xgbTree)#, pdp_classif_logicBag)
```
```{r}
#Partial Dependence organism type
pdp_classif_rf  <- model_profile(explainer_classif_rf, variable = "organism.group", type = "partial")
pdp_classif_glm  <- model_profile(explainer_classif_glm, variable = "organism.group", type = "partial")
pdp_classif_svm  <- model_profile(explainer_classif_svm, variable = "organism.group", type = "partial")
pdp_classif_decTree  <- model_profile(explainer_classif_decTree, variable = "organism.group", type = "partial")
pdp_classif_nnet  <- model_profile(explainer_classif_nnet, variable = "organism.group", type = "partial")
pdp_classif_xgbTree  <- model_profile(explainer_classif_xgbTree, variable = "organism.group", type = "partial")
#pdp_classif_logicBag  <- model_profile(explainer_classif_logicBag, variable = "dose.um3.mL.master", type = "partial")

plot(pdp_classif_rf$agr_profiles, pdp_classif_glm$agr_profiles, pdp_classif_svm$agr_profiles, pdp_classif_decTree$agr_profiles, pdp_classif_nnet$agr_profiles, pdp_classif_xgbTree$agr_profiles) +
  ggtitle("Contrastive Partial Dependence Profiles", "") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

#### Breakdown

```{r}
# create a single observation
new_cust <- test$effect_10 %>% as.data.frame()

# compute breakdown distances
new_cust_glm <- predict_parts(explainer_classif_glm, new_observation = test, type = "break_down")
new_cust_rf <- predict_parts(explainer_classif_rf, new_observation = test, type = "break_down")
new_cust_svm <- predict_parts(explainer_classif_svm, new_observation = test, type = "break_down")
new_cust_decTree <- predict_parts(explainer_classif_decTree, new_observation = test, type = "break_down")
new_cust_nnet <- predict_parts(explainer_classif_nnet, new_observation = test, type = "break_down")
new_cust_xgbTree <- predict_parts(explainer_classif_xgbTree, new_observation = test, type = "break_down")
#new_cust_logicBag <- predict_parts(explainer_classif_logicBag, new_observation = test, type = "break_down")

# class of prediction_breakdown output
class(new_cust_rf)

# check out the top 10 influential variables for this observation
new_cust_rf[1:10, 1:5]
```

```{r}
plot(new_cust_glm, new_cust_rf, new_cust_svm, new_cust_decTree, new_cust_nnet, new_cust_xgbTree)#,
     #new_cust_logicBag)
```
```{r}
library(ggplot2)

# filter for top 10 influential variables for each model and plot
list(new_cust_glm, new_cust_rf, new_cust_svm, new_cust_decTree, new_cust_nnet, 
     #new_cust_logicBag,
     new_cust_xgbTree) %>%
  purrr::map(~ top_n(., 11, wt = abs(contribution))) %>%
  do.call(rbind, .) %>%
  mutate(variable = paste0(variable, " (", label, ")")) %>%
  ggplot(aes(contribution, reorder(variable, contribution))) +
  geom_point() +
  geom_vline(xintercept = 0, size = 3, color = "white") +
  facet_wrap(~ label, scales = "free_y", ncol = 1) +
  ylab(NULL)
```
# Final Model
All in all random forest is my final model of choice: it appears the more balanced and is the most accurate overall. I will further refine this model using recursive feature elimination and compare accuracy with the other models.

```{r}

my_ctrl <- rfeControl(functions = rfFuncs,
                      method = "repeatedcv",
                      verbose = FALSE,
                      repeats = 5,
                      returnResamp = "all")

rfProfile <- rfe(y = df$effect_10, # set dependent variable
              x = df %>% 
                dplyr::select(-effect_10),
              rfeControl = my_ctrl,
               size = c(1:2, 4, 6, 8, 10, 12, 13))
rfProfile
```
```{r}
y <- "effect_10"
x <- setdiff(names(df), y)

#alt method
rfRFE <-  list(summary = defaultSummary,
               fit = function(x, y, first, last, ...){
                 library(randomForest)
                 randomForest(x, y, importance = first, ...)
                 },
               pred = function(object, x)  predict(object, x),
               rank = function(object, x, y) {
                 vimp <- varImp(object)
                 vimp <- vimp[order(vimp$Overall,decreasing = TRUE),,drop = FALSE]
                 vimp$var <- rownames(vimp)                  
                 vimp
                 },
               selectSize = pickSizeBest,
               selectVar = pickVars)
```

```{r}
predictors(rfProfile)
```

```{r}
head(rfProfile$resample)
```

```{r}
trellis.par.set(caretTheme())
plot(rfProfile, type = c("g", "o"))
```

```{r}
sessionInfo()
```