### 2021-02-24 DALEX.Rmd
more plots!

---
title: "DALEX_TechQualityControl"
author: "Scott Coffin"
date: "2/23/2021"
output: 
  html_document:
    code_folding: hide
    theme: sandstone
    toc: yes
    toc_float: yes
    toc_depth: 6
    number_sections: true
  word_document:
    toc: yes
---
```{r global-options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      warning=FALSE, message=FALSE, echo = FALSE)
```

```{r message=FALSE, warning=FALSE}
# load required packages
library(rsample)
library(dplyr)
library(DALEX)
library(tidyverse)
library(tigerstats)
library(caret)
library(skimr)
library(ggeffects)
library(knitr)
library(doParallel)
library(modelplotr)
library(gridExtra)
```
# Setup
## Data Import
```{r}
# Load finalized dataset.
aoc <- read_csv("AquaticOrganisms_Clean_final.csv", guess_max = 10000) #%>% 
  #mutate(ID = paste0("ID",row_number()))

#### Introduction Setup ####

# All text inputs below.

#### Overview AO Setup ####

#Final_effect_dataset <- read_csv("Final_effect_dataset.csv")%>%
  #mutate(plot_f = case_when(
    #plot_f == "Polymer" ~ "Polymer",
    #plot_f == "Size" ~ "Size",
    #plot_f == "Shape" ~ "Shape",
    #plot_f == "Organism" ~ "Organism",
    #plot_f == "Lvl1" ~ "Endpoint Category",
    #plot_f == "Life.stage" ~ "Life Stage",
    #plot_f == "Invivo.invivo" ~ "In Vivo or In Vitro",
    #plot_f == "Exposure.route" ~ "Exposure Route"))%>%
  #mutate(plot_f = factor(plot_f))%>%
  #mutate(logEndpoints = log(Endpoints))%>%
  #rename(Percent = Freq)

polydf<-rowPerc(xtabs( ~polymer +effect, aoc)) #pulls polymers by effect 
polyf<-as.data.frame(polydf)%>% #Makes data frame 
  filter(effect %in% c("Y","N"))%>% #Sorts into Yes and No
  mutate(polymer = case_when(
    polymer == "BIO" ~ "Biopolymer",
    polymer == "EVA" ~ "Polyethylene Vinyl Acetate",
    polymer == "LTX" ~ "Latex",
    polymer == "PA" ~ "Polyamide",
    polymer == "PE" ~ "Polyethylene",
    polymer == "PC" ~ "Polycarbonate",
    polymer == "PET" ~ "Polyethylene Terephthalate",
    polymer == "PI" ~ "Polyisoprene",
    polymer == "PMMA" ~ "Polymethylmethacrylate",
    polymer == "PP" ~ "Polypropylene",
    polymer == "PS" ~ "Polystyrene",
    polymer == "PUR" ~ "Polyurathane",
    polymer == "PVC" ~ "Polyvinylchloride",
    polymer == "PLA" ~ "Polylactic Acid"))%>%
  mutate_if(is.numeric, round,0) #rounds percents 
Endpoints<-xtabs(~polymer +effect ,aoc) #Pulls all study obs. for polymer from dataset
polyfinal<- data.frame(cbind(polyf, Endpoints))%>% #adds it as a column
  rename(Endpoints='Freq.1')%>% #renames column
  mutate(logEndpoints = log(Endpoints))%>%
  rename(Percent = Freq)#renames column

sizedf<-rowPerc(xtabs(~size.category +effect, aoc))
sizef<-as.data.frame(sizedf)%>%
  filter(effect %in% c("Y","N"))%>%
  mutate(size.category = case_when(
    size.category == 1 ~ "<1µm",
    size.category == 2 ~ "1µm < 10µm",
    size.category == 3 ~ "10µm < 100µm",
    size.category == 4 ~ "100µm < 1mm",
    size.category == 5 ~ "1mm < 5mm",
    size.category == 0 ~ "unavailable"))%>%
  rename(Type = "size.category")%>%
  mutate_if(is.numeric, round,0)%>%
  mutate(plot="Size")
study_s<-xtabs(~size.category +effect ,aoc)
sizefinal<- data.frame(cbind(sizef, study_s))%>% 
  rename(Endpoints='Freq.1')%>%
  rename(category='size.category')%>%
  mutate(logEndpoints = log(Endpoints))%>%
  rename(Percent = Freq)#renames column

shapedf<-rowPerc(xtabs(~shape + effect, aoc))
shapef<-as.data.frame(shapedf)%>%
  filter(effect %in% c("Y","N"))%>%
  rename(Type="shape")%>%
  mutate_if(is.numeric, round,0)%>%
  mutate(plot="Shape")%>%
  mutate(Type = case_when(
    Type == "cube" ~ "Cube",
    Type == "sphere" ~ "Sphere",
    Type == "fragment" ~ "Fragment",
    Type == "fiber" ~ "Fiber"))
study_sh<-xtabs(~shape + effect,aoc)
shapefinal<- data.frame(cbind(shapef, study_sh))%>% 
  rename(Endpoints='Freq.1')%>%
  rename(category='shape')%>%
  mutate(logEndpoints = log(Endpoints))%>%
  rename(Percent = Freq)#renames column

taxdf<-rowPerc(xtabs(~organism.group +effect, aoc))
taxf<-as.data.frame(taxdf)%>%
  filter(effect %in% c("Y","N"))%>%
  rename(Type= "organism.group")%>%
  mutate_if(is.numeric, round,0)%>%
  mutate(plot="Organism")
study_t<-xtabs(~organism.group +effect,aoc)
taxfinal<- data.frame(cbind(taxf, study_t))%>% 
  rename(Endpoints='Freq.1')%>%
  rename(category='organism.group')%>%
  mutate(logEndpoints = log(Endpoints))%>%
  rename(Percent = Freq)#renames column

lvl1df<-rowPerc(xtabs(~lvl1 +effect, aoc))
lvl1f<-as.data.frame(lvl1df)%>%
  filter(effect %in% c("Y","N"))%>%
  rename(Type= "lvl1")%>%
  mutate_if(is.numeric, round,0)%>%
  mutate(plot="Lvl1")%>%
  mutate(Type = case_when(
    Type == "alimentary.excretory" ~ "Alimentary, Excretory",
    Type == "behavioral.sense.neuro" ~ "Behavioral, Sensory, Neurological",
    Type == "circulatory.respiratory" ~ "Circulatory, Respiratory",
    Type == "community" ~ "Community",
    Type == "fitness" ~ "Fitness",
    Type == "immune" ~ "Immune",
    Type == "metabolism" ~ "Metabolism",
    Type == "microbiome" ~ "Microbiome",
    Type == "stress" ~ "Stress")) 
study_l<-xtabs(~lvl1 +effect,aoc)
lvl1final<- data.frame(cbind(lvl1f, study_l))%>% 
  rename(Endpoints='Freq.1')%>%
  rename(category='lvl1')%>%
  mutate(logEndpoints = log(Endpoints))%>%
  rename(Percent = Freq)#renames column

lifedf<-rowPerc(xtabs(~life.stage +effect, aoc))
lifef<-as.data.frame(lifedf)%>%
  filter(effect %in% c("Y","N"))%>%
  rename(Type= "life.stage")%>%
  mutate_if(is.numeric, round,0)%>%
  mutate(plot="Life.stage")
studyli<-xtabs(~life.stage +effect ,aoc)
lifefinal<- data.frame(cbind(lifef, studyli))%>% 
  rename(Endpoints='Freq.1')%>%
  rename(category='life.stage')%>%
  mutate(logEndpoints = log(Endpoints))%>%
  rename(Percent = Freq)#renames column

vivodf<-rowPerc(xtabs(~invitro.invivo +effect, aoc))
vivof<-as.data.frame(vivodf)%>%
  filter(effect %in% c("Y","N"))%>%
  rename(Type= "invitro.invivo")%>%
  mutate_if(is.numeric, round,0)%>%
  mutate(plot="Invivo.invivo")%>%
  mutate(Type = case_when(
    Type=="invivo"~"In Vivo",
    Type=="invitro"~"In Vitro"))
study_v<-xtabs(~invitro.invivo +effect,aoc)
vivofinal<- data.frame(cbind(vivof, study_v))%>% 
  rename(Endpoints='Freq.1')%>%
  rename(category='invitro.invivo')%>%
  mutate(logEndpoints = log(Endpoints))%>%
  rename(Percent = Freq)#renames column

routedf<-rowPerc(xtabs(~exposure.route +effect, aoc))
routef<-as.data.frame(routedf)%>%
  filter(effect %in% c("Y","N"))%>%
  rename(Type= "exposure.route")%>%
  mutate_if(is.numeric, round,0)%>%
  mutate(plot="Exposure.route")%>%
  mutate(Type = case_when(
    Type == "coparental.exposure" ~"Co-Parental Exposure",
    Type == "paternal.exposure" ~ "Paternal Exposure",
    Type == "maternal.exposure" ~ "Maternal Exposure",
    Type == "food" ~ "Food",
    Type == "water" ~ "Water",
    Type == "sediment" ~ "Sediment",
    Type == "media" ~ "Media"))
study_r<-xtabs(~exposure.route +effect,aoc)
routefinal<- data.frame(cbind(routef, study_r))%>% 
  rename(Endpoints='Freq.1')%>%
  rename(category='exposure.route')%>%
  mutate(logEndpoints = log(Endpoints))%>%
  rename(Percent = Freq)#renames column

#### Exploration AO Setup ####

# Master dataset for scatterplots - for Heili's tab.
aoc_v1 <- aoc %>% # start with original dataset
   # full dataset filters.
  mutate(effect_f = factor(case_when(effect == "Y" ~ "Yes",
    effect == "N" ~ "No"),
    levels = c("No", "Yes"))) %>%
  # removing NAs to make data set nicer
  replace_na(list(size.category = 0, shape = "Not Reported", polymer = "Not Reported", life.stage = "Not Reported"))

aoc_setup <- aoc_v1 %>% # start with original dataset
  mutate(size_f = factor(case_when(
    size.category == 1 ~ "1nm < 100nm",
    size.category == 2 ~ "100nm < 1µm",
    size.category == 3 ~ "1µm < 100µm",
    size.category == 4 ~ "100µm < 1mm",
    size.category == 5 ~ "1mm < 5mm",
    size.category == 0 ~ "Not Reported"),
    levels = c("1nm < 100nm", "100nm < 1µm", "1µm < 100µm", "100µm < 1mm", "1mm < 5mm", "Not Reported"))) %>% # creates new column with nicer names and order by size levels.
  # shape category data tidying.
  mutate(shape_f = factor(case_when(
    shape == "fiber" ~ "Fiber",
    shape == "fragment" ~ "Fragment",
    shape == "sphere" ~ "Sphere",
    shape == "Not Reported" ~ "Not Reported"),
    levels = c("Fiber", "Fragment", "Sphere", "Not Reported"))) %>% # order our different shapes.
  # polymer category data tidying.
  mutate(poly_f = factor(case_when(
    polymer == "BIO" ~ "Biopolymer",
    polymer == "EVA" ~ "Polyethylene Vinyl Acetate",
    polymer == "LTX" ~ "Latex",
    polymer == "PA" ~ "Polyamide",
    polymer == "PE" ~ "Polyethylene",
    polymer == "PC" ~ "Polycarbonate",
    polymer == "PET" ~ "Polyethylene Terephthalate",
    polymer == "PI" ~ "Polyisoprene",
    polymer == "PMMA" ~ "Polymethylmethacrylate",
    polymer == "PP" ~ "Polypropylene",
    polymer == "PS" ~ "Polystyrene",
    polymer == "PUR" ~ "Polyurathane",
    polymer == "PVC" ~ "Polyvinylchloride",
    polymer == "PLA" ~ "Polylactic Acid",
    polymer == "Not Reported" ~ "Not Reported"))) %>%
  # taxonomic category data tidying.
  mutate(org_f = factor(organism.group, levels = c("Algae", "Annelida", "Bacterium", "Cnidaria", "Crustacea",
                                                   "Echinoderm", "Fish", "Insect", "Mollusca", "Nematoda", "Plant", "Rotifera", "Mixed"))) %>% # order our different organisms.
  mutate(lvl1_f = factor(case_when(lvl1 == "alimentary.excretory" ~ "Alimentary, Excretory",
    lvl1 == "behavioral.sense.neuro" ~ "Behavioral, Sensory, Neurological",
    lvl1 == "circulatory.respiratory" ~ "Circulatory, Respiratory",
    lvl1 == "community" ~ "Community",
    lvl1 == "fitness" ~ "Fitness",
    lvl1 == "immune" ~ "Immune",
    lvl1 == "metabolism" ~ "Metabolism",
    lvl1 == "microbiome" ~ "Microbiome",
    lvl1 == "stress" ~ "Stress"))) %>% # creates new column with nicer names.
  # Level 2 Data tidying
  mutate(lvl2_f = factor(case_when(lvl2 == "abundance"~"Abundance",
    lvl2 == "actinobacteria" ~ "Actinobacteria",
    lvl2 == "aggressivity"~"Agressivity",
    lvl2 == "ammonia.excretion" ~ "Ammonia Excretion",
    lvl2 == "bacteroidetes"~ "Bacteriodetes",
    lvl2 == "blood"~"Blood",
    lvl2 == "body.condition"~"Body Condition",
    lvl2 == "boldness"~"Boldness",
    lvl2 == "brain.histo"~"Brain Histological Abnormalities",
    lvl2 == "burrowing"~"Burrowing",
    lvl2 == "carb.metabolism"~"Carb Metabolism",
    lvl2 == "chemokines.cytokines"~"Chemokines",
    lvl2 == "circulatory"~"Circulatory",
    lvl2 == "detoxification"~"Detoxification",
    lvl2 == "development"~"Development",
    lvl2 == "digestion"~"Digestion",
    lvl2 == "digestive.enzymes"~"Digestive Enzymes",
    lvl2 == "digestive.tract.histo"~"Digestive Tract Histological Abnormalities",
    lvl2 == "diversity"~ "Diversity",
    lvl2 == "feeding"~ "Feeding",
    lvl2 == "firmicutes"~ "Firmicutes",
    lvl2 == "gall.bladder.histo" ~ "Gall Bladder Histological Abnormalities",
    lvl2 == "gen.metabolism"~ "General Metabolism",
    lvl2 == "gill.histo"~ "Gill Histological Abnormalities",
    lvl2 == "gonad.histo"~"Gonad Histological Abnormalities",
    lvl2 == "growth"~ "Growth",
    lvl2 == "immune.cells"~"Immune Cells",
    lvl2 == "immune.other"~"Immune Other ",
    lvl2 == "intestinal.permeability"~"Intestinal Permeability",
    lvl2 == "kidney.histo"~"Kidney Histological abnormalities",
    lvl2 == "lipid.metabolism"~"Lipid Metabolism",
    lvl2 == "liver.histo"~"Liver Histological Abnormalities",
    lvl2 == "liver.kidney.products" ~ "Liver and Kidney Products",
    lvl2 == "locomotion"~"Locomotion",
    lvl2 == "mortality"~"Mortality",
    lvl2 == "nervous.system"~"Nervous System",
    lvl2 == "oxidative.stress"~"Oxidative Stress",
    lvl2 == "photosynthesis"~ "Photosynthesis",
    lvl2 == "proteobacteria"~"Protebacteria",
    lvl2 == "reproduction"~"Reproduction",
    lvl2 == "respiration"~"Respiration",
    lvl2 == "sexhormones"~"Sex Hormones",
    lvl2 == "shoaling"~"Shoaling",
    lvl2 == "stress"~"Stress",
    lvl2 == "vision.system"~"Vision System"))) %>% #Renames for widget
  mutate(bio_f = factor(case_when(bio.org == "cell"~"Cell", #Bio Org Data Tidying
    bio.org == "organism"~"Organism",
    bio.org == "population"~ "Population",
    bio.org == "subcell"~"Subcell",
    bio.org == "tissue" ~ "Tissue")))%>%
  mutate(vivo_f = factor(case_when(invitro.invivo == "invivo"~"In Vivo",
    invitro.invivo == "invitro"~"In Vitro")))%>% ##Renames for widget (Not using a widget right now, but saving for human health database)
  mutate(life_f = factor(case_when(life.stage == "Early"~"Early",
    life.stage == "Juvenile"~"Juvenile",
    life.stage == "Adult"~"Adult",
    life.stage == "Not Reported"~"Not Reported")))%>% #Renames for widget
  mutate(env_f = factor(case_when(environment == "Freshwater"~"Freshwater",
    environment == "Marine" ~ "Marine",
    environment == "Terrestrial" ~ "Terrestrial"))) %>%
  mutate(dose.mg.L.master.converted.reported = factor(dose.mg.L.master.converted.reported)) %>%
  mutate(dose.particles.mL.master.converted.reported = factor(dose.particles.mL.master.converted.reported)) %>% 
   mutate(dose.um3.mL.master = particle.volume.um3 * dose.particles.mL.master) %>%   #calculate volume/mL
  mutate(af.time_noNA = replace_na(af.time, "Unavailable")) %>% 
  mutate(acute.chronic_f = factor(case_when(af.time_noNA == 10 ~ "Acute",
                                            af.time_noNA == 1 ~ "Chronic",
                                            af.time_noNA == "Unavailable" ~ "Unavailable"))) %>%    #factorize assesment factor time into chronic/acute
  mutate(dose.mg.L.master.AF.noec = dose.mg.L.master * af.noec) %>% 
  mutate(dose.particles.mL.master.AF.noec = dose.particles.mL.master * af.noec) %>% 
  mutate(effect_f = factor(effect)) %>% 
  mutate_if(is.character, as.factor) %>% 
  mutate(effect_10 = case_when(
     effect_f == "Y" ~ 1,
     effect_f == "N" ~ 0))

#### SSD AO Setup ####

# Master dataset for SSDs
aoc_z <- aoc_setup %>% # start with Heili's altered dataset (no filtration for terrestrial data)
  # environment category data tidying.
  mutate(environment.noNA = replace_na(environment, "Not Reported")) %>% # replaces NA to better relabel.
  mutate(env_f = factor(environment.noNA, levels = c("Marine", "Freshwater", "Terrestrial", "Not Reported")))

# final cleanup and factoring  
aoc_z$Species <- as.factor(paste(aoc_z$genus,aoc_z$species)) #must make value 'Species" (uppercase)
aoc_z$Group <- as.factor(aoc_z$organism.group) #must make value "Group"
aoc_z$Group <- fct_explicit_na(aoc_z$Group) #makes sure that species get counted even if they're missing a group
```
## Data Selection
```{r}
df <- aoc_z %>% 
  filter(tech.tier.zero == "N") %>% #gives studies that pass technical quality criteria
  dplyr::select(c(organism.group, exposure.duration.d, lvl1_f, dose.um3.mL.master, dose.mg.L.master, dose.particles.mL.master, life.stage, bio.org, polymer, shape, size.length.um.used.for.conversions, treatments, effect, effect.metric, particle.volume.um3, density.mg.um.3, environment, exposure.route, lvl2_f)) %>% 
  mutate(effect.metric = as.character(effect.metric)) %>% 
  mutate(effect.metric = (case_when(
    effect.metric == "NONEC" ~ "NOEC",
    effect.metric == "HONEC" ~ "NOEC",
    effect.metric == "LOEC" ~ "LOEC",
    effect.metric == "LC50" ~ "LC50",
    effect.metric == "EC50" ~"EC50",
    effect.metric == "EC10" ~ "EC10"
  ))) %>% 
  mutate(effect.metric = replace_na(effect.metric,"not_available")) %>% 
  mutate(effect.metric = as.factor(effect.metric)) %>% 
  drop_na() %>%  #drop missing
mutate_if(~is.numeric(.) && (.) > 0, log10) %>% 
  mutate(effect_10 = case_when( #convert ordinal to numeric
      effect == "Y" ~ 1,
      effect == "N" ~ 0
    )) %>%
  mutate(effect_10 = factor(effect_10)) %>% 
  dplyr::select(-(effect))

#ensure completeness
skim(df)
```

```{r}
# create train, validation, and test splits
# Create calibration and validation splits with tidymodels initial_split() function.
set.seed(4)
df_split <- df %>%
  initial_split(prop = 0.75)
# default is 3/4ths split (but 75% training, 25% testing).

# Create a training data set with the training() function
# Pulls from training and testing sets created by initial_split()
train <- training(df_split)
test <- testing(df_split)

# variable names for resonse & features
y <- "effect_10"
x <- setdiff(names(df), y) 
```

```{r All Model Training, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
# Examine the environment to be sure # of observations looks like the 75/25 split. 3199:1066.
classif_rf <- train(effect_10~., data = train, method = "rf", ntree = 100, tuneLength = 1)

classif_glm <- train(effect_10~., data = train, method = "glm", family = "binomial")

classif_svm <- train(effect_10~., data = train, method = "svmRadial", prob.model = TRUE, tuneLength = 1)#,
                    # preProcess = c("pca","scale","center"))

classif_decTree <- train(effect_10~., data = train, method = "C5.0", preProcess = c("scale","center"))

classif_nnet <- train(effect_10~., data = train, method = "nnet", preProcess=c("scale","center"))

classif_xgbTree <- train(effect_10~., data = train, method = "xgbTree")

#classif_logicBag <- train(effect_10~., data = train, method = "logicBag", ntree = 100)

# explain
yTest <- as.numeric(as.character(test$effect_10))

explainer_classif_rf <- DALEX::explain(classif_rf, label = "rf",
                                       data = test, 
                                       y = yTest)

explainer_classif_glm <- DALEX::explain(classif_glm, label = "glm", 
                                        data = test, y = yTest)

explainer_classif_svm <- DALEX::explain(classif_svm,  label = "svm", 
                                        data = test, y = yTest)

explainer_classif_decTree <- DALEX::explain(classif_rf, label = "Decision Tree",
                                       data = test, 
                                       y = yTest)

explainer_classif_nnet <- DALEX::explain(classif_glm, label = "Neural Net", 
                                        data = test, y = yTest)

explainer_classif_xgbTree <- DALEX::explain(classif_svm,  label = "eXtreme Gradient Boosting Trees", 
                                        data = test, y = yTest)

#explainer_classif_logicBag <- DALEX::explain(classif_svm,  label = "Bagged Logic Regression", 
                                       # data = test, y = yTest)

```
# Modelling
## Model Performance
```{r model performance, include=FALSE}
mp_classif_rf <- model_performance(explainer_classif_rf)
mp_classif_glm <- model_performance(explainer_classif_glm)
mp_classif_svm <- model_performance(explainer_classif_svm)
mp_classif_decTree <- model_performance(explainer_classif_decTree)
mp_classif_nnet <- model_performance(explainer_classif_nnet)
mp_classif_xgbTree <- model_performance(explainer_classif_xgbTree)
#mp_classif_logicBag <- model_performance(explainer_classif_logicBag)
```

### Histogram of Residuals
```{r}
plot(mp_classif_rf, mp_classif_glm, geom ="histogram")
```
```{r}
plot(mp_classif_svm, mp_classif_decTree, geom ="histogram")
```

```{r}
plot(mp_classif_nnet, mp_classif_xgbTree, geom ="histogram")
```
### Precision Recall Curves
```{r}
plot(mp_classif_rf, mp_classif_glm, mp_classif_svm, mp_classif_decTree, mp_classif_nnet, mp_classif_xgbTree, geom ="prc")
```

### Reverse Cumulative Distribution of Residuals
```{r}
# compare residuals plots
resid_dist <- plot(mp_classif_rf, mp_classif_glm, mp_classif_svm, mp_classif_decTree, mp_classif_nnet,
                   mp_classif_xgbTree) + #, mp_classif_logicBag) +
  theme_minimal() +
        theme(legend.position = 'bottom',
              plot.title = element_text(hjust = 0.5)) + 
        labs(y = '')
resid_dist
```
#### Residuals
```{r}
resid_box <- plot(mp_classif_rf, mp_classif_glm, mp_classif_svm, mp_classif_decTree, mp_classif_nnet, mp_classif_xgbTree,
     #mp_classif_logicBag, 
     geom = "boxplot") +
  theme_minimal() +
        theme(legend.position = 'bottom',
              plot.title = element_text(hjust = 0.5)) 
resid_box
```
```{r}
require(gridExtra)
grid.arrange(resid_box,resid_dist, ncol=2)
```

### Lift Curves

Lift curves describe a performance coefficient (lift) over the cumulative proportion of a population. Lift is calculated as the ratio of "yes's" on a certain sample point (for toxicity) divided by the ratio of "yes's" on the whole dataset. $Lift = Predicted Rate/Average Rate$.

```{r eval=FALSE, include=FALSE}
#### DO NOT RUN - MAKES WEIRD PLOT###

# #recode class as character
# df2 <- df %>% 
#  mutate(effect = factor(case_when( #convert ordinal to numeric
#       effect_10 == "1" ~ "Y",
#       effect_10 == "0" ~ "N"
#     )))
# #split test/train
# set.seed(4)
# df2_split <- df2 %>%
#   initial_split(prop = 0.75)
# train2 <- training(df2_split)
# test2 <- testing(df2_split)
# 
# #set controls
# ctrl <- trainControl(method = "cv", classProbs = TRUE,
#                      summaryFunction = twoClassSummary)
# #create models for lift plots
# lift_rf <- train(effect ~., data = train2, method = "rf", ntree = 100, tuneLength = 1,
#                     trControl = ctrl)
# 
# lift_glm <- train(effect~., data = train2, method = "glm", family = "binomial", 
#                      trControl = ctrl)
# 
# lift_svm <- train(effect~., data = train2, method = "svmRadial", prob.model = TRUE, 
#                      tuneLength = 1, trControl = ctrl)#,
#                     # preProcess = c("pca","scale","center"))
# 
# lift_decTree <- train(effect~., data = train2, method = "C5.0", 
#                          preProcess =c("scale","center"), trControl = ctrl) 
# 
# lift_nnet <- train(effect~., data = train2, method = "nnet", preProcess=c("scale","center"),
#                       trControl = ctrl)
# 
# lift_xgbTree <- train(effect~., data = train2, method = "xgbTree", trControl = ctrl)
# 
# 
# ## Generate the test set results
# lift_results <- data.frame(effect = test2$effect)
# lift_results$rf <- predict(lift_rf, test, type = "prob")[,"Y"]
# lift_results$glm <- predict(lift_glm, test, type = "prob")[,"Y"]
# lift_results$svm <- predict(lift_svm, test, type = "prob")[,"Y"]
# lift_results$decTree <- predict(lift_decTree, test, type = "prob")[,"Y"]
# lift_results$nnet <- predict(lift_nnet, test, type = "prob")[,"Y"]
# lift_results$xgbTree <- predict(lift_xgbTree, test, type = "prob")[,"Y"]
# head(lift_results)
# 
# #plot results
# trellis.par.set(caretTheme())
# lift_obj <- lift(effect ~ rf + glm + svm + decTree + nnet + xgbTree, data = lift_results)
# plot(lift_obj, values = 60, auto.key = list(columns = 3,
#                                             lines = TRUE,
#                                             points = FALSE))
```
##### ALT METHOD
http://rstudio-pubs-static.s3.amazonaws.com/436131_3212dcf341cc422590f1a9f52830cfd6.html

```{r}
# transform datasets and model objects into scored data and calculate deciles 
scores_and_ntiles <- prepare_scores_and_ntiles(datasets=list("train","test"),
                                               dataset_labels = list("train data","test data"),
                                               models = list("classif_rf", "classif_decTree",
                                                             "classif_glm", "classif_nnet",
                                                             "classif_svm", "classif_xgbTree"),
                                               model_labels = list("random forest", "Decision
                                                                   tree", "General linear model",
                                                                   "Neural net", "Support Vector
                                                                   machine", "eXtreme Gradient
                                                                   Boosting Trees"),
                                               target_column="effect_10",
                                               ntiles = 100)

# transform data generated with prepare_scores_and_deciles into aggregated data for chosen plotting scope 
plot_input <- plotting_scope(prepared_input = scores_and_ntiles, scope = 'compare_models')
plot_cumgains(data = plot_input, custom_line_colors = RColorBrewer::brewer.pal(2,'Accent'))
```

```{r}
plot_cumlift(data = plot_input,custom_line_colors = RColorBrewer::brewer.pal(2,'Accent'))
```
```{r}
plot_cumresponse(data = plot_input,highlight_ntile = 20, 
                 custom_line_colors = RColorBrewer::brewer.pal(2,'Accent'))
```
```{r}
plot_multiplot(data = plot_input,  custom_line_colors = RColorBrewer::brewer.pal(2,'Accent'))
```

### Roc Curves
```{r}
plot(mp_classif_rf, 
     #mp_classif_glm, 
     mp_classif_svm, 
     #mp_classif_decTree, 
     mp_classif_nnet,
     #mp_classif_xgbTree,
     geom = "roc") +
  ggtitle("ROC Curves - All Models",  
          paste("AUC_rf = ",round(mp_classif_rf$measures$auc,3), 
                paste("AUC_svm = ",round(mp_classif_svm$measures$auc,3)),
          paste("AUC_nnet = ",round(mp_classif_nnet$measures$auc,3))
          )) +
#,  "AUC_glm = 0.799  AUC_svm = 0.798 AUC_decTree = AUC_nnet = AUC_xgbTree = ") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

### Variable Importance
```{r fig.height=7, fig.width=5, warning=FALSE}
vi_classif_rf <- model_parts(explainer_classif_rf, loss_function = loss_root_mean_square)
vi_classif_glm <- model_parts(explainer_classif_glm, loss_function = loss_root_mean_square)
vi_classif_svm <- model_parts(explainer_classif_svm, loss_function = loss_root_mean_square)
vi_classif_decTree <- model_parts(explainer_classif_decTree, loss_function = loss_root_mean_square)
vi_classif_nnet <- model_parts(explainer_classif_nnet, loss_function = loss_root_mean_square)
vi_classif_xgbTree <- model_parts(explainer_classif_xgbTree, loss_function = loss_root_mean_square)
#vi_classif_logicBag <- model_parts(explainer_classif_logicBag, loss_function = loss_root_mean_square)

plot(vi_classif_rf, vi_classif_glm, vi_classif_svm)#, #vi_classif_logicBag)
```
```{r}
plot(vi_classif_decTree, vi_classif_nnet, vi_classif_xgbTree)
```

### Partial Dependence Plot
```{r}
pdp_classif_rf  <- model_profile(explainer_classif_rf, variable = "dose.um3.mL.master", type = "partial")
pdp_classif_glm  <- model_profile(explainer_classif_glm, variable = "dose.um3.mL.master", type = "partial")
pdp_classif_svm  <- model_profile(explainer_classif_svm, variable = "dose.um3.mL.master", type = "partial")
pdp_classif_decTree  <- model_profile(explainer_classif_decTree, variable = "dose.um3.mL.master", type = "partial")
pdp_classif_nnet  <- model_profile(explainer_classif_nnet, variable = "dose.um3.mL.master", type = "partial")
pdp_classif_xgbTree  <- model_profile(explainer_classif_xgbTree, variable = "dose.um3.mL.master", type = "partial")
#pdp_classif_logicBag  <- model_profile(explainer_classif_logicBag, variable = "dose.um3.mL.master", type = "partial")

plot(pdp_classif_rf, pdp_classif_glm, pdp_classif_svm, pdp_classif_decTree, pdp_classif_nnet, pdp_classif_xgbTree)#, pdp_classif_logicBag)
```

```{r}
#Partial Dependence organism type
pdp_classif_rf  <- model_profile(explainer_classif_rf, variable = "organism.group", type = "partial")
pdp_classif_glm  <- model_profile(explainer_classif_glm, variable = "organism.group", type = "partial")
pdp_classif_svm  <- model_profile(explainer_classif_svm, variable = "organism.group", type = "partial")
pdp_classif_decTree  <- model_profile(explainer_classif_decTree, variable = "organism.group", type = "partial")
pdp_classif_nnet  <- model_profile(explainer_classif_nnet, variable = "organism.group", type = "partial")
pdp_classif_xgbTree  <- model_profile(explainer_classif_xgbTree, variable = "organism.group", type = "partial")
#pdp_classif_logicBag  <- model_profile(explainer_classif_logicBag, variable = "dose.um3.mL.master", type = "partial")

plot(pdp_classif_rf$agr_profiles, pdp_classif_glm$agr_profiles, pdp_classif_svm$agr_profiles, pdp_classif_decTree$agr_profiles, pdp_classif_nnet$agr_profiles, pdp_classif_xgbTree$agr_profiles) +
  ggtitle("Contrastive Partial Dependence Profiles", "") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

##### Grouped Partial Dependence Plots
```{r include=FALSE}
plot(pdp_classif_rf)
#Partial Dependence by dose.particles.mL.master and effect.metric
pdp_classif_rf  <- model_profile(explainer_classif_rf, variable = "dose.particles.mL.master", 
                                 groups = "effect.metric")
pdp_classif_glm  <- model_profile(explainer_classif_glm, variable = "dose.particles.mL.master", 
                                 groups = "effect.metric")
pdp_classif_svm  <- model_profile(explainer_classif_svm, variable = "dose.particles.mL.master", 
                                 groups = "effect.metric")
pdp_classif_decTree  <- model_profile(explainer_classif_decTree, variable = "dose.particles.mL.master", 
                                 groups = "effect.metric")
pdp_classif_nnet  <- model_profile(explainer_classif_nnet, variable = "dose.particles.mL.master", 
                                 groups = "effect.metric")
pdp_classif_xgbTree  <- model_profile(explainer_classif_xgbTree, variable = "dose.particles.mL.master", 
                                 groups = "effect.metric")
```

```{r}
#plot neural net and glm
glm <- plot(pdp_classif_glm$agr_profiles) +
  ggtitle("GLM: Partial Dependence Profiles by Dose and Effect Metric", "") +
  xlab("log10(dose particles/mL)") +
  ylab("Average Prediction for Effect (1 = yes, 0 = no)") +
  scale_color_discrete(name = "Model x Effect Metric") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

nnet <- plot(pdp_classif_nnet$agr_profiles) +
  ggtitle("Neural Net: Partial Dependence Profiles by Dose and Effect Metric", "") +
  xlab("log10(dose particles/mL)") +
  ylab("Average Prediction for Effect (1 = yes, 0 = no)") +
  scale_color_discrete(name = "Model x Effect Metric") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
grid.arrange(glm, nnet)
```

```{r}
#plot rf and decision tree
rf <- plot(pdp_classif_rf$agr_profiles) +
  ggtitle("Random Forest:Partial Dependence Profiles by Dose and Effect Metric", "") +
  xlab("log10(dose particles/mL)") +
  ylab("Average Prediction for Effect (1 = yes, 0 = no)") +
  scale_color_discrete(name = "Model x Effect Metric") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

decTree <- plot(pdp_classif_decTree$agr_profiles) +
  ggtitle("Decision Tree: Partial Dependence Profiles by Dose and Effect Metric", "") +
  xlab("log10(dose particles/mL)") +
  ylab("Average Prediction for Effect (1 = yes, 0 = no)") +
  scale_color_discrete(name = "Model x Effect Metric") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
grid.arrange(rf, decTree)
```

```{r}
#plot rf and decision tree
svm <- plot(pdp_classif_svm$agr_profiles) +
  ggtitle("Support Vector Machine:Partial Dependence Profiles by Dose and Effect Metric", "") +
  xlab("log10(dose particles/mL)") +
  ylab("Average Prediction for Effect (1 = yes, 0 = no)") +
  scale_color_discrete(name = "Model x Effect Metric") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

xgbTree <- plot(pdp_classif_xgbTree$agr_profiles) +
  ggtitle("eXtreme Gradient Boosing Trees: Partial Dependence Profiles by Dose and Effect Metric", "") +
  xlab("log10(dose particles/mL)") +
  ylab("Average Prediction for Effect (1 = yes, 0 = no)") +
  scale_color_discrete(name = "Model x Effect Metric") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
grid.arrange(svm, xgbTree)
```

### Confusion Matrices

```{r include=FALSE}
#predict test data
pred_rf <- predict(classif_rf, newdata = test)
pred_glm <- predict(classif_glm, newdata = test)
pred_svm <- predict(classif_svm, newdata = test)
pred_decTree <- predict(classif_decTree, newdata = test)
pred_nnet <- predict(classif_nnet, newdata = test)
pred_xgbTree <- predict(classif_xgbTree, newdata = test)

#build confusion matrices
table_rf <- data.frame(confusionMatrix(pred_rf, test$effect_10)$table)
table_glm <- data.frame(confusionMatrix(pred_glm, test$effect_10)$table)
table_svm <- data.frame(confusionMatrix(pred_svm, test$effect_10)$table)
table_decTree <- data.frame(confusionMatrix(pred_decTree, test$effect_10)$table)
table_nnet <- data.frame(confusionMatrix(pred_nnet, test$effect_10)$table)
table_xgbTree <- data.frame(confusionMatrix(pred_xgbTree, test$effect_10)$table)

#build plots
plotTable_rf <- table_rf %>% mutate(goodbad = ifelse(table_rf$Prediction == table_rf$Reference, "good", "bad")) %>%
  group_by(Reference) %>% mutate(prop = Freq/sum(Freq))

plotTable_glm <- table_glm %>% mutate(goodbad = ifelse(table_glm$Prediction == table_glm$Reference, "good", "bad")) %>%
  group_by(Reference) %>% mutate(prop = Freq/sum(Freq))

plotTable_svm <- table_svm %>% mutate(goodbad = ifelse(table_svm$Prediction == table_svm$Reference, "good", "bad")) %>%
  group_by(Reference) %>% mutate(prop = Freq/sum(Freq))

plotTable_decTree <- table_decTree %>% mutate(goodbad = ifelse(table_decTree$Prediction == table_decTree$Reference, "good", "bad")) %>%
  group_by(Reference) %>% mutate(prop = Freq/sum(Freq))

plotTable_nnet <- table_nnet %>% mutate(goodbad = ifelse(table_nnet$Prediction == table_nnet$Reference, "good", "bad")) %>%
  group_by(Reference) %>% mutate(prop = Freq/sum(Freq))

plotTable_xgbTree <- table_xgbTree %>% mutate(goodbad = ifelse(table_xgbTree$Prediction == table_xgbTree$Reference, "good", "bad")) %>%
  group_by(Reference) %>% mutate(prop = Freq/sum(Freq))
```

Confusion Matrices for the six tested models are below.
```{r}
# fill alpha relative to sensitivity/specificity by proportional outcomes within reference groups (see dplyr code above as well as original confusion matrix for comparison)
CM_rf <- ggplot(data = plotTable_rf, mapping = aes(x = Reference, y = Prediction, fill = prop)) +
  geom_tile() +
  geom_text(aes(label = Freq), vjust = .5, fontface  = "bold", alpha = 1) +
  scale_fill_gradient(low = "white", high = "cyan4", name = "Proportion") +
  scale_x_discrete(labels = c("No Effect", "Effect")) +
  scale_y_discrete(labels = c("No Effect", "Effect")) +
  ggtitle("Random Forest",
          paste("Accuracy = ", 100 * round(mp_classif_rf$measures$accuracy,3), "%")) +
  theme_bw() +
    theme(plot.title = element_text(hjust = 0.5),
          legend.position = "none")

CM_glm <- ggplot(data = plotTable_glm, mapping = aes(x = Reference, y = Prediction, fill = prop)) +
  geom_tile() +
  geom_text(aes(label = Freq), vjust = .5, fontface  = "bold", alpha = 1) +
  scale_fill_gradient(low = "white", high = "cyan4", name = "Proportion") +
  scale_x_discrete(labels = c("No Effect", "Effect")) +
  scale_y_discrete(labels = c("No Effect", "Effect")) +
  ggtitle("General Linear Model",
          paste("Accuracy = ", 100 * round(mp_classif_glm$measures$accuracy,3), "%")) +
  theme_bw() +
    theme(plot.title = element_text(hjust = 0.5),
          legend.position = "none")

CM_svm <- ggplot(data = plotTable_svm, mapping = aes(x = Reference, y = Prediction, fill = prop)) +
  geom_tile() +
  geom_text(aes(label = Freq), vjust = .5, fontface  = "bold", alpha = 1) +
  scale_fill_gradient(low = "white", high = "cyan4", name = "Proportion") +
  scale_x_discrete(labels = c("No Effect", "Effect")) +
  scale_y_discrete(labels = c("No Effect", "Effect")) +
  ggtitle("Support Vector Machine",
          paste("Accuracy = ", 100 * round(mp_classif_svm$measures$accuracy,3), "%")) +
  theme_bw() +
 theme(plot.title = element_text(hjust = 0.5),
          legend.position = "none")

CM_decTree <- ggplot(data = plotTable_decTree, mapping = aes(x = Reference, y = Prediction, fill = prop)) +
  geom_tile() +
  geom_text(aes(label = Freq), vjust = .5, fontface  = "bold", alpha = 1) +
  scale_fill_gradient(low = "white", high = "cyan4", name = "Proportion") +
  scale_x_discrete(labels = c("No Effect", "Effect")) +
  scale_y_discrete(labels = c("No Effect", "Effect")) +
  ggtitle("Decision Tree",
          paste("Accuracy = ", 100 * round(mp_classif_decTree$measures$accuracy,3), "%")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),
          legend.position = "none")

CM_nnet <- ggplot(data = plotTable_nnet, mapping = aes(x = Reference, y = Prediction, fill = prop)) +
  geom_tile() +
  geom_text(aes(label = Freq), vjust = .5, fontface  = "bold", alpha = 1) +
  scale_fill_gradient(low = "white", high = "cyan4", name = "Proportion") +
  scale_x_discrete(labels = c("No Effect", "Effect")) +
  scale_y_discrete(labels = c("No Effect", "Effect")) +
  ggtitle("Neural Net",
          paste("Accuracy = ", 100 * round(mp_classif_nnet$measures$accuracy,3), "%")) +
  theme_bw() +
    theme(plot.title = element_text(hjust = 0.5),
          legend.position = "none")

CM_xgbTree <- ggplot(data = plotTable_xgbTree, mapping = aes(x = Reference, y = Prediction, fill = prop)) +
  geom_tile() +
  geom_text(aes(label = Freq), vjust = .5, fontface  = "bold", alpha = 1) +
  scale_fill_gradient(low = "white", high = "cyan4", name = "Proportion") +
  scale_x_discrete(labels = c("No Effect", "Effect")) +
  scale_y_discrete(labels = c("No Effect", "Effect")) +
  ggtitle("eXtreme Gradient Boosting Trees",
          paste("Accuracy = ", 100 * round(mp_classif_xgbTree$measures$accuracy,3), "%")) +
  theme_bw() +
    theme(plot.title = element_text(hjust = 0.5),
          legend.position = "none")

grid.arrange(CM_rf, CM_glm, CM_svm, CM_decTree, CM_nnet, CM_xgbTree,
             ncol = 2, top = "Confusion Matrices for ML Models")
```

### Breakdown

```{r}
# create a single observation
new_cust <- test$effect_10 %>% as.data.frame()

# compute breakdown distances
new_cust_glm <- predict_parts(explainer_classif_glm, new_observation = test, type = "break_down")
new_cust_rf <- predict_parts(explainer_classif_rf, new_observation = test, type = "break_down")
new_cust_svm <- predict_parts(explainer_classif_svm, new_observation = test, type = "break_down")
new_cust_decTree <- predict_parts(explainer_classif_decTree, new_observation = test, type = "break_down")
new_cust_nnet <- predict_parts(explainer_classif_nnet, new_observation = test, type = "break_down")
new_cust_xgbTree <- predict_parts(explainer_classif_xgbTree, new_observation = test, type = "break_down")
#new_cust_logicBag <- predict_parts(explainer_classif_logicBag, new_observation = test, type = "break_down")

# class of prediction_breakdown output
class(new_cust_rf)

# check out the top 10 influential variables for this observation
new_cust_rf[1:10, 1:5]
```

```{r}
plot(new_cust_glm, new_cust_rf, new_cust_svm, new_cust_decTree, new_cust_nnet, new_cust_xgbTree)#,
     #new_cust_logicBag)
```
```{r}
library(ggplot2)

# filter for top 10 influential variables for each model and plot
list(new_cust_glm, new_cust_rf) %>%
  purrr::map(~ top_n(., 11, wt = abs(contribution))) %>%
  do.call(rbind, .) %>%
  mutate(variable = paste0(variable, " (", label, ")")) %>%
  ggplot(aes(contribution, reorder(variable, contribution))) +
  geom_point() +
  geom_vline(xintercept = 0, size = 3, color = "white") +
  facet_wrap(~ label, scales = "free_y", ncol = 1) +
  ylab(NULL)
```

```{r}
library(ggplot2)

# filter for top 10 influential variables for each model and plot
list(new_cust_svm, new_cust_decTree) %>%
  purrr::map(~ top_n(., 11, wt = abs(contribution))) %>%
  do.call(rbind, .) %>%
  mutate(variable = paste0(variable, " (", label, ")")) %>%
  ggplot(aes(contribution, reorder(variable, contribution))) +
  geom_point() +
  geom_vline(xintercept = 0, size = 3, color = "white") +
  facet_wrap(~ label, scales = "free_y", ncol = 1) +
  ylab(NULL)
```

```{r}
library(ggplot2)

# filter for top 10 influential variables for each model and plot
list(new_cust_nnet,new_cust_xgbTree) %>%
  purrr::map(~ top_n(., 11, wt = abs(contribution))) %>%
  do.call(rbind, .) %>%
  mutate(variable = paste0(variable, " (", label, ")")) %>%
  ggplot(aes(contribution, reorder(variable, contribution))) +
  geom_point() +
  geom_vline(xintercept = 0, size = 3, color = "white") +
  facet_wrap(~ label, scales = "free_y", ncol = 1) +
  ylab(NULL)
```

# Final Model
All in all random forest is my final model of choice: it appears the more balanced and is the most accurate overall. This model will now be tuned and refined for maximum performance.

## Tuning
http://rstudio-pubs-static.s3.amazonaws.com/480890_237ad52b09b6440e9c849a3c07a04d2f.html
```{r Tuning, include=FALSE}
cache = TRUE
set.seed(1000)
train_control <- trainControl(method = "repeatedcv", 
                              number = 10,
                              repeats = 5,
                              verboseIter = TRUE,
                              allowParallel = TRUE,
                              summaryFunction = multiClassSummary)
cache = TRUE
set.seed(1000)

start_time <- Sys.time() # Start timer

my_grid1 <- expand.grid(mtry = 1:17)

rf1 <- train(effect_10 ~ .,
             data = train,
             method = "rf",
             metric = "Accuracy",
             tuneGrid = my_grid1,
             trControl = train_control)
rf1
```

Plot tuning.

```{r}
cache = TRUE

my_plot <- function(model) {
    theme_set(theme_minimal())
    u <- model$results %>%
        select(mtry, Accuracy, Kappa, F1, Sensitivity,
              Specificity,Pos_Pred_Value, Neg_Pred_Value, 
               Precision, Recall, Detection_Rate) %>%
        gather(a, b, -mtry)

    u %>% ggplot(aes(mtry, b)) + geom_line() + geom_point() + 
        facet_wrap(~ a, scales = "free") + 
        labs(x = "Number of mtry", y = NULL, 
             title = "The Relationship between Model Performance and mtry")
}

rf1 %>% my_plot()
```

I will further refine this model using recursive feature elimination and compare accuracy with the other models.

```{r Recursive Feature Elimination}

my_ctrl <- rfeControl(functions = rfFuncs, #random forests
                      method = "repeatedcv",
                      verbose = FALSE,
                      repeats = 5,
                      returnResamp = "all")

rfProfile <- rfe(y = df$effect_10, # set dependent variable
              x = df %>% 
                dplyr::select(-effect_10),
              rfeControl = my_ctrl,
               size = c(1:2, 4, 6, 8, 10, 12, 13))
rfProfile
```

## Predictor Selection
The following variables are those that were picked in the final (most accurate) model. 
```{r}
#get variable names picked in the final model
predictors(rfProfile)
```
The first 6 models are shown below with their corresponding accuracy and kappa values. 
```{r}
head(rfProfile$resample)
```

```{r}
trellis.par.set(caretTheme())
#plot(rfProfile, type = c("g", "o"), metric = "Accuracy")
ggplot(rfProfile) + theme_bw()
```
```{r}
plot1 <- xyplot(rfProfile, type = c("g", "p", "smooth"), main = "Accuracy of individual resampling results")
plot2 <- densityplot(rfProfile, 
                     subset = Variables < 5, 
                     adjust = 1.25, 
                     as.table = TRUE, 
                     xlab = "Accuracy Estimates", 
                     pch = "|")
print(plot1, split=c(1,1,1,2), more=TRUE)
print(plot2, split=c(1,2,1,2))
```

### Final Predictor Selection
```{r}
my_size <- pickSizeTolerance(rfProfile$results, metric = "Accuracy", tol = 1, maximize = TRUE)
# higher tol (~10) gives you less variables
# lower tol (~1) gives you more variables - "I'd like the simplest model within 1% of the best model."
accuracy1 <- pickVars(rfProfile$variables, size = my_size)
accuracy1
```
A random forest model with the above four predictors is within 1% of the best model. If a more accurate model is desired, and data is available, additional factors can be included in the model. Shown below are factors that should be included for a model that is within 0.5% accuracy of the best model.

```{r}
my_size <- pickSizeTolerance(rfProfile$results, metric = "Accuracy", tol = 0.5, maximize = TRUE)
# higher tol (~10) gives you less variables
# lower tol (~1) gives you more variables - "I'd like the simplest model within 1% of the best model."
accuracy0.5 <- pickVars(rfProfile$variables, size = my_size)
accuracy0.5
```
Below is a table showing which variables are included in models of varying accuracies.

```{r}
my_size <- pickSizeTolerance(rfProfile$results, metric = "Accuracy", tol = 0.1, maximize = TRUE)
accuracy0.1 <- pickVars(rfProfile$variables, size = my_size)
my_size <- pickSizeTolerance(rfProfile$results, metric = "Accuracy", tol = 0.3, maximize = TRUE)
accuracy0.3 <- pickVars(rfProfile$variables, size = my_size)
my_size <- pickSizeTolerance(rfProfile$results, metric = "Accuracy", tol = 5, maximize = TRUE)
accuracy5 <- pickVars(rfProfile$variables, size = my_size)
my_size <- pickSizeTolerance(rfProfile$results, metric = "Accuracy", tol = 10, maximize = TRUE)
accuracy10 <- pickVars(rfProfile$variables, size = my_size)

varsTable = list('0.1%' =accuracy0.1, '1%' = accuracy1, '5%' = accuracy5, '10%' = accuracy10)

#$make padded dataframe
na.pad <- function(x,len){
    x[1:len]
}

makePaddedDataFrame <- function(l,...){
    maxlen <- max(sapply(l,length))
    data.frame(lapply(l,na.pad,len=maxlen),...)
}
#fancy print
kable(makePaddedDataFrame(
  list('Accuracy (0.1)' = accuracy0.1, 'Accuracy (0.3)' = accuracy0.3, "Accuracy (1)" = accuracy1, 'Accuracy (5)' = accuracy5, 'Accuracy (10)' = accuracy10)
  ),
      caption = "Variables to include in Random Forest to achieve Model Accuracy (within x% of best model)",
    footnote = "Random Forest Recursive Feature Elimination")
```
### Final Model Tuning
While the absolute best model contains 18 variables with an estimated accuracy of 93.89%, it is possible to achieve ~93% accuracy with just eight variables, as displayed in the above table. The advantage of including fewer variables is that less information is needed to accurately predict toxicity. The final model with these eight variables will be further refined through an iterative tuning process to determine the optimal number of variables to be randomly collected for sampling at each split (mtry), and the number of branches to grow after each split.Code for the following section is inspired from https://rpubs.com/phamdinhkhanh/389752. 
```{r}
final_df <-df %>% 
  dplyr::select(c(effect.metric, dose.mg.L.master, lvl2_f, dose.particles.mL.master, effect_10, dose.um3.mL.master, organism.group, exposure.duration.d, treatments)) %>% 
  droplevels()

#ensure completeness
skim(final_df)
```

```{r}
# create train, validation, and test splits
# Create calibration and validation splits with tidymodels initial_split() function.
set.seed(4)
df_final_split <- final_df %>%
  initial_split(prop = 0.75)
# default is 3/4ths split (but 75% training, 25% testing).

# Create a training data set with the training() function
# Pulls from training and testing sets created by initial_split()
train_final <- training(df_final_split)
test_final <- testing(df_final_split)

# variable names for resonse & features
y <- "effect_10"
x <- setdiff(names(final_df), y) 
```

The mtry parameter will be optimized below:

```{r}
tunegrid <- expand.grid(.mtry=(1:7))
                       # , .ntree=c(500, 1000, 1500, 2000, 2500)) #would like to optimize but can't figure out how!

fitControl <- trainControl(## 10-fold CV
                           method = "repeatedcv",
                           number = 10,
                           ## repeat 3 times
                           repeats = 3)

nrow(tunegrid)

set.seed(825)
tuneFit <- train(effect_10 ~ ., data = train_final, 
                 method = "rf", 
                 trControl = fitControl, 
                 verbose = FALSE, 
                 ## Now specify the exact models 
                 ## to evaluate:
                 metric = 'Accuracy',
                 tuneGrid = tunegrid)
tuneFit
```

```{r}
plot(tuneFit)
```
Now that mtry is optimized, the number of trees will be optimized, holding mtry constant.

```{r}
# Manual Search
control <- trainControl(method="repeatedcv", number=10, repeats=3, search="grid")
tunegrid <- expand.grid(.mtry=c(7)) #replace with optimal value from previous chunk
modellist <- list()
for (ntree in c(1000, 1500, 2000, 2500)) {
	set.seed(123)
	fit <- train(effect_10~., data=train_final, method="rf", metric='Accuracy', 
	             tuneGrid=tunegrid, trControl=control, ntree=ntree)
	key <- toString(ntree)
	modellist[[key]] <- fit
}
# compare results
results <- resamples(modellist)
summary(results)
```

```{r}
dotplot(results)
```

### Final Model
Now that we have optimized the variables for our final model, we will build and save the final simplified model that has the highest accuracy.
```{r Final Model}
tunegrid <- expand.grid(.mtry=c(7))

final_model <- train(effect_10~., data = train_final, method = "rf", ntree = 2500, tuneLength = 5)

# explain
yTest <- as.numeric(as.character(test_final$effect_10))

explainer_final_model <- DALEX::explain(final_model, label = "rf",
                                       data = test_final, 
                                       y = yTest)

mp_classif_final_model <- model_performance(explainer_final_model)
mp_classif_final_model
```
#### Performance
Performance metrics for the final model are below.

##### ROC Curve

```{r}
plot(mp_classif_final_model,
     geom = "roc") +
  ggtitle("ROC Curve - Final Model",  
          paste("AUC = ",round(mp_classif_final_model$measures$auc,2),
                paste("Accuracy = ", 100 * round(mp_classif_final_model$measures$accuracy,3), "%")
          )) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        legend.position = "none")
```

##### Confusion Matrix
The tuned, final, simplified model is now validated using test set data.
```{r}
test_pred <- predict(final_model, newdata = test_final)
confusionMatrix(table(test_pred, test_final$effect_10))
```
This confusion matrix is plotted below.
```{r}
table <- data.frame(confusionMatrix(test_pred, test_final$effect_10)$table)

plotTable <- table %>%
  mutate(goodbad = ifelse(table$Prediction == table$Reference, "good", "bad")) %>%
  group_by(Reference) %>%
  mutate(prop = Freq/sum(Freq))

# fill alpha relative to sensitivity/specificity by proportional outcomes within reference groups (see dplyr code above as well as original confusion matrix for comparison)
ggplot(data = plotTable, mapping = aes(x = Reference, y = Prediction, fill = prop)) +
  geom_tile() +
  geom_text(aes(label = Freq), vjust = .5, fontface  = "bold", alpha = 1) +
  scale_fill_gradient(low = "white", high = "cyan4", name = "Proportion") +
  scale_x_discrete(labels = c("No Effect", "Effect")) +
  scale_y_discrete(labels = c("No Effect", "Effect")) +
  theme_bw()
```

# Machine Settings

Machine Info:
```{r}
Sys.info()[c(1:3,5)]
```

Session Info:
```{r}
sessionInfo()
```