### 2021-02-27 Modelling/simulated distributions.Rmd
simulated distributions

---
title: "Simulated Probabilistic Distributions"
author: "Scott Coffin"
date: "2/24/2021"
output:   
  html_document:
    code_folding: hide
    theme: journal
    toc: yes
    toc_float: yes
    toc_depth: 6
    number_sections: true
    includes:
     # after_body: footer.html
  word_document:
    toc: yes
---
#Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      warning=FALSE, message=FALSE,time_it = TRUE) #report
```

```{r library}
# Bimodal universal shape distribution for environmental microplastic from Kooi and Koelmans (2019)
library(truncnorm)
library(tidyverse)
library(MonteCarlo)# Monte Carlo Simulation 
library(gridExtra)
library(msm) ## rtnorm - get upper and lower limit of shape distribution
library(GeneralizedHyperbolic) ## normal-inverse Gaussian
library(skimr)
```
# Resources: 
https://pubs.acs.org/doi/suppl/10.1021/acs.est.0c02982/suppl_file/es0c02982_si_001.pdf https://pubs.acs.org/doi/suppl/10.1021/acs.estlett.9b00379/suppl_file/ez9b00379_si_001.pdf
https://pubs.acs.org/doi/10.1021/acs.estlett.9b00379

<!-- ## Parameters -->
<!-- ```{r} -->
<!-- #average parameters from distribution -->
<!-- f1 = 0.06 -->
<!-- f2 = 0.94 -->
<!-- delta1 = 0.03 -->
<!-- delta2 = 0.19 -->
<!-- mu1 = 0.08 -->
<!-- mu2 = 0.44 -->
<!-- x = runif(n=100) -->

<!-- #define list from above -->
<!-- param_list = list("f1" = f1, "f2" = f2, "delta1" = delta1, "delta2" = delta2, "mu1" = mu1, "mu2" = mu2, "x" = x) -->

<!-- #permanently define euler's number -->
<!-- e <- exp(1) -->
<!-- lockBinding("e", globalenv()) -->

<!-- #starting conditions for x -->
<!-- x = runif(n = 100) #particle size in micrometers -->

<!-- #equation 4 -->
<!-- y = f1 * (1/sqrt(2 * pi * delta1 ^ 2)) * (e^-(x - mu1)^2 /2 *delta1 ^ 2)+ f2 * (1/sqrt(2 * pi * delta2 ^ 2)) * (e^-(x - mu2)^2 /2 *delta2 ^ 2) -->

<!-- #define equation 4 -->
<!-- binom <- function(x, f1, f2, delta1, delta2, mu1, mu2){ -->
<!--   #generate sample -->
<!--   #x = runif(n=1000) -->
<!--   #calculate CSF -->
<!--   y <- f1 * (1/sqrt(2 * pi * delta1 ^ 2)) * (e^-(x - mu1)^2 /2 *delta1 ^ 2)+ f2 * (1/sqrt(2 * pi * delta2 ^ 2)) * (e^-(x - mu2)^2 /2 *delta2 ^ 2) -->
<!--   #return result -->
<!--   return(list("y" = y)) -->
<!-- } -->

<!-- MC_result <- MonteCarlo(func=binom, nrep=1, param_list=param_list) -->

<!-- df<-MakeFrame(MC_result) -->
<!-- head(df) -->
<!-- tbl <- tibble(df) -->

<!-- ``` -->
<!-- #### Size -->

<!-- ```{r} -->
<!-- # From Merel Kooi -->
<!-- #rm(list=ls()) -->

<!-- ## Parameters -->
<!-- n = 1E5     #number of values to generate -->
<!-- xmin = 1   #minimum size (um) -->
<!-- xmax = 5000 #maximum size (um) -->
<!-- alpha = 1.6 #exponent of power law distribution -->

<!-- ## Generate random data given parameters of the power law distribution -->
<!-- ## Based on PDF -->

<!-- ## First approach  -->
<!-- a = -alpha     # It has to be negative -->
<!-- y = runif(n)   # Number of samples -->
<!-- Size1 = ((xmax^(a+1) - xmin^(a+1))*y + xmin^(a+1))^(1/(a+1)) -->

<!-- ## Second approach -->
<!-- ## Function that will randomly generate 1 value from the distribution -->
<!-- size.func <- function (X){ -->
<!--   success <- FALSE -->
<!--   while (!success){ -->
<!--     U = runif(1, 0, 1) -->
<!--     X = xmin*(1-U)^(1/(1-alpha)) -->
<!--     success <- X < xmax } -->
<!--   return(X) -->
<!-- } -->

<!-- ## df to store your generated results in -->
<!-- Size2 <- data.frame(Size = numeric(0)) -->

<!-- ## loop through function n times (can probably do this in a function too) -->
<!-- for(i in 1:n){ -->
<!--   X <- size.func() -->
<!--   Size2 <- rbind(Size2, X) -->
<!-- } -->

<!-- colnames(Size2) <- "Size2" -->

<!-- ## check results - visual comparison -->
<!-- Size <- cbind(Size1, Size2) -->

<!-- p1 <- ggplot(Size, aes(x = Size1)) + geom_histogram() + scale_x_log10() -->
<!-- p2 <- ggplot(Size, aes(x = Size2)) + geom_histogram() + scale_x_log10() -->
<!-- grid.arrange(p1, p2, -->
<!--              top = "Comparison of Two Methods to Generate Distributions") -->
<!-- ``` -->

<!-- #### Shape -->

<!-- The most abundant shape category of microplastic in water and sediment is fibers (48.5%), followed by fragments (31%), beads (6.5%), films (5.5%), and foam (3.5%)(3). Combining these abundance data with the triangular shape distributions (Table S2 and Figure S2) resulted in a continuous bimodal microplastic shape distribution (Figure 2A). The fitted parameter values for eq 4 were as follows: f1 = 0.06, f2 = 0.94, σ1 = 0.03, σ2 = 0.19, μ1 = 0.08, and μ2 = 0.44 (see Table S5 for standard errors). A Pearson χ2 test indicated that the optimized distribution fits the data well (the fitted model did not differ significantly from the data; p = 0.231 > 0.05). The distribution is dominated by fibers and fragments (CSF = 0.25–0.75) but also has a distinct second peak at a CSF of 0.07, which is mainly attributed to sheets (Figure 2A). The distribution captures the main features of shapes encountered in the environment as well as the relative abundances of these (now continuous) shapes in one go. Most illustrative though is the continuous character of microplastic shape. -->

<!-- ```{r} -->
<!-- # equation for shape -->

<!-- ## Parameters -->
<!-- n = 1E5     #number of values to generate -->
<!-- f1 = 0.059    -->
<!-- sigma1 = 0.076  -->
<!-- mu1 =  0.030 -->
<!-- f2 = 0.941 -->
<!-- sigma2 = 0.441 -->
<!-- mu2 =  0.189 -->

<!-- y=runif(n) #number of samples -->

<!-- ## Generate random data given parameters of the bimodal normal distributions Based on PDF -->

<!-- ## First approach  -->
<!-- CSF1 = (f1 * (1 / sqrt(2 * pi() * sigma1 ^ 2) * e ^ -(x - mu1) ^ 2 / (2 * sigma1 ^ 2) + f2 * (1 / sqrt(2 * pi() * sigma2 ^ 2) * e ^ -(x - mu2) ^ 2 / (2 * sigma2 ^ 2)))) -->

<!-- #a = -alpha     # Ithas to be negative -->
<!-- #y = runif(n)   # Number of samples -->
<!-- #Size1 = ((xmax^(a+1) - xmin^(a+1))*y + xmin^(a+1))^(1/(a+1)) -->

<!-- ``` -->

<!-- #### Density -->

# Generate Particles

Shape is defined as:
Simplifed equation for corey shape factor:

$CSF = H/sqrt(LW)$

Equation for abundance of particles based on density (two distributions). (Note that this is *equation 4* in Kooi and Koelmans (2019).)

$y = f1(\frac{1}{sqrt(2pi()\sigma_1^2})e^{-(x-\mu_1)^2/2\sigma_1^2} + f2(\frac{1}{sqrt(2pi()\sigma_2^2})e^{-(x-\mu_2)^2/2\sigma_2^2}$

For polymer abundance, for the K&K2019 paper I used the fixed values from Burns & Boxall review, and went straight for density. Given these fixed relative abundances, you could calculate polymer type randomly for N particles via:

pols <- c("PE", "PP", "PS")
rel.ab <- c(0.5, 0.3, 0.2)
n = 1000
sample(pols, size = n, prob = rel.ab, replace = TRUE)

Particle volume is determined according to shape as follows:

$V = \frac{\pi}{6}L^3CSF^2$

Where CSF = corey shape factor and L = length.

Particle mass is estimated as follows:

$m = pV*\frac{1}{1e12}*1000$

Where *m* is the mass (mg), *p* is density (g/cm^3), *V* is volume (um^3) - which is calculated by the cube of the length (um) of each particle, and additional conversion factors for mg to g (x1000) and cm^3 to um^3 (1e-12).
```{r generate size, include=FALSE}
#rm(list=ls())
n = 10000
## Generate values for the three distributions
set.seed(123)

##++++++++++++++++++++++++++++++++++++++++
## SIZE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

xmin = 1 #UM
alpha = 1.6 #2.7 is preferred in unpublished Kooi et al.

X.func <- function (X){
  success <- FALSE
  while (!success){
    U = runif(1, 0, 1)
    X = xmin*(1-U)^(1/(1-alpha))
    success <- X < 5000} ##should be smaller than 5000 um 
  return(X)
}

Data <- data.frame(Size = numeric(0));

for(i in 1:n){
  X <- X.func()
  Data <- rbind(Data, X)
}

colnames(Data) <- c("Size")

min(Data$Size) ##20 um
max(Data$Size) ##5000 um

##++++++++++++++++++++++++++++++++++++++++
## SHAPE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

mu1 <- 0.08
mu2 <- 0.44
sd1 <- 0.03
sd2 <- 0.19
lambda1 <- 0.06
lambda2 <- 0.94

#Sample N random uniforms U
U =runif(n)

#Sampling from the mixture
for(i in 1:n){
  if(U[i]<lambda1){
    Data$Shape[i] = rtnorm(1,mu1,sd1, lower = 0, upper = 1)
  }else{
    Data$Shape[i] = rtnorm(1,mu2,sd2, lower = 0, upper = 1)
  }
}

min(Data$Shape) ##0
max(Data$Shape) ##1

##++++++++++++++++++++++++++++++++++++++++
## DENSITY DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

d.alpha = 73.8 #tail heaviness
d.beta = 69.9  #asymmetry
d.mu = 0.840   #location
d.delta = 0.0972 #scale

D.func <- function (D){
  success <- FALSE
  while (!success){
    D = rnig(1, mu = d.mu, alpha = d.alpha, beta = d.beta, delta = d.delta)
    success <- D < 2.63} ## include upper limit of 2.63, the max. 
  return(D)
}

Dens <- data.frame(Density = numeric(0));

for(i in 1:n){
  X <- D.func()
  Dens <- rbind(Dens, X)
}

colnames(Dens) <- c("Density")

Data <- cbind(Data, Dens)

##++++++++++++++++++++++++++++++++++++++++
## Polymer DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

pols <- c("PE", "PP", "PS")
rel.ab <- c(0.5, 0.3, 0.2)
polymer <- sample(pols, size = n, prob = rel.ab, replace = TRUE)
Data <- cbind(Data, polymer)
```

```{r}
##++++++++++++++++++++++++++++++++++++++++
## Calculate Volume/Volume and estimate mass/volume
##++++++++++++++++++++++++++++++++++++++++
Data_final <- Data %>% 
  mutate(vol.L = 1000) %>% 
  mutate(vol.mL = vol.L * 1000) %>% 
  mutate(mass.mg = Density * Size^3 * 1E-9) %>% 
  mutate(um3 = (Size^3)) %>% 
  mutate(dose.um3.mL.master = um3 / vol.mL) %>% 
  mutate(dose.particles.mL.master = 1 / vol.mL) %>% 
  mutate(dose.mg.L.master = mass.mg / vol.L)
skim(Data_final)
```

Parameters:
Shape, density, size, 
why do 3 polymers? Leave polymer out for now. 
For mechanistical point of view- polymer does not matter.

```{r}
##++++++++++++++++++++++++++++++++++++++++
## Quick check results
##++++++++++++++++++++++++++++++++++++++++
#p1 <- ggplot(Data_final, aes(x = Size, fill = polymer)) + geom_histogram(bins = 50, alpha = 0.7) + scale_x_log10()
p2 <- ggplot(Data_final, aes(x = Shape, fill = polymer)) + geom_histogram(bins = 50, alpha = 0.7)
p3 <- ggplot(Data_final, aes(x = Density, fill = polymer)) + geom_histogram(bins = 50, alpha = 0.7)
p4 <- ggplot(Data_final, aes(x = polymer, fill = polymer)) + geom_histogram(stat = "count")
#p5 <- ggplot(Data_final, aes(x = mass.mg, fill = polymer)) + geom_histogram(bins = 50, alpha = 0.7 )+ scale_x_log10()
#p6 <- ggplot(Data_final, aes(x = um3, fill = polymer)) + geom_histogram(bins = 50, alpha = 0.7)+ scale_x_log10()
p7 <- ggplot(Data_final, aes(x = dose.um3.mL.master, fill = polymer)) + geom_histogram(bins = 50, alpha = 0.7)+ scale_x_log10()
p8 <- ggplot(Data_final, aes(x = dose.particles.mL.master, fill = polymer)) + geom_histogram(bins = 50, alpha = 0.7)+ scale_x_log10()
p9 <- ggplot(Data_final, aes(x = dose.mg.L.master, fill = polymer)) + geom_histogram(bins = 50, alpha = 0.7)+ scale_x_log10()
grid.arrange(#p1,
             p2,
             p3,
             p4,
             #p5, 
             #p6,
             p7,
             p8,
             p9,
             top = "Simulated Microplastics Dataset (n = 1e5)")
```

# Predict Toxicities
```{r}
#pull in model from DALEX
final_model <- readRDS("final_rf_model.rds")

### parameters ###
df_1 <- Data_final %>% 
  mutate(treatments = 1) %>% 
  mutate(organism.group = "Crustacea") %>% 
  mutate(exposure.duration.d = 28) %>% 
  mutate(lvl1_f = "Fitness") %>% 
  mutate(life.stage= "Juvenile") %>% 
  mutate(bio.org = "organism.group") %>% 
  mutate(effect.metric = "NOEC") %>% 
  mutate(environment = "Marine") %>% 
  mutate(exposure.route = "water") %>% 
  mutate(lvl2_f = "Mortality")

df_1_pred <- predict(final_model, newdata = df_1, type = "prob")
df_1_pred_final <- cbind(df_1, df_1_pred) %>% 
  rename("ToxicityProbability" = `1`)

df_1_pred_final %>% ggplot(aes(x = Size, y = ToxicityProbability)) + 
  stat_ecdf(geom = "point", size = 2, color = "blue")+
  stat_ecdf(geom = "step", linetype = 'solid', alpha = 0.6, size = 1.5) +
  scale_x_log10() +
  scale_y_continuous(labels = scales::percent)+
  labs(title = "Cumulative Toxicity Probability",
       subtitle = "n = 1000 particles/L; Crustacea acute mortality",
       caption = "Predicted Toxicity from Random Forest")
```
HIGHLIGHT AREA UNDER THE CURVE
If we go down to lower number, we would not adequately mimic distribution. Stick to creating 1E5 particles, then change volume. 
Could capture variability for small bumbers by simulating small number of particles

How to calculate uncertainties? 
  Replace effect parameters by median +- SD, then predict tox for each distribution, plot, then if it is minor, FORGET about it! REsiduals in Merel's models were not randomly distributed -  so imperfect models.

What sizes should be used?
  1-5,000 um.

Wait for Merel to update interactions between shape/size/density/count.

What about max size ingest?
  Tox studies typically do not exposure animals to particles bigger than their mouth size opening, so model may not understand that relationship. 
  Could impose bioavailable fraction AFTER model. JUST do predictions for species that have bioavailable fraction data. Could estimate mouth size from prey to predator fraction. JAMS et al has mouth-size opening parameters.

Non-linear regression to get dose-response.

Mattson, Hasselof, Frontiers in Science - measures size distribution.

Final figure: SINGLE toxicity prediction for each dose. Use environmental data. Make SSD. Risk Characterization. The model's SSD is fundamentally different from the SSD made from monodisperse particles, so it is MORE reliable. 
Big assumption: all particle toxicities are ADDITIVE! Depends on effect mechanism - food dilution.

## Dose-response curves

## SSD

## Sensitivity Analysis
it would be great to use the full approach to find out which model parameters and MP subclasses the HC5 is most sensitive to.And also if that HC5 is sensitive to which of Merel’s compartment-specific MP parameterisations is used.

## Shiny App Tabs
1) ML predictor
2) Chemical data
  Bioaccumulation model, partitioning with chemicals/particles. Would add toxicity from chemicals to toxicity from particles! Use same model from Worm paper. Would show how it would compare with toxicities from particles. We have criticial body burden concept (lipids)