### 2021-06-21 simulated-distributions_FishDev.rmd
bits and bobs for migration

---
title: "Simulated Probabilistic Distributions_ Algae DEV"
author: "Scott Coffin"
date: "2/24/2021"
output:   
  html_document:
    code_folding: hide
    theme: journal
    toc: yes
    toc_float: yes
    toc_depth: 6
    number_sections: true
    includes:
     # after_body: footer.html
  word_document:
    toc: yes
---
#Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      warning=FALSE, message=FALSE,time_it = TRUE) #report
```

```{r library}
# Bimodal universal shape distribution for environmental microplastic from Kooi and Koelmans (2019)
library(truncnorm)
library(tidyverse)
library(MonteCarlo)# Monte Carlo Simulation 
library(gridExtra)
library(msm) ## rtnorm - get upper and lower limit of shape distribution
library(GeneralizedHyperbolic) ## normal-inverse Gaussian
library(skimr)
library(drc) #dose-response curves
```

```{r}
# load tox data from RShiny app
aoc_z <- readRDS(file = "aoc_z.Rda")
```

# Resources: 
https://pubs.acs.org/doi/suppl/10.1021/acs.est.0c02982/suppl_file/es0c02982_si_001.pdf https://pubs.acs.org/doi/suppl/10.1021/acs.estlett.9b00379/suppl_file/ez9b00379_si_001.pdf
https://pubs.acs.org/doi/10.1021/acs.estlett.9b00379

# Particles constant, volume different
## Generate Particles

Shape is defined as:
Simplifed equation for corey shape factor:

$CSF = H/sqrt(LW)$

Equation for abundance of particles based on density (two distributions). (Note that this is *equation 4* in Kooi and Koelmans (2019).)

The most abundant shape category of microplastic in water and sediment is fibers (48.5%), followed by fragments (31%), beads (6.5%), films (5.5%), and foam (3.5%)(3). Combining these abundance data with the triangular shape distributions (Table S2 and Figure S2) resulted in a continuous bimodal microplastic shape distribution (Figure 2A). The fitted parameter values for eq 4 were as follows: f1 = 0.06, f2 = 0.94, σ1 = 0.03, σ2 = 0.19, μ1 = 0.08, and μ2 = 0.44 (see Table S5 for standard errors). A Pearson χ2 test indicated that the optimized distribution fits the data well (the fitted model did not differ significantly from the data; p = 0.231 > 0.05). The distribution is dominated by fibers and fragments (CSF = 0.25–0.75) but also has a distinct second peak at a CSF of 0.07, which is mainly attributed to sheets (Figure 2A). The distribution captures the main features of shapes encountered in the environment as well as the relative abundances of these (now continuous) shapes in one go. Most illustrative though is the continuous character of microplastic shape.

$y = f1(\frac{1}{sqrt(2pi()\sigma_1^2})e^{-(x-\mu_1)^2/2\sigma_1^2} + f2(\frac{1}{sqrt(2pi()\sigma_2^2})e^{-(x-\mu_2)^2/2\sigma_2^2}$

For polymer abundance, for the K&K2019 paper I used the fixed values from Burns & Boxall review, and went straight for density. Given these fixed relative abundances, you could calculate polymer type randomly for N particles via:

pols <- c("PE", "PP", "PS")
rel.ab <- c(0.5, 0.3, 0.2)
n = 1000
sample(pols, size = n, prob = rel.ab, replace = TRUE)

Particle volume is determined according to shape as follows:

$V = \frac{\pi}{6}L^3CSF^2$

Where CSF = corey shape factor and L = length.

Particle mass is estimated as follows:

$m = pV*\frac{1}{1e12}*1000$

Where *m* is the mass (mg), *p* is density (g/cm^3), *V* is volume (um^3) - which is calculated by the cube of the length (um) of each particle, and additional conversion factors for mg to g (x1000) and cm^3 to um^3 (1e-12).
```{r, include=FALSE}
#rm(list=ls())
n = 10000
## Generate values for the three distributions
set.seed(123)

##++++++++++++++++++++++++++++++++++++++++
## SIZE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

xmin = 1 #UM
alpha = 1.6 #2.7 is preferred in unpublished Kooi et al.

X.func <- function (X){
  success <- FALSE
  while (!success){
    U = runif(1, 0, 1)
    X = xmin*(1-U)^(1/(1-alpha))
    success <- X < 5000} ##should be smaller than 5000 um 
  return(X)
}

Data <- data.frame(Size = numeric(0));

for(i in 1:n){
  X <- X.func()
  Data <- rbind(Data, X)
}

colnames(Data) <- c("Size")

min(Data$Size) ##20 um
max(Data$Size) ##5000 um

##++++++++++++++++++++++++++++++++++++++++
## SHAPE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

mu1 <- 0.08
mu2 <- 0.44
sd1 <- 0.03
sd2 <- 0.19
lambda1 <- 0.06
lambda2 <- 0.94

#Sample N random uniforms U
U =runif(n)

#Sampling from the mixture
for(i in 1:n){
  if(U[i]<lambda1){
    Data$Shape[i] = rtnorm(1,mu1,sd1, lower = 0, upper = 1)
  }else{
    Data$Shape[i] = rtnorm(1,mu2,sd2, lower = 0, upper = 1)
  }
}

min(Data$Shape) ##0
max(Data$Shape) ##1

##++++++++++++++++++++++++++++++++++++++++
## DENSITY DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

d.alpha = 73.8 #tail heaviness
d.beta = 69.9  #asymmetry
d.mu = 0.840   #location
d.delta = 0.0972 #scale

D.func <- function (D){
  success <- FALSE
  while (!success){
    D = rnig(1, mu = d.mu, alpha = d.alpha, beta = d.beta, delta = d.delta)
    success <- D < 2.63} ## include upper limit of 2.63, the max. 
  return(D)
}

Dens <- data.frame(Density = numeric(0));

for(i in 1:n){
  X <- D.func()
  Dens <- rbind(Dens, X)
}

colnames(Dens) <- c("Density")

Data <- cbind(Data, Dens)

##++++++++++++++++++++++++++++++++++++++++
## Polymer DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

pols <- c("PE", "PP", "PS")
rel.ab <- c(0.5, 0.3, 0.2)
polymer <- sample(pols, size = n, prob = rel.ab, replace = TRUE)
Data <- cbind(Data, polymer)

Data <- Data %>%  
  mutate(size.category = factor(case_when(
    Size < 10 & Size >= 1 ~ "1µm < 10µm",
    Size < 100 & Size >= 10 ~ "10µm < 100µm",
    Size < 1000 & Size >= 100 ~ "100µm < 1mm",
    Size < 5000 & Size >= 1000 ~ "1mm < 5mm"))) %>% 
  mutate(mass.mg = Density * Size^3 * 1E-9) %>% 
  mutate(um3 = (Size^3))
```

```{r}
#prediction paramaters
Data <- Data %>% 
  mutate(treatments = 4) %>% 
  mutate(organism.group = "Algae") %>% 
  mutate(exposure.duration.d = 14) %>% 
  mutate(lvl1_f = "Fitness") %>% 
  mutate(life.stage= "Early") %>% 
  mutate(bio.org = "organism") %>% 
  mutate(effect.metric = "NOEC") %>% 
  mutate(environment = "Marine") %>% 
  mutate(exposure.route = "water") %>% 
  mutate(lvl2_f = "Development")
```

```{r}
##++++++++++++++++++++++++++++++++++++++++
## Calculate Volume/Volume and estimate mass/volume
##++++++++++++++++++++++++++++++++++++++++

### 10,000 particles/L
df_1L <- Data %>% 
  mutate(vol.L = 1) %>% 
  mutate(vol.mL = vol.L * 1000) %>% 
  mutate(dose.um3.mL.master = um3 / vol.mL) %>% 
  mutate(dose.particles.mL.master = 1 / vol.mL) %>% 
  mutate(dose.mg.L.master = mass.mg / vol.L)

skim(df_1L)

### 1,000 particles/L
df_10L <- Data %>% 
  mutate(vol.L = 10) %>% 
  mutate(vol.mL = vol.L * 1000) %>% 
  mutate(dose.um3.mL.master = um3 / vol.mL) %>% 
  mutate(dose.particles.mL.master = 1 / vol.mL) %>% 
  mutate(dose.mg.L.master = mass.mg / vol.L)

### 100 particles/L
df_100L <- Data %>% 
  mutate(vol.L = 100) %>% 
  mutate(vol.mL = vol.L * 1000) %>% 
  mutate(dose.um3.mL.master = um3 / vol.mL) %>% 
  mutate(dose.particles.mL.master = 1 / vol.mL) %>% 
  mutate(dose.mg.L.master = mass.mg / vol.L)

### 10 particles/L
df_1000L <- Data %>% 
  mutate(vol.L = 1000) %>% 
  mutate(vol.mL = vol.L * 1000) %>% 
  mutate(dose.um3.mL.master = um3 / vol.mL) %>% 
  mutate(dose.particles.mL.master = 1 / vol.mL) %>% 
  mutate(dose.mg.L.master = mass.mg / vol.L)
```

```{r}
sum_1L <- df_1L %>% group_by(size.category) %>% summarize(particles.L =
                                                       sum(dose.particles.mL.master) * 1000) %>% mutate(volume.L = factor("1L"))
sum_10L <- df_10L %>% group_by(size.category) %>% summarize(particles.L =
                                                       sum(dose.particles.mL.master)* 1000) %>% mutate(volume.L = factor("10L"))
sum_100L <- df_100L %>% group_by(size.category) %>% summarize(particles.L =
                                                       sum(dose.particles.mL.master)* 1000) %>% mutate(volume.L = factor("100L"))
sum_1000L <- df_1000L %>% group_by(size.category) %>% summarize(particles.L =
                                                       sum(dose.particles.mL.master)* 1000) %>% mutate(volume.L = factor("1000L"))

summary_df <- rbind(sum_1L, sum_10L, sum_100L, sum_1000L)
ggplot() +
  geom_col(data = summary_df, aes(x = volume.L, y = particles.L, fill = volume.L)) +
  labs(title = "Particles/L in Simulated Datasets",
       subtitle = "n = 10,000 particles diluted in various liters") +
  theme_minimal(base_size = 20)
```

Parameters:
Shape, density, size, 
why do 3 polymers? Leave polymer out for now. 
For mechanistic point of view- polymer does not matter.

```{r}
##++++++++++++++++++++++++++++++++++++++++
## Quick check results
##++++++++++++++++++++++++++++++++++++++++
p1 <- ggplot(df_1000L, aes(x = Size, fill = size.category)) + geom_histogram(bins = 50, alpha = 0.7) + scale_x_log10()
p2 <- ggplot(df_1000L, aes(x = Shape, fill = polymer)) + geom_histogram(bins = 50, alpha = 0.7)
p3 <- ggplot(df_1000L, aes(x = Density, fill = polymer)) + geom_histogram(bins = 50, alpha = 0.7)
p4 <- ggplot(df_1000L, aes(x = polymer, fill = polymer)) + geom_histogram(stat = "count")
#p5 <- ggplot(df_1000L, aes(x = mass.mg, fill = polymer)) + geom_histogram(bins = 50, alpha = 0.7 )+ scale_x_log10()
#p6 <- ggplot(df_1000L, aes(x = um3, fill = polymer)) + geom_histogram(bins = 50, alpha = 0.7)+ scale_x_log10()
p7 <- ggplot(df_1000L, aes(x = dose.um3.mL.master, fill = polymer)) + geom_histogram(bins = 50, alpha = 0.7)+ scale_x_log10()

p8 <- df_1000L %>% group_by(size.category) %>% summarize(particles.mL =
                                                       sum(dose.particles.mL.master)) %>% 
  ggplot(aes(x = size.category, y = particles.mL, fill = size.category)) + geom_col()

p9 <- ggplot(df_1000L, aes(x = dose.mg.L.master, fill = polymer)) + geom_histogram(bins = 50, alpha = 0.7)+ scale_x_log10()
grid.arrange(p1,
             p2,
             p3,
             p4,
             #p5, 
             #p6,
             p7,
             p8,
             #p9,
             top = "Simulated Microplastics Dataset (n = 1e5)",
             ncol = 2)
```

#### Determine which variables are most available
```{r}
#determine which lvl2_f are most abundant for Algae data
crust <- aoc_z %>% 
  dplyr::select(c(life.stage, bio.org, effect.metric, environment, exposure.route, lvl2_f, exposure.duration.d, treatments, lvl1_f, organism.group, dose.particles.mL.master, dose.mg.L.master, effect_f, effect_10, size.length.um.used.for.conversions, density.mg.um.3, csf)) %>% 
 #filter(treatments == 4) %>% 
  filter(organism.group == "Algae") %>% 
  filter(exposure.duration.d == 14) %>% 
  filter(lvl1_f == "Fitness") %>% 
  #filter(life.stage == "Early") %>% 
  filter(bio.org == "organism") %>% 
 # filter(effect.metric == "NOEC") %>% 
  #filter(environment == "Marine") %>% 
  filter(exposure.route == "water") %>% 
 filter(lvl2_f == "Development")

skim(crust)
```

```{r eval=FALSE, include=FALSE}
crust2 <- crust %>% drop_na(effect_f) %>%  mutate_if(~is.numeric(.) && (.) > 0, log10)

#skim(crust2)
summary(glm(data = crust2, effect_10 ~ dose.particles.mL.master + dose.mg.L.master + size.length.um.used.for.conversions, na.action = "na.exclude", family = "binomial"))
```

## Predict Toxicities
```{r}
#pull in model from DALEX
final_model <- readRDS("final_rf_model.rds")

#### prep data frames ####
df_1L_f <- df_1L %>% 
  mutate_if(~is.numeric(.) && (.) > 0, log10) #model has everything log-transformed

df_10L_f <- df_10L %>% 
  mutate_if(~is.numeric(.) && (.) > 0, log10) #model has everything log-transformed

df_100L_f <- df_100L %>% 
  mutate_if(~is.numeric(.) && (.) > 0, log10) #model has everything log-transformed

df_1000L_f <- df_1000L %>% 
  mutate_if(~is.numeric(.) && (.) > 0, log10) #model has everything log-transformed

#### Predict Toxicity ####
df_1L_p <- predict(final_model, newdata = df_1L_f, type = "prob") %>%  rename("ToxicityProbability" = `1`)
df_1L_p_f <- cbind(df_1L, df_1L_p)

df_10L_p <- predict(final_model, newdata = df_10L_f, type = "prob") %>%  rename("ToxicityProbability" = `1`)
df_10L_p_f <- cbind(df_10L, df_10L_p)

df_100L_p <- predict(final_model, newdata = df_100L_f, type = "prob") %>%  rename("ToxicityProbability" = `1`)
df_100L_p_f <- cbind(df_100L, df_100L_p)

df_1000L_p <- predict(final_model, newdata = df_1000L_f, type = "prob") %>%  rename("ToxicityProbability" = `1`)
df_1000L_p_f <- cbind(df_1000L, df_1000L_p)

df_pred <- rbind(df_1L_p_f, df_10L_p_f, df_100L_p_f, df_1000L_p_f)

prob_summ <- df_pred %>% mutate(volume.L = as.factor(vol.L)) %>%  group_by(volume.L) %>% summarize(tox = sum(ToxicityProbability))

prob_summ %>% ggplot(aes(x = volume.L, y = tox, fill = volume.L)) +geom_col()

df_pred %>% ggplot(aes(x = factor(vol.L), y = ToxicityProbability)) + 
  stat_ecdf(geom = "point", size = 2, color = "blue")+
  stat_ecdf(geom = "step", linetype = 'solid', alpha = 0.6, size = 1.5) +
  #scale_x_log10() +
  scale_y_continuous(labels = scales::percent)+
  labs(title = "Cumulative Toxicity Probability",
       subtitle = "n = 1000 particles/L; Algae acute Development",
       caption = "Predicted Toxicity from Random Forest")
```

# Volume Constant, Particles different
## Generate Particles
### 5 particles in 1 L
```{r, include=FALSE}
#rm(list=ls())
n = 5
## Generate values for the three distributions
set.seed(123)

##++++++++++++++++++++++++++++++++++++++++
## SIZE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

xmin = 1 #UM
alpha = 1.6 #2.7 is preferred in unpublished Kooi et al.

X.func <- function (X){
  success <- FALSE
  while (!success){
    U = runif(1, 0, 1)
    X = xmin*(1-U)^(1/(1-alpha))
    success <- X < 5000} ##should be smaller than 5000 um 
  return(X)
}

Data <- data.frame(Size = numeric(0));

for(i in 1:n){
  X <- X.func()
  Data <- rbind(Data, X)
}

colnames(Data) <- c("Size")

min(Data$Size) ##20 um
max(Data$Size) ##5000 um

##++++++++++++++++++++++++++++++++++++++++
## SHAPE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

mu1 <- 0.08
mu2 <- 0.44
sd1 <- 0.03
sd2 <- 0.19
lambda1 <- 0.06
lambda2 <- 0.94

#Sample N random uniforms U
U =runif(n)

#Sampling from the mixture
for(i in 1:n){
  if(U[i]<lambda1){
    Data$Shape[i] = rtnorm(1,mu1,sd1, lower = 0, upper = 1)
  }else{
    Data$Shape[i] = rtnorm(1,mu2,sd2, lower = 0, upper = 1)
  }
}

min(Data$Shape) ##0
max(Data$Shape) ##1

##++++++++++++++++++++++++++++++++++++++++
## DENSITY DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

d.alpha = 73.8 #tail heaviness
d.beta = 69.9  #asymmetry
d.mu = 0.840   #location
d.delta = 0.0972 #scale

D.func <- function (D){
  success <- FALSE
  while (!success){
    D = rnig(1, mu = d.mu, alpha = d.alpha, beta = d.beta, delta = d.delta)
    success <- D < 2.63} ## include upper limit of 2.63, the max. 
  return(D)
}

Dens <- data.frame(Density = numeric(0));

for(i in 1:n){
  X <- D.func()
  Dens <- rbind(Dens, X)
}

colnames(Dens) <- c("Density")

Data <- cbind(Data, Dens)

##++++++++++++++++++++++++++++++++++++++++
## Polymer DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

pols <- c("PE", "PP", "PS")
rel.ab <- c(0.5, 0.3, 0.2)
polymer <- sample(pols, size = n, prob = rel.ab, replace = TRUE)
Data <- cbind(Data, polymer)

Data_5p <- Data %>%  
  mutate(size.category = factor(case_when(
    Size < 10 & Size >= 1 ~ "1µm < 10µm",
    Size < 100 & Size >= 10 ~ "10µm < 100µm",
    Size < 1000 & Size >= 100 ~ "100µm < 1mm",
    Size < 5000 & Size >= 1000 ~ "1mm < 5mm"))) %>% 
  mutate(mass.mg = Density * Size^3 * 1E-9) %>% 
  mutate(um3 = (Size^3)) %>% 
  mutate(treatments = 4) %>% 
  mutate(organism.group = "Algae") %>% 
  mutate(exposure.duration.d = 14) %>% 
  mutate(lvl1_f = "Fitness") %>% 
  mutate(life.stage= "Early") %>% 
  mutate(bio.org = "organism") %>% 
  mutate(effect.metric = "NOEC") %>% 
  mutate(environment = "Marine") %>% 
  mutate(exposure.route = "water") %>% 
  mutate(lvl2_f = "Development") %>% 
  mutate(vol.mL = 1) %>% 
  mutate(vol.L = vol.mL / 1000) %>% 
  mutate(dose.um3.mL.master = um3 / vol.mL) %>% 
  mutate(dose.particles.mL.master = 1 / vol.mL) %>% 
  mutate(dose.mg.L.master = mass.mg / vol.L) %>% 
  mutate(particles.total = factor("5p"))
```

### 10 particles in 1 L
```{r, include=FALSE}
#rm(list=ls())
n = 10
## Generate values for the three distributions
set.seed(123)

##++++++++++++++++++++++++++++++++++++++++
## SIZE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

xmin = 1 #UM
alpha = 1.6 #2.7 is preferred in unpublished Kooi et al.

X.func <- function (X){
  success <- FALSE
  while (!success){
    U = runif(1, 0, 1)
    X = xmin*(1-U)^(1/(1-alpha))
    success <- X < 5000} ##should be smaller than 5000 um 
  return(X)
}

Data <- data.frame(Size = numeric(0));

for(i in 1:n){
  X <- X.func()
  Data <- rbind(Data, X)
}

colnames(Data) <- c("Size")

min(Data$Size) ##20 um
max(Data$Size) ##5000 um

##++++++++++++++++++++++++++++++++++++++++
## SHAPE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

mu1 <- 0.08
mu2 <- 0.44
sd1 <- 0.03
sd2 <- 0.19
lambda1 <- 0.06
lambda2 <- 0.94

#Sample N random uniforms U
U =runif(n)

#Sampling from the mixture
for(i in 1:n){
  if(U[i]<lambda1){
    Data$Shape[i] = rtnorm(1,mu1,sd1, lower = 0, upper = 1)
  }else{
    Data$Shape[i] = rtnorm(1,mu2,sd2, lower = 0, upper = 1)
  }
}

min(Data$Shape) ##0
max(Data$Shape) ##1

##++++++++++++++++++++++++++++++++++++++++
## DENSITY DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

d.alpha = 73.8 #tail heaviness
d.beta = 69.9  #asymmetry
d.mu = 0.840   #location
d.delta = 0.0972 #scale

D.func <- function (D){
  success <- FALSE
  while (!success){
    D = rnig(1, mu = d.mu, alpha = d.alpha, beta = d.beta, delta = d.delta)
    success <- D < 2.63} ## include upper limit of 2.63, the max. 
  return(D)
}

Dens <- data.frame(Density = numeric(0));

for(i in 1:n){
  X <- D.func()
  Dens <- rbind(Dens, X)
}

colnames(Dens) <- c("Density")

Data <- cbind(Data, Dens)

##++++++++++++++++++++++++++++++++++++++++
## Polymer DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

pols <- c("PE", "PP", "PS")
rel.ab <- c(0.5, 0.3, 0.2)
polymer <- sample(pols, size = n, prob = rel.ab, replace = TRUE)
Data <- cbind(Data, polymer)

Data_10p <- Data %>%  
  mutate(size.category = factor(case_when(
    Size < 10 & Size >= 1 ~ "1µm < 10µm",
    Size < 100 & Size >= 10 ~ "10µm < 100µm",
    Size < 1000 & Size >= 100 ~ "100µm < 1mm",
    Size < 5000 & Size >= 1000 ~ "1mm < 5mm"))) %>% 
  mutate(mass.mg = Density * Size^3 * 1E-9) %>% 
  mutate(um3 = (Size^3)) %>% 
  mutate(treatments = 4) %>% 
  mutate(organism.group = "Algae") %>% 
  mutate(exposure.duration.d = 14) %>% 
  mutate(lvl1_f = "Fitness") %>% 
  mutate(life.stage= "Early") %>% 
  mutate(bio.org = "organism") %>% 
  mutate(effect.metric = "NOEC") %>% 
  mutate(environment = "Marine") %>% 
  mutate(exposure.route = "water") %>% 
  mutate(lvl2_f = "Development") %>% 
  mutate(vol.mL = 1) %>% 
  mutate(vol.L = vol.mL / 1000) %>% 
  mutate(dose.um3.mL.master = um3 / vol.mL) %>% 
  mutate(dose.particles.mL.master = 1 / vol.mL) %>% 
  mutate(dose.mg.L.master = mass.mg / vol.L) %>% 
  mutate(particles.total = factor("10p"))
```
### 20 particles in 1 L
```{r, include=FALSE}
#rm(list=ls())
n = 20
## Generate values for the three distributions
set.seed(123)

##++++++++++++++++++++++++++++++++++++++++
## SIZE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

xmin = 1 #UM
alpha = 1.6 #2.7 is preferred in unpublished Kooi et al.

X.func <- function (X){
  success <- FALSE
  while (!success){
    U = runif(1, 0, 1)
    X = xmin*(1-U)^(1/(1-alpha))
    success <- X < 5000} ##should be smaller than 5000 um 
  return(X)
}

Data <- data.frame(Size = numeric(0));

for(i in 1:n){
  X <- X.func()
  Data <- rbind(Data, X)
}

colnames(Data) <- c("Size")

min(Data$Size) ##20 um
max(Data$Size) ##5000 um

##++++++++++++++++++++++++++++++++++++++++
## SHAPE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

mu1 <- 0.08
mu2 <- 0.44
sd1 <- 0.03
sd2 <- 0.19
lambda1 <- 0.06
lambda2 <- 0.94

#Sample N random uniforms U
U =runif(n)

#Sampling from the mixture
for(i in 1:n){
  if(U[i]<lambda1){
    Data$Shape[i] = rtnorm(1,mu1,sd1, lower = 0, upper = 1)
  }else{
    Data$Shape[i] = rtnorm(1,mu2,sd2, lower = 0, upper = 1)
  }
}

min(Data$Shape) ##0
max(Data$Shape) ##1

##++++++++++++++++++++++++++++++++++++++++
## DENSITY DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

d.alpha = 73.8 #tail heaviness
d.beta = 69.9  #asymmetry
d.mu = 0.840   #location
d.delta = 0.0972 #scale

D.func <- function (D){
  success <- FALSE
  while (!success){
    D = rnig(1, mu = d.mu, alpha = d.alpha, beta = d.beta, delta = d.delta)
    success <- D < 2.63} ## include upper limit of 2.63, the max. 
  return(D)
}

Dens <- data.frame(Density = numeric(0));

for(i in 1:n){
  X <- D.func()
  Dens <- rbind(Dens, X)
}

colnames(Dens) <- c("Density")

Data <- cbind(Data, Dens)

##++++++++++++++++++++++++++++++++++++++++
## Polymer DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

pols <- c("PE", "PP", "PS")
rel.ab <- c(0.5, 0.3, 0.2)
polymer <- sample(pols, size = n, prob = rel.ab, replace = TRUE)
Data <- cbind(Data, polymer)

Data_20p <- Data %>%  
  mutate(size.category = factor(case_when(
    Size < 10 & Size >= 1 ~ "1µm < 10µm",
    Size < 100 & Size >= 10 ~ "10µm < 100µm",
    Size < 1000 & Size >= 100 ~ "100µm < 1mm",
    Size < 5000 & Size >= 1000 ~ "1mm < 5mm"))) %>% 
  mutate(mass.mg = Density * Size^3 * 1E-9) %>% 
  mutate(um3 = (Size^3)) %>% 
  mutate(treatments = 4) %>% 
  mutate(organism.group = "Algae") %>% 
  mutate(exposure.duration.d = 14) %>% 
  mutate(lvl1_f = "Fitness") %>% 
  mutate(life.stage= "Early") %>% 
  mutate(bio.org = "organism") %>% 
  mutate(effect.metric = "NOEC") %>% 
  mutate(environment = "Marine") %>% 
  mutate(exposure.route = "water") %>% 
  mutate(lvl2_f = "Development") %>% 
  mutate(vol.mL = 1) %>% 
  mutate(vol.L = vol.mL / 1000) %>% 
  mutate(dose.um3.mL.master = um3 / vol.mL) %>% 
  mutate(dose.particles.mL.master = 1 / vol.mL) %>% 
  mutate(dose.mg.L.master = mass.mg / vol.L) %>% 
  mutate(particles.total = factor("20p"))
```

### 40 particles in 1 L
```{r, include=FALSE}
#rm(list=ls())
n = 40
## Generate values for the three distributions
set.seed(123)

##++++++++++++++++++++++++++++++++++++++++
## SIZE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

xmin = 1 #UM
alpha = 1.6 #2.7 is preferred in unpublished Kooi et al.

X.func <- function (X){
  success <- FALSE
  while (!success){
    U = runif(1, 0, 1)
    X = xmin*(1-U)^(1/(1-alpha))
    success <- X < 5000} ##should be smaller than 5000 um 
  return(X)
}

Data <- data.frame(Size = numeric(0));

for(i in 1:n){
  X <- X.func()
  Data <- rbind(Data, X)
}

colnames(Data) <- c("Size")

min(Data$Size) ##20 um
max(Data$Size) ##5000 um

##++++++++++++++++++++++++++++++++++++++++
## SHAPE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

mu1 <- 0.08
mu2 <- 0.44
sd1 <- 0.03
sd2 <- 0.19
lambda1 <- 0.06
lambda2 <- 0.94

#Sample N random uniforms U
U =runif(n)

#Sampling from the mixture
for(i in 1:n){
  if(U[i]<lambda1){
    Data$Shape[i] = rtnorm(1,mu1,sd1, lower = 0, upper = 1)
  }else{
    Data$Shape[i] = rtnorm(1,mu2,sd2, lower = 0, upper = 1)
  }
}

min(Data$Shape) ##0
max(Data$Shape) ##1

##++++++++++++++++++++++++++++++++++++++++
## DENSITY DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

d.alpha = 73.8 #tail heaviness
d.beta = 69.9  #asymmetry
d.mu = 0.840   #location
d.delta = 0.0972 #scale

D.func <- function (D){
  success <- FALSE
  while (!success){
    D = rnig(1, mu = d.mu, alpha = d.alpha, beta = d.beta, delta = d.delta)
    success <- D < 2.63} ## include upper limit of 2.63, the max. 
  return(D)
}

Dens <- data.frame(Density = numeric(0));

for(i in 1:n){
  X <- D.func()
  Dens <- rbind(Dens, X)
}

colnames(Dens) <- c("Density")

Data <- cbind(Data, Dens)

##++++++++++++++++++++++++++++++++++++++++
## Polymer DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

pols <- c("PE", "PP", "PS")
rel.ab <- c(0.5, 0.3, 0.2)
polymer <- sample(pols, size = n, prob = rel.ab, replace = TRUE)
Data <- cbind(Data, polymer)

Data_40p <- Data %>%  
  mutate(size.category = factor(case_when(
    Size < 10 & Size >= 1 ~ "1µm < 10µm",
    Size < 100 & Size >= 10 ~ "10µm < 100µm",
    Size < 1000 & Size >= 100 ~ "100µm < 1mm",
    Size < 5000 & Size >= 1000 ~ "1mm < 5mm"))) %>% 
  mutate(mass.mg = Density * Size^3 * 1E-9) %>% 
  mutate(um3 = (Size^3)) %>% 
  mutate(treatments = 4) %>% 
  mutate(organism.group = "Algae") %>% 
  mutate(exposure.duration.d = 14) %>% 
  mutate(lvl1_f = "Fitness") %>% 
  mutate(life.stage= "Early") %>% 
  mutate(bio.org = "organism") %>% 
  mutate(effect.metric = "NOEC") %>% 
  mutate(environment = "Marine") %>% 
  mutate(exposure.route = "water") %>% 
  mutate(lvl2_f = "Development") %>% 
  mutate(vol.mL = 1) %>% 
  mutate(vol.L = vol.mL / 1000) %>%  
  mutate(dose.um3.mL.master = um3 / vol.mL) %>% 
  mutate(dose.particles.mL.master = 1 / vol.mL) %>% 
  mutate(dose.mg.L.master = mass.mg / vol.L) %>% 
  mutate(particles.total = factor("40p"))
```

### 80 particles in 1 L
```{r, include=FALSE}
#rm(list=ls())
n = 80
## Generate values for the three distributions
set.seed(123)

##++++++++++++++++++++++++++++++++++++++++
## SIZE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

xmin = 1 #UM
alpha = 1.6 #2.7 is preferred in unpublished Kooi et al.

X.func <- function (X){
  success <- FALSE
  while (!success){
    U = runif(1, 0, 1)
    X = xmin*(1-U)^(1/(1-alpha))
    success <- X < 5000} ##should be smaller than 5000 um 
  return(X)
}

Data <- data.frame(Size = numeric(0));

for(i in 1:n){
  X <- X.func()
  Data <- rbind(Data, X)
}

colnames(Data) <- c("Size")

min(Data$Size) ##20 um
max(Data$Size) ##5000 um

##++++++++++++++++++++++++++++++++++++++++
## SHAPE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

mu1 <- 0.08
mu2 <- 0.44
sd1 <- 0.03
sd2 <- 0.19
lambda1 <- 0.06
lambda2 <- 0.94

#Sample N random uniforms U
U =runif(n)

#Sampling from the mixture
for(i in 1:n){
  if(U[i]<lambda1){
    Data$Shape[i] = rtnorm(1,mu1,sd1, lower = 0, upper = 1)
  }else{
    Data$Shape[i] = rtnorm(1,mu2,sd2, lower = 0, upper = 1)
  }
}

min(Data$Shape) ##0
max(Data$Shape) ##1

##++++++++++++++++++++++++++++++++++++++++
## DENSITY DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

d.alpha = 73.8 #tail heaviness
d.beta = 69.9  #asymmetry
d.mu = 0.840   #location
d.delta = 0.0972 #scale

D.func <- function (D){
  success <- FALSE
  while (!success){
    D = rnig(1, mu = d.mu, alpha = d.alpha, beta = d.beta, delta = d.delta)
    success <- D < 2.63} ## include upper limit of 2.63, the max. 
  return(D)
}

Dens <- data.frame(Density = numeric(0));

for(i in 1:n){
  X <- D.func()
  Dens <- rbind(Dens, X)
}

colnames(Dens) <- c("Density")

Data <- cbind(Data, Dens)

##++++++++++++++++++++++++++++++++++++++++
## Polymer DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

pols <- c("PE", "PP", "PS")
rel.ab <- c(0.5, 0.3, 0.2)
polymer <- sample(pols, size = n, prob = rel.ab, replace = TRUE)
Data <- cbind(Data, polymer)

Data_80p <- Data %>%  
  mutate(size.category = factor(case_when(
    Size < 10 & Size >= 1 ~ "1µm < 10µm",
    Size < 100 & Size >= 10 ~ "10µm < 100µm",
    Size < 1000 & Size >= 100 ~ "100µm < 1mm",
    Size < 5000 & Size >= 1000 ~ "1mm < 5mm"))) %>% 
  mutate(mass.mg = Density * Size^3 * 1E-9) %>% 
  mutate(um3 = (Size^3)) %>% 
  mutate(treatments = 4) %>% 
  mutate(organism.group = "Algae") %>% 
  mutate(exposure.duration.d = 14) %>% 
  mutate(lvl1_f = "Fitness") %>% 
  mutate(life.stage= "Early") %>% 
  mutate(bio.org = "organism") %>% 
  mutate(effect.metric = "NOEC") %>% 
  mutate(environment = "Marine") %>% 
  mutate(exposure.route = "water") %>% 
  mutate(lvl2_f = "Development") %>% 
  mutate(vol.mL = 1) %>% 
  mutate(vol.L = vol.mL / 1000) %>% 
  mutate(dose.um3.mL.master = um3 / vol.mL) %>% 
  mutate(dose.particles.mL.master = 1 / vol.mL) %>% 
  mutate(dose.mg.L.master = mass.mg / vol.L) %>% 
  mutate(particles.total = factor("80p"))
```

### 160 particles in 1 L
```{r, include=FALSE}
#rm(list=ls())
n = 160
## Generate values for the three distributions
set.seed(123)

##++++++++++++++++++++++++++++++++++++++++
## SIZE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

xmin = 1 #UM
alpha = 1.6 #2.7 is preferred in unpublished Kooi et al.

X.func <- function (X){
  success <- FALSE
  while (!success){
    U = runif(1, 0, 1)
    X = xmin*(1-U)^(1/(1-alpha))
    success <- X < 5000} ##should be smaller than 5000 um 
  return(X)
}

Data <- data.frame(Size = numeric(0));

for(i in 1:n){
  X <- X.func()
  Data <- rbind(Data, X)
}

colnames(Data) <- c("Size")

min(Data$Size) ##20 um
max(Data$Size) ##5000 um

##++++++++++++++++++++++++++++++++++++++++
## SHAPE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

mu1 <- 0.08
mu2 <- 0.44
sd1 <- 0.03
sd2 <- 0.19
lambda1 <- 0.06
lambda2 <- 0.94

#Sample N random uniforms U
U =runif(n)

#Sampling from the mixture
for(i in 1:n){
  if(U[i]<lambda1){
    Data$Shape[i] = rtnorm(1,mu1,sd1, lower = 0, upper = 1)
  }else{
    Data$Shape[i] = rtnorm(1,mu2,sd2, lower = 0, upper = 1)
  }
}

min(Data$Shape) ##0
max(Data$Shape) ##1

##++++++++++++++++++++++++++++++++++++++++
## DENSITY DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

d.alpha = 73.8 #tail heaviness
d.beta = 69.9  #asymmetry
d.mu = 0.840   #location
d.delta = 0.0972 #scale

D.func <- function (D){
  success <- FALSE
  while (!success){
    D = rnig(1, mu = d.mu, alpha = d.alpha, beta = d.beta, delta = d.delta)
    success <- D < 2.63} ## include upper limit of 2.63, the max. 
  return(D)
}

Dens <- data.frame(Density = numeric(0));

for(i in 1:n){
  X <- D.func()
  Dens <- rbind(Dens, X)
}

colnames(Dens) <- c("Density")

Data <- cbind(Data, Dens)

##++++++++++++++++++++++++++++++++++++++++
## Polymer DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

pols <- c("PE", "PP", "PS")
rel.ab <- c(0.5, 0.3, 0.2)
polymer <- sample(pols, size = n, prob = rel.ab, replace = TRUE)
Data <- cbind(Data, polymer)

Data_160p <- Data %>%  
  mutate(size.category = factor(case_when(
    Size < 10 & Size >= 1 ~ "1µm < 10µm",
    Size < 100 & Size >= 10 ~ "10µm < 100µm",
    Size < 1000 & Size >= 100 ~ "100µm < 1mm",
    Size < 5000 & Size >= 1000 ~ "1mm < 5mm"))) %>% 
  mutate(mass.mg = Density * Size^3 * 1E-9) %>% 
  mutate(um3 = (Size^3)) %>% 
  mutate(treatments = 4) %>% 
  mutate(organism.group = "Algae") %>% 
  mutate(exposure.duration.d = 14) %>% 
  mutate(lvl1_f = "Fitness") %>% 
  mutate(life.stage= "Early") %>% 
  mutate(bio.org = "organism") %>% 
  mutate(effect.metric = "NOEC") %>% 
  mutate(environment = "Marine") %>% 
  mutate(exposure.route = "water") %>% 
  mutate(lvl2_f = "Development") %>% 
  mutate(vol.mL = 1) %>% 
  mutate(vol.L = vol.mL / 1000) %>% 
  mutate(dose.um3.mL.master = um3 / vol.mL) %>% 
  mutate(dose.particles.mL.master = 1 / vol.mL) %>% 
  mutate(dose.mg.L.master = mass.mg / vol.L) %>% 
  mutate(particles.total = factor("160p"))
```

### 320 particles in 1 L
```{r, include=FALSE}
#rm(list=ls())
n = 320
## Generate values for the three distributions
set.seed(123)

##++++++++++++++++++++++++++++++++++++++++
## SIZE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

xmin = 1 #UM
alpha = 1.6 #2.7 is preferred in unpublished Kooi et al.

X.func <- function (X){
  success <- FALSE
  while (!success){
    U = runif(1, 0, 1)
    X = xmin*(1-U)^(1/(1-alpha))
    success <- X < 5000} ##should be smaller than 5000 um 
  return(X)
}

Data <- data.frame(Size = numeric(0));

for(i in 1:n){
  X <- X.func()
  Data <- rbind(Data, X)
}

colnames(Data) <- c("Size")

min(Data$Size) ##20 um
max(Data$Size) ##5000 um

##++++++++++++++++++++++++++++++++++++++++
## SHAPE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

mu1 <- 0.08
mu2 <- 0.44
sd1 <- 0.03
sd2 <- 0.19
lambda1 <- 0.06
lambda2 <- 0.94

#Sample N random uniforms U
U =runif(n)

#Sampling from the mixture
for(i in 1:n){
  if(U[i]<lambda1){
    Data$Shape[i] = rtnorm(1,mu1,sd1, lower = 0, upper = 1)
  }else{
    Data$Shape[i] = rtnorm(1,mu2,sd2, lower = 0, upper = 1)
  }
}

min(Data$Shape) ##0
max(Data$Shape) ##1

##++++++++++++++++++++++++++++++++++++++++
## DENSITY DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

d.alpha = 73.8 #tail heaviness
d.beta = 69.9  #asymmetry
d.mu = 0.840   #location
d.delta = 0.0972 #scale

D.func <- function (D){
  success <- FALSE
  while (!success){
    D = rnig(1, mu = d.mu, alpha = d.alpha, beta = d.beta, delta = d.delta)
    success <- D < 2.63} ## include upper limit of 2.63, the max. 
  return(D)
}

Dens <- data.frame(Density = numeric(0));

for(i in 1:n){
  X <- D.func()
  Dens <- rbind(Dens, X)
}

colnames(Dens) <- c("Density")

Data <- cbind(Data, Dens)

##++++++++++++++++++++++++++++++++++++++++
## Polymer DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

pols <- c("PE", "PP", "PS")
rel.ab <- c(0.5, 0.3, 0.2)
polymer <- sample(pols, size = n, prob = rel.ab, replace = TRUE)
Data <- cbind(Data, polymer)

Data_320p <- Data %>%  
  mutate(size.category = factor(case_when(
    Size < 10 & Size >= 1 ~ "1µm < 10µm",
    Size < 100 & Size >= 10 ~ "10µm < 100µm",
    Size < 1000 & Size >= 100 ~ "100µm < 1mm",
    Size < 5000 & Size >= 1000 ~ "1mm < 5mm"))) %>% 
  mutate(mass.mg = Density * Size^3 * 1E-9) %>% 
  mutate(um3 = (Size^3)) %>% 
  mutate(treatments = 4) %>% 
  mutate(organism.group = "Algae") %>% 
  mutate(exposure.duration.d = 14) %>% 
  mutate(lvl1_f = "Fitness") %>% 
  mutate(life.stage= "Early") %>% 
  mutate(bio.org = "organism") %>% 
  mutate(effect.metric = "NOEC") %>% 
  mutate(environment = "Marine") %>% 
  mutate(exposure.route = "water") %>% 
  mutate(lvl2_f = "Development") %>% 
 mutate(vol.mL = 1) %>% 
  mutate(vol.L = vol.mL / 1000) %>% 
  mutate(dose.um3.mL.master = um3 / vol.mL) %>% 
  mutate(dose.particles.mL.master = 1 / vol.mL) %>% 
  mutate(dose.mg.L.master = mass.mg / vol.L) %>% 
  mutate(particles.total = factor("320p"))
```

### 480 particles in 1 L
```{r, include=FALSE}
#rm(list=ls())
n = 480
## Generate values for the three distributions
set.seed(123)

##++++++++++++++++++++++++++++++++++++++++
## SIZE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

xmin = 1 #UM
alpha = 1.6 #2.7 is preferred in unpublished Kooi et al.

X.func <- function (X){
  success <- FALSE
  while (!success){
    U = runif(1, 0, 1)
    X = xmin*(1-U)^(1/(1-alpha))
    success <- X < 5000} ##should be smaller than 5000 um 
  return(X)
}

Data <- data.frame(Size = numeric(0));

for(i in 1:n){
  X <- X.func()
  Data <- rbind(Data, X)
}

colnames(Data) <- c("Size")

min(Data$Size) ##20 um
max(Data$Size) ##5000 um

##++++++++++++++++++++++++++++++++++++++++
## SHAPE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

mu1 <- 0.08
mu2 <- 0.44
sd1 <- 0.03
sd2 <- 0.19
lambda1 <- 0.06
lambda2 <- 0.94

#Sample N random uniforms U
U =runif(n)

#Sampling from the mixture
for(i in 1:n){
  if(U[i]<lambda1){
    Data$Shape[i] = rtnorm(1,mu1,sd1, lower = 0, upper = 1)
  }else{
    Data$Shape[i] = rtnorm(1,mu2,sd2, lower = 0, upper = 1)
  }
}

min(Data$Shape) ##0
max(Data$Shape) ##1

##++++++++++++++++++++++++++++++++++++++++
## DENSITY DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

d.alpha = 73.8 #tail heaviness
d.beta = 69.9  #asymmetry
d.mu = 0.840   #location
d.delta = 0.0972 #scale

D.func <- function (D){
  success <- FALSE
  while (!success){
    D = rnig(1, mu = d.mu, alpha = d.alpha, beta = d.beta, delta = d.delta)
    success <- D < 2.63} ## include upper limit of 2.63, the max. 
  return(D)
}

Dens <- data.frame(Density = numeric(0));

for(i in 1:n){
  X <- D.func()
  Dens <- rbind(Dens, X)
}

colnames(Dens) <- c("Density")

Data <- cbind(Data, Dens)

##++++++++++++++++++++++++++++++++++++++++
## Polymer DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

pols <- c("PE", "PP", "PS")
rel.ab <- c(0.5, 0.3, 0.2)
polymer <- sample(pols, size = n, prob = rel.ab, replace = TRUE)
Data <- cbind(Data, polymer)

Data_480p <- Data %>%  
  mutate(size.category = factor(case_when(
    Size < 10 & Size >= 1 ~ "1µm < 10µm",
    Size < 100 & Size >= 10 ~ "10µm < 100µm",
    Size < 1000 & Size >= 100 ~ "100µm < 1mm",
    Size < 5000 & Size >= 1000 ~ "1mm < 5mm"))) %>% 
  mutate(mass.mg = Density * Size^3 * 1E-9) %>% 
  mutate(um3 = (Size^3)) %>% 
  mutate(treatments = 4) %>% 
  mutate(organism.group = "Algae") %>% 
  mutate(exposure.duration.d = 14) %>% 
  mutate(lvl1_f = "Fitness") %>% 
  mutate(life.stage= "Early") %>% 
  mutate(bio.org = "organism") %>% 
  mutate(effect.metric = "NOEC") %>% 
  mutate(environment = "Marine") %>% 
  mutate(exposure.route = "water") %>% 
  mutate(lvl2_f = "Development") %>% 
 mutate(vol.mL = 1) %>% 
  mutate(vol.L = vol.mL / 1000) %>% 
  mutate(dose.um3.mL.master = um3 / vol.mL) %>% 
  mutate(dose.particles.mL.master = 1 / vol.mL) %>% 
  mutate(dose.mg.L.master = mass.mg / vol.L) %>% 
  mutate(particles.total = factor("480p"))
```

### 640 particles in 1 L
```{r, include=FALSE}
#rm(list=ls())
n = 640
## Generate values for the three distributions
set.seed(123)

##++++++++++++++++++++++++++++++++++++++++
## SIZE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

xmin = 1 #UM
alpha = 1.6 #2.7 is preferred in unpublished Kooi et al.

X.func <- function (X){
  success <- FALSE
  while (!success){
    U = runif(1, 0, 1)
    X = xmin*(1-U)^(1/(1-alpha))
    success <- X < 5000} ##should be smaller than 5000 um 
  return(X)
}

Data <- data.frame(Size = numeric(0));

for(i in 1:n){
  X <- X.func()
  Data <- rbind(Data, X)
}

colnames(Data) <- c("Size")

min(Data$Size) ##20 um
max(Data$Size) ##5000 um

##++++++++++++++++++++++++++++++++++++++++
## SHAPE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

mu1 <- 0.08
mu2 <- 0.44
sd1 <- 0.03
sd2 <- 0.19
lambda1 <- 0.06
lambda2 <- 0.94

#Sample N random uniforms U
U =runif(n)

#Sampling from the mixture
for(i in 1:n){
  if(U[i]<lambda1){
    Data$Shape[i] = rtnorm(1,mu1,sd1, lower = 0, upper = 1)
  }else{
    Data$Shape[i] = rtnorm(1,mu2,sd2, lower = 0, upper = 1)
  }
}

min(Data$Shape) ##0
max(Data$Shape) ##1

##++++++++++++++++++++++++++++++++++++++++
## DENSITY DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

d.alpha = 73.8 #tail heaviness
d.beta = 69.9  #asymmetry
d.mu = 0.840   #location
d.delta = 0.0972 #scale

D.func <- function (D){
  success <- FALSE
  while (!success){
    D = rnig(1, mu = d.mu, alpha = d.alpha, beta = d.beta, delta = d.delta)
    success <- D < 2.63} ## include upper limit of 2.63, the max. 
  return(D)
}

Dens <- data.frame(Density = numeric(0));

for(i in 1:n){
  X <- D.func()
  Dens <- rbind(Dens, X)
}

colnames(Dens) <- c("Density")

Data <- cbind(Data, Dens)

##++++++++++++++++++++++++++++++++++++++++
## Polymer DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

pols <- c("PE", "PP", "PS")
rel.ab <- c(0.5, 0.3, 0.2)
polymer <- sample(pols, size = n, prob = rel.ab, replace = TRUE)
Data <- cbind(Data, polymer)

Data_640p <- Data %>%  
  mutate(size.category = factor(case_when(
    Size < 10 & Size >= 1 ~ "1µm < 10µm",
    Size < 100 & Size >= 10 ~ "10µm < 100µm",
    Size < 1000 & Size >= 100 ~ "100µm < 1mm",
    Size < 5000 & Size >= 1000 ~ "1mm < 5mm"))) %>% 
  mutate(mass.mg = Density * Size^3 * 1E-9) %>% 
  mutate(um3 = (Size^3)) %>% 
  mutate(treatments = 4) %>% 
  mutate(organism.group = "Algae") %>% 
  mutate(exposure.duration.d = 14) %>% 
  mutate(lvl1_f = "Fitness") %>% 
  mutate(life.stage= "Early") %>% 
  mutate(bio.org = "organism") %>% 
  mutate(effect.metric = "NOEC") %>% 
  mutate(environment = "Marine") %>% 
  mutate(exposure.route = "water") %>% 
  mutate(lvl2_f = "Development") %>% 
 mutate(vol.mL = 1) %>% 
  mutate(vol.L = vol.mL / 1000) %>%  
  mutate(dose.um3.mL.master = um3 / vol.mL) %>% 
  mutate(dose.particles.mL.master = 1 / vol.mL) %>% 
  mutate(dose.mg.L.master = mass.mg / vol.L) %>% 
  mutate(particles.total = factor("640p"))
```

### 1000 particles
```{r, include=FALSE}
#rm(list=ls())
n = 1000
## Generate values for the three distributions
set.seed(123)

##++++++++++++++++++++++++++++++++++++++++
## SIZE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

xmin = 1 #UM
alpha = 1.6 #2.7 is preferred in unpublished Kooi et al.

X.func <- function (X){
  success <- FALSE
  while (!success){
    U = runif(1, 0, 1)
    X = xmin*(1-U)^(1/(1-alpha))
    success <- X < 5000} ##should be smaller than 5000 um 
  return(X)
}

Data <- data.frame(Size = numeric(0));

for(i in 1:n){
  X <- X.func()
  Data <- rbind(Data, X)
}

colnames(Data) <- c("Size")

min(Data$Size) ##20 um
max(Data$Size) ##5000 um

##++++++++++++++++++++++++++++++++++++++++
## SHAPE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

mu1 <- 0.08
mu2 <- 0.44
sd1 <- 0.03
sd2 <- 0.19
lambda1 <- 0.06
lambda2 <- 0.94

#Sample N random uniforms U
U =runif(n)

#Sampling from the mixture
for(i in 1:n){
  if(U[i]<lambda1){
    Data$Shape[i] = rtnorm(1,mu1,sd1, lower = 0, upper = 1)
  }else{
    Data$Shape[i] = rtnorm(1,mu2,sd2, lower = 0, upper = 1)
  }
}

min(Data$Shape) ##0
max(Data$Shape) ##1

##++++++++++++++++++++++++++++++++++++++++
## DENSITY DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

d.alpha = 73.8 #tail heaviness
d.beta = 69.9  #asymmetry
d.mu = 0.840   #location
d.delta = 0.0972 #scale

D.func <- function (D){
  success <- FALSE
  while (!success){
    D = rnig(1, mu = d.mu, alpha = d.alpha, beta = d.beta, delta = d.delta)
    success <- D < 2.63} ## include upper limit of 2.63, the max. 
  return(D)
}

Dens <- data.frame(Density = numeric(0));

for(i in 1:n){
  X <- D.func()
  Dens <- rbind(Dens, X)
}

colnames(Dens) <- c("Density")

Data <- cbind(Data, Dens)

##++++++++++++++++++++++++++++++++++++++++
## Polymer DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

pols <- c("PE", "PP", "PS")
rel.ab <- c(0.5, 0.3, 0.2)
polymer <- sample(pols, size = n, prob = rel.ab, replace = TRUE)
Data <- cbind(Data, polymer)

Data_1000p <- Data %>%  
  mutate(size.category = factor(case_when(
    Size < 10 & Size >= 1 ~ "1µm < 10µm",
    Size < 100 & Size >= 10 ~ "10µm < 100µm",
    Size < 1000 & Size >= 100 ~ "100µm < 1mm",
    Size < 5000 & Size >= 1000 ~ "1mm < 5mm"))) %>% 
  mutate(mass.mg = Density * Size^3 * 1E-9) %>% 
  mutate(um3 = (Size^3)) %>% 
  mutate(treatments = 4) %>% 
  mutate(organism.group = "Algae") %>% 
  mutate(exposure.duration.d = 14) %>% 
  mutate(lvl1_f = "Fitness") %>% 
  mutate(life.stage= "Early") %>% 
  mutate(bio.org = "organism") %>% 
  mutate(effect.metric = "NOEC") %>% 
  mutate(environment = "Marine") %>% 
  mutate(exposure.route = "water") %>% 
  mutate(lvl2_f = "Development") %>% 
 mutate(vol.mL = 1) %>% 
  mutate(vol.L = vol.mL / 1000) %>% 
  mutate(dose.um3.mL.master = um3 / vol.mL) %>% 
  mutate(dose.particles.mL.master = 1 / vol.mL) %>% 
  mutate(dose.mg.L.master = mass.mg / vol.L) %>% 
  mutate(particles.total = factor("1000p"))
```

### 1280 particles
```{r, include=FALSE}
#rm(list=ls())
n = 1280
## Generate values for the three distributions
set.seed(123)

##++++++++++++++++++++++++++++++++++++++++
## SIZE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

xmin = 1 #UM
alpha = 1.6 #2.7 is preferred in unpublished Kooi et al.

X.func <- function (X){
  success <- FALSE
  while (!success){
    U = runif(1, 0, 1)
    X = xmin*(1-U)^(1/(1-alpha))
    success <- X < 5000} ##should be smaller than 5000 um 
  return(X)
}

Data <- data.frame(Size = numeric(0));

for(i in 1:n){
  X <- X.func()
  Data <- rbind(Data, X)
}

colnames(Data) <- c("Size")

min(Data$Size) ##20 um
max(Data$Size) ##5000 um

##++++++++++++++++++++++++++++++++++++++++
## SHAPE DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

mu1 <- 0.08
mu2 <- 0.44
sd1 <- 0.03
sd2 <- 0.19
lambda1 <- 0.06
lambda2 <- 0.94

#Sample N random uniforms U
U =runif(n)

#Sampling from the mixture
for(i in 1:n){
  if(U[i]<lambda1){
    Data$Shape[i] = rtnorm(1,mu1,sd1, lower = 0, upper = 1)
  }else{
    Data$Shape[i] = rtnorm(1,mu2,sd2, lower = 0, upper = 1)
  }
}

min(Data$Shape) ##0
max(Data$Shape) ##1

##++++++++++++++++++++++++++++++++++++++++
## DENSITY DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

d.alpha = 73.8 #tail heaviness
d.beta = 69.9  #asymmetry
d.mu = 0.840   #location
d.delta = 0.0972 #scale

D.func <- function (D){
  success <- FALSE
  while (!success){
    D = rnig(1, mu = d.mu, alpha = d.alpha, beta = d.beta, delta = d.delta)
    success <- D < 2.63} ## include upper limit of 2.63, the max. 
  return(D)
}

Dens <- data.frame(Density = numeric(0));

for(i in 1:n){
  X <- D.func()
  Dens <- rbind(Dens, X)
}

colnames(Dens) <- c("Density")

Data <- cbind(Data, Dens)

##++++++++++++++++++++++++++++++++++++++++
## Polymer DISTRIBUTION
##++++++++++++++++++++++++++++++++++++++++

pols <- c("PE", "PP", "PS")
rel.ab <- c(0.5, 0.3, 0.2)
polymer <- sample(pols, size = n, prob = rel.ab, replace = TRUE)
Data <- cbind(Data, polymer)

Data_1280p <- Data %>%  
  mutate(size.category = factor(case_when(
    Size < 10 & Size >= 1 ~ "1µm < 10µm",
    Size < 100 & Size >= 10 ~ "10µm < 100µm",
    Size < 1000 & Size >= 100 ~ "100µm < 1mm",
    Size < 5000 & Size >= 1000 ~ "1mm < 5mm"))) %>% 
  mutate(mass.mg = Density * Size^3 * 1E-9) %>% 
  mutate(um3 = (Size^3)) %>% 
  mutate(treatments = 4) %>% 
  mutate(organism.group = "Algae") %>% 
  mutate(exposure.duration.d = 14) %>% 
  mutate(lvl1_f = "Fitness") %>% 
  mutate(life.stage= "Early") %>% 
  mutate(bio.org = "organism") %>% 
  mutate(effect.metric = "NOEC") %>% 
  mutate(environment = "Marine") %>% 
  mutate(exposure.route = "water") %>% 
  mutate(lvl2_f = "Development") %>% 
 mutate(vol.mL = 1) %>% 
  mutate(vol.L = vol.mL / 1000) %>% 
  mutate(dose.um3.mL.master = um3 / vol.mL) %>% 
  mutate(dose.particles.mL.master = 1 / vol.mL) %>% 
  mutate(dose.mg.L.master = mass.mg / vol.L) %>% 
  mutate(particles.total = factor("1280p"))
```

Parameters:
Shape, density, size, 
why do 3 polymers? Leave polymer out for now. 
For mechanistic point of view- polymer does not matter.

## Predict Toxicities
```{r}
#pull in model from DALEX
final_model <- readRDS("final_rf_model.rds")

#### prep data frames ####
df_5p_f <- Data_5p %>% 
  mutate_if(~is.numeric(.) && (.) > 0, log10) #model has everything log-transformed

df_10p_f <- Data_10p %>% 
  mutate_if(~is.numeric(.) && (.) > 0, log10) #model has everything log-transformed

df_20p_f <- Data_20p %>% 
  mutate_if(~is.numeric(.) && (.) > 0, log10) #model has everything log-transformed

df_40p_f <- Data_40p %>% 
  mutate_if(~is.numeric(.) && (.) > 0, log10) #model has everything log-transformed

df_80p_f <- Data_80p %>% 
  mutate_if(~is.numeric(.) && (.) > 0, log10) #model has everything log-transformed

df_160p_f <- Data_160p %>% 
  mutate_if(~is.numeric(.) && (.) > 0, log10) #model has everything log-transformed

df_320p_f <- Data_320p %>% 
  mutate_if(~is.numeric(.) && (.) > 0, log10) #model has everything log-transformed

df_480p_f <- Data_480p %>% 
  mutate_if(~is.numeric(.) && (.) > 0, log10) #model has everything log-transformed

df_640p_f <- Data_640p %>% 
  mutate_if(~is.numeric(.) && (.) > 0, log10) #model has everything log-transformed

df_1000p_f <- Data_1000p %>% 
  mutate_if(~is.numeric(.) && (.) > 0, log10) #model has everything log-transformed

df_1280p_f <- Data_1280p %>% 
  mutate_if(~is.numeric(.) && (.) > 0, log10) #model has everything log-transformed

#### Predict Toxicity ####
df_5p_p <- predict(final_model, newdata = df_5p_f, type = "prob") %>%  rename("ToxicityProbability" = `1`)
df_5p_p_f <- cbind(df_5p_f, df_5p_p)

df_10p_p <- predict(final_model, newdata = df_10p_f, type = "prob") %>%  rename("ToxicityProbability" = `1`)
df_10p_p_f <- cbind(df_10p_f, df_10p_p)

df_20p_p <- predict(final_model, newdata = df_20p_f, type = "prob") %>%  rename("ToxicityProbability" = `1`)
df_20p_p_f <- cbind(df_20p_f, df_20p_p)

df_40p_p <- predict(final_model, newdata = df_40p_f, type = "prob") %>%  rename("ToxicityProbability" = `1`)
df_40p_p_f <- cbind(df_40p_f, df_40p_p)

df_10p_p <- predict(final_model, newdata = df_10p_f, type = "prob") %>%  rename("ToxicityProbability" = `1`)
df_10p_p_f <- cbind(df_10p_f, df_10p_p)

df_80p_p <- predict(final_model, newdata = df_80p_f, type = "prob") %>%  rename("ToxicityProbability" = `1`)
df_80p_p_f <- cbind(df_80p_f, df_80p_p)

df_160p_p <- predict(final_model, newdata = df_160p_f, type = "prob") %>%  rename("ToxicityProbability" = `1`)
df_160p_p_f <- cbind(df_160p_f, df_160p_p)

df_320p_p <- predict(final_model, newdata = df_320p_f, type = "prob") %>%  rename("ToxicityProbability" = `1`)
df_320p_p_f <- cbind(df_320p_f, df_320p_p)

df_480p_p <- predict(final_model, newdata = df_480p_f, type = "prob") %>%  rename("ToxicityProbability" = `1`)
df_480p_p_f <- cbind(df_480p_f, df_480p_p)

df_640p_p <- predict(final_model, newdata = df_640p_f, type = "prob") %>%  rename("ToxicityProbability" = `1`)
df_640p_p_f <- cbind(df_640p_f, df_640p_p)

df_1000p_p <- predict(final_model, newdata = df_1000p_f, type = "prob") %>%  rename("ToxicityProbability" = `1`)
df_1000p_p_f <- cbind(df_1000p_f, df_1000p_p)

df_1280p_p <- predict(final_model, newdata = df_1280p_f, type = "prob") %>%  rename("ToxicityProbability" = `1`)
df_1280p_p_f <- cbind(df_1280p_f, df_1280p_p)

df_pred <- rbind(df_5p_p_f, df_10p_p_f, df_20p_p_f, df_40p_p_f, df_80p_p_f, df_160p_p_f, df_320p_p_f, df_480p_p_f, df_640p_p_f, df_1000p_p_f, df_1280p_p_f)

prob_summ <- df_pred %>% group_by(particles.total) %>% summarize(tox = sum(ToxicityProbability))

####convert >1 prob to 1 ###
prob_summ <- prob_summ %>%  mutate(probability = case_when(tox < 1 ~ tox, tox >=1 ~ 1)) %>% 
  mutate(particles = as.numeric(str_sub(particles.total, end = -2))) %>% 
  mutate(log.particles = log(particles))

#plot dose response relationship
prob_summ %>% 
  ggplot(aes(x = particles, y = probability)) +
  geom_point() + 
  stat_smooth(method = drm, method.args = list(fct = LL.4()), se = FALSE) + #logit-4 paramater curve
  scale_x_log10(name = "Particles/mL (1-5,000 um)") +
  scale_y_continuous(labels = scales::percent_format(),
                     name = "Probability of NOEC") +
  labs(title = "Model-Predicted Probability of Toxicity",
       subtitle = "Algae; 14-d exp; Development; NOEC; Early; Marine",
       caption = "Random Forest Model; Simulated particle distribution (alpha = 1.6)") +
  theme_minimal(base_size = 14) +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

# df_pred %>% 
#   group_by(particles.total) %>% 
#   summarize(tox_n = count(ToxicityProbability)) %>% 
#   mutate(cum_tox = cumsum(tox_n)) %>% 
#   ggplot(aes(x = Size, y = cum_tox)) + 
#   geom_step() +
#   facet_wrap(~particles.total)

  #scale_x_log10() +
  # scale_y_continuous(labels = scales::percent)+
  # labs(title = "Cumulative Toxicity Probability",
  #      subtitle = "n = 1000 particles/mL; Algae acute Development",
  #      caption = "Predicted Toxicity from Random Forest")
```

```{r}
dose_response <- drm(data = prob_summ, probability ~ log.particles, fct = LL.4())
summary(dose_response)
```

```{r}
summary(glm(ToxicityProbability ~ Size * Shape * polymer * um3, data = df_640p_p_f))
p1 <- df_640p_p_f %>% ggplot(aes(x = 10^Size, y = ToxicityProbability, color = polymer)) + geom_point()

p2 <- df_640p_p_f %>% ggplot(aes(x = ToxicityProbability, fill = size.category)) + geom_histogram()
grid.arrange(p1,p2)
```

```{r}
#test toxicity prediction with a single row

dataFrame <- data.frame(Doubles=double(),
                 Ints=integer(),
                 Factors=factor(),
                 Logicals=logical(),
                 Characters=character(),
                 stringsAsFactors=FALSE)
newrow =data.frame(dose.particles.mL.master = 1)
# Compare with monodisperse prediction
monoDisperse_100 <- newrow %>%  
 # mutate(dose.particles.mL.master = 120) %>%  #120 particles/mL is about EC50 for above
  mutate(mass.mg = 0.2263657 * 120) %>% #average mass x 120 particles
  mutate(um3 = 242182027) %>% 
  mutate(treatments = 4) %>% 
  mutate(organism.group = "Algae") %>% 
  mutate(exposure.duration.d = 14) %>% 
  mutate(lvl1_f = "Fitness") %>% 
  mutate(life.stage= "Early") %>% 
  mutate(bio.org = "organism") %>% 
  mutate(effect.metric = "NOEC") %>% 
  mutate(environment = "Marine") %>% 
  mutate(exposure.route = "water") %>% 
  mutate(lvl2_f = "Development") %>% 
  mutate(vol.mL = 1) %>% 
  mutate(vol.L = vol.mL / 1000) %>% 
  mutate(dose.um3.mL.master = um3 / vol.mL) %>% 
  mutate(dose.mg.L.master = mass.mg / vol.L) %>% 
  mutate(particles.total = factor("120p"))

predict(final_model, newdata = monoDisperse_100, type = "prob")
```
### Uniform Distribution

It may be possible to visualize discrete relationships of toxicity by creating a uniform distribution of particles within the parameter space of the training dataset, then plotting the predicted toxicities for different predictors.

```{r}
###### determine bounds for training dataset #####
##NOTE: WILL NOT WORK IN MARKDOWN UNLESS WE PULL IN CODE FOR AOC FROM OTHER SCRIPT ###

#first extract 
particleCharacteristics <- aoc_z %>% 
  dplyr::select(c(mass.per.particle.mg, polymer, shape, density.mg.um.3, size.length.um.used.for.conversions,
                  particle.volume.um3, csf))

skim(particleCharacteristics)
```

#### Generate particles
```{r}

n = 10000

#Create uniform characteristics for base particle characteristics (size, shape, density)
csf <- runif(n, min = 0.01, 0.99)
#mass.per.particle.mg <- runif(n, min = 0, max = 1.65e+1)
density.mg.um.3 <- runif(n, min = 8.75e-10, max = 1.53e-9)
size.length.um.used.for.conversions <- runif(n, min = 1, max = 1000)

#create uniform distribution of polymer types
pols <- c("PE", "PP", "PS", "PVC", "PET")
rel.ab <- c(0.2, 0.2, 0.2, 0.2, 0.2)
polymer <- sample(pols, size = n, prob = rel.ab, replace = TRUE)

#create data frame with uniform distirbutions
uniform <- data.frame(cbind(csf, density.mg.um.3, size.length.um.used.for.conversions)) %>%
  mutate_all(funs(as.numeric(as.character(.))))
#bind polymer and convert to factor
uniform <- cbind(uniform, polymer) %>% mutate_if(is.character, as.factor)

#calculate higher order parameters and assign organism traits
unif_df <- uniform %>%  
  #particle characteristics
  mutate(size.category = factor(case_when(
    size.length.um.used.for.conversions < 10 & size.length.um.used.for.conversions >= 1 ~ "1µm < 10µm",
    size.length.um.used.for.conversions < 100 & size.length.um.used.for.conversions >= 10 ~ "10µm < 100µm",
    size.length.um.used.for.conversions < 1000 & size.length.um.used.for.conversions >= 100 ~ "100µm < 1mm",
    size.length.um.used.for.conversions < 5000 & size.length.um.used.for.conversions >= 1000 ~ "1mm < 5mm"))) %>% 
  mutate(mass.per.particle.mg = density.mg.um.3 * size.length.um.used.for.conversions ^ 3 * 1E-9) %>% 
  mutate(particle.volume.um3 = pi/6 * (size.length.um.used.for.conversions ^ 3) * csf ^2 ) %>% #equation 4
  #volume
  mutate(vol.mL = 0.0001) %>% 
  mutate(vol.L = vol.mL / 1000) %>%  
  mutate(dose.um3.mL.master = particle.volume.um3 / vol.mL) %>% 
  mutate(dose.particles.mL.master = 1 / vol.mL) %>% 
  mutate(dose.mg.L.master = mass.per.particle.mg / vol.L) %>% 
    #organism/test system characteristics
  mutate(treatments = 4) %>% 
  mutate(organism.group = "Algae") %>% 
  mutate(exposure.duration.d = 14) %>% 
  mutate(lvl1_f = "Fitness") %>% 
  mutate(life.stage= "Early") %>% 
  mutate(bio.org = "organism") %>% 
  mutate(effect.metric = "NOEC") %>% 
  mutate(environment = "Marine") %>% 
  mutate(exposure.route = "water") %>% 
  mutate(lvl2_f = "Development") %>% 
  mutate_if(is.character, as.factor)

skim(unif_df)
```

##### Inspect distributions of calculated parameters
```{r}
#create a 2 x 3 plotting matrix
par(mfrow = c(3,3))
#plot
hist(unif_df$csf)
hist(unif_df$size.length.um.used.for.conversions)
hist(unif_df$density.mg.um.3)
hist(unif_df$dose.mg.L.master)
hist(unif_df$dose.particles.mL.master)
hist(unif_df$dose.um3.mL.master)
hist(unif_df$mass.per.particle.mg)
hist(unif_df$particle.volume.um3)
#ggplot(unif_df, aes(x = size.category, y =..count..)) + geom_histogram(stat = "count")
```

##### Predict toxicities on uniform distribution of particles
```{r}
#log transform numeric variables first
unif_df <- unif_df %>%  mutate_if(~is.numeric(.) && (.) > 0, log10) #model has everything log-transformed

#predict tox
unif_pred <- predict(final_model, newdata = unif_df, type = "prob") %>%  rename("ToxicityProbability" = `1`)
uniform_pred <- cbind(unif_pred, unif_df)

#summary(glm(ToxicityProbability ~ Size * Shape * polymer * um3, data = df_640p_p_f))
p1 <- uniform_pred %>% ggplot(aes(x = dose.particles.mL.master, y = ToxicityProbability, color = polymer)) + geom_point()

p2 <- uniform_pred %>% ggplot(aes(x = dose.mg.L.master, y = ToxicityProbability, color = polymer)) + geom_point()

p3 <- uniform_pred %>% ggplot(aes(x = dose.um3.mL.master, y = ToxicityProbability, color = polymer)) + geom_point()

p4 <- uniform_pred %>% ggplot(aes(x = ToxicityProbability, fill = polymer)) + geom_histogram()

p5 <- uniform_pred %>%  ggplot(aes(x = size.length.um.used.for.conversions, y = ToxicityProbability, color = csf)) + geom_point()

p6 <- uniform_pred %>%  ggplot(aes(x = csf, y = ToxicityProbability, color = size.length.um.used.for.conversions)) + geom_point()

p7 <- uniform_pred %>%  ggplot(aes(x = particle.volume.um3, y = ToxicityProbability, color =
                                     size.length.um.used.for.conversions)) + geom_point()

p8 <- uniform_pred %>%  ggplot(aes(x = mass.per.particle.mg, y = ToxicityProbability, color =
                                     size.length.um.used.for.conversions)) + geom_point()

p9 <- uniform_pred %>%  ggplot(aes(x = density.mg.um.3, y = ToxicityProbability, color =
                                     size.length.um.used.for.conversions)) + geom_point()

grid.arrange(p1, p2, p3, p4, ncol = 2)
```
```{r}
grid.arrange(p5, p6, p7, p8, p9, ncol = 2)
```
```{r}
summary(lm(ToxicityProbability ~ size.length.um.used.for.conversions * csf * particle.volume.um3, data = uniform_pred))
```

HIGHLIGHT AREA UNDER THE CURVE
If we go down to lower number, we would not adequately mimic distribution. Stick to creating 1E5 particles, then change volume. 
Could capture variability for small bumbers by simulating small number of particles

How to calculate uncertainties? 
  Replace effect parameters by median +- SD, then predict tox for each distribution, plot, then if it is minor, FORGET about it! REsiduals in Merel's models were not randomly distributed -  so imperfect models.

What sizes should be used?
  1-5,000 um.

Wait for Merel to update interactions between shape/size/density/count.

What about max size ingest?
  Tox studies typically do not exposure animals to particles bigger than their mouth size opening, so model may not understand that relationship. 
  Could impose bioavailable fraction AFTER model. JUST do predictions for species that have bioavailable fraction data. Could estimate mouth size from prey to predator fraction. JAMS et al has mouth-size opening parameters.

Non-linear regression to get dose-response.

Mattson, Hasselof, Frontiers in Science - measures size distribution.

Final figure: SINGLE toxicity prediction for each dose. Use environmental data. Make SSD. Risk Characterization. The model's SSD is fundamentally different from the SSD made from monodisperse particles, so it is MORE reliable. 
Big assumption: all particle toxicities are ADDITIVE! Depends on effect mechanism - food dilution.

# Dose-response curves

# SSD

Create simulated SSD
Test time.

BOOTSTRAP to see if it fits the same each time.

TRY DIFFERENT MODEL TO PROVE THAT OVERFITTING IS NOT OCCURING.

# Sensitivity Analysis
it would be great to use the full approach to find out which model parameters and MP subclasses the HC5 is most sensitive to.And also if that HC5 is sensitive to which of Merel’s compartment-specific MP parameterisations is used.

# Shiny App Tabs
1) ML predictor
2) Chemical data
  Bioaccumulation model, partitioning with chemicals/particles. Would add toxicity from chemicals to toxicity from particles! Use same model from Worm paper. Would show how it would compare with toxicities from particles. We have criticial body burden concept (lipids)